### 2021-02-17 README.Rmd
Added README file.

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# yowie

<!-- badges: start -->
[![R-CMD-check](https://github.com/numbats/yowie/workflows/R-CMD-check/badge.svg)](https://github.com/numbats/yowie/actions)
<!-- badges: end -->

The goal of yowie is to provide longitudinal wages data sets along with several demographic variables from The National Longitudinal Survey of Youth (NLSY) from survey year 1979 to 2018. There are three data sets in this package:

-  `wages_hs2020`: The wages data from the cohort whose highest grade completed is up to 12th grade.
-  `wages_hs_dropout`: The wages data of the high school dropouts.
-  `demographic_nlsy79`: The demographic data of the NLSY79 cohort.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("numbats/yowie")
```

## Example

Here is the example of plotting some samples of subjects in `wages_hs2020` using `brolgar` (Tierney, Cook, and Prvan, 2020).

```{r}
library(yowie)
library(brolgar)
library(ggplot2)

set.seed(20210217)

wages_hs2020 <- as_tsibble(x = wages_hs2020,
                    key = id,
                    index = year,
                    regular = FALSE)

ggplot(wages_hs2020, 
       aes(x = year,
                y = mean_hourly_wage,
                group = id)) +
  geom_line(alpha = 0.7) +
  facet_sample() +
  ylab("mean hourly wage") +
  theme_bw()
```

### 2021-02-20 R/demographic_nlsy79.R
Edit the data documentation.

@@ -1,9 +1,9 @@
-#' Demographic Data from National Longitudinal Survey of Youth
+#' Demographic Data from the National Longitudinal Survey of Youth (NLSY79)
 #'
-#' A data set contains demographic information from The NLSY 1979.
+#' A data set contains demographic information of the NLSY79 cohort.
 #' The cohort included 12,686 respondents ages 14-22 when first interviewed in 1979.
 #'
-#' @source [The NLSY79 Database](https://www.nlsinfo.org/content/cohorts/nlsy79/get-data)
+#' @source The U.S. Bureau of Labor Statistics. (2021, January 6). *National Longitudinal Survey of Youth 1979*. https://www.nlsinfo.org/content/cohorts/nlsy79/get-data
 #'
 #' @format A data frame contains 12,686 rows and 7 variables:
 #' \describe{

### 2021-02-20 R/wages_hs2020.R

@@ -1,18 +1,24 @@
-#' Wages Data from National Longitudinal Survey of Youth
+#' Wages Data from the National Longitudinal Survey of Youth (NLSY79)
 #'
 #' A data set contains longitudinal data of mean hourly wages along
-#' with several demographic variables from National Longitudinal Survey of Youth (NLSY)
+#' with several demographic variables from the National Longitudinal Survey of Youth (NLSY79)
+#' held by the U.S. Bureau of Labor Statistics
 #' from Round 1 (1979 survey year) to Round 28 (2018 survey year).
-#' The cohort provided in this data set are those with maximum highest grade completed is
-#' high school and participated at least 5 years in the survey.
+#' The cohort provided in this data set are those highest grade completed is up to 12th grade
+#' and participated at least 5 rounds of surveys.
 #'
-#' @source [The NLSY79 Database](https://www.nlsinfo.org/content/cohorts/nlsy79/get-data)
+#' @source The U.S. Bureau of Labor Statistics. (2021, January 6). *National Longitudinal Survey of Youth 1979*. https://www.nlsinfo.org/content/cohorts/nlsy79/get-data
 #'
 #' @format A data frame with 103,994 rows and 15 variables:
 #' \describe{
 #'   \item{id}{A unique individual's ID number. This is the `key` of the data.}
 #'   \item{year}{The year the observation was taken. This could be the `index` of the data.}
-#'   \item{mean_hourly_wage}{The mean of the hourly wages the individual gets at each of their different jobs.}
+#'   \item{mean_hourly_wage}{The mean of the hourly wages the individual gets at each of their different jobs.
+#'                           The value could be a weighted or an arithmetic mean.
+#'                           The weighted mean is used when the information of hours of work as the weight is available.
+#'                           The mean hourly wage could also be a predicted value,
+#'                           if the original value is considered influential as the result of robust
+#'                           linear regression as the part of data cleaning. }
 #'   \item{age_1979}{The age of the subject in 1979.}
 #'   \item{gender}{Gender of the subject, FEMALE and MALE.}
 #'   \item{race}{Race of the subject, NON-BLACK,NON-HISPANIC; HISPANIC; BLACK.}
@@ -22,17 +28,10 @@
 #'   \item{number_of_jobs}{Number of jobs that an individual has.}
 #'   \item{total_hours}{The number of hours the individual usually works per week.}
 #'   \item{is_wm}{Whether the mean hourly wage is calculated by weighted
[…]

### 2021-02-20 R/wages_hs_dropout.R

@@ -1,20 +1,26 @@
-#' Wages Data of High School Dropout from The NLSY79
+#' Wages Data of High School Dropout from the National Longitudinal Survey of Youth (NLSY79)
 #'
 #' A data set contains longitudinal data of mean hourly wages along
-#' with several demographic variables from National Longitudinal Survey of Youth (NLSY)
+#' with several demographic variables from the National Longitudinal Survey of Youth (NLSY79)
+#' held by the U.S. Bureau of Labor Statistics
 #' from Round 1 (1979 survey year) to Round 28 (2018 survey year).
 #' The cohort provided in this data set are high school dropout i.e the cohort who only completed
-#' either 9th, 10th, 11th grade, or at least 19 years old when completed high school,
-#' and participated at least 5 years in the survey.
+#' either 9th, 10th, 11th grade, or aged at least 19 years old when completed high school,
+#' and participated at least 5 rounds of surveys.
 #'
-#' @source [The NLSY79 Database](https://www.nlsinfo.org/content/cohorts/nlsy79/get-data)
+#' @source The U.S. Bureau of Labor Statistics. (2021, January 6). *National Longitudinal Survey of Youth 1979*. https://www.nlsinfo.org/content/cohorts/nlsy79/get-data
 #'
 #' @format A data frame contains 97,087 rows and 15 variables:
 #'
 #' \describe{
 #'   \item{id}{A unique individual's ID number. This is the `key` of the data.}
 #'   \item{year}{The year the observation was taken. This could be the `index` of the data.}
-#'   \item{mean_hourly_wage}{The mean of the hourly wages the individual gets at each of their different jobs.}
+#'   \item{mean_hourly_wage}{The mean of the hourly wages the individual gets at each of their different jobs.
+#'                           The value could be a weighted or an arithmetic mean.
+#'                           The weighted mean is used when the information of hours of work as the weight is available.
+#'                           The mean hourly wage could also be a predicted value,
+#'                           if the original value is considered influential as the result of robust
+#'                           linear regression as the part of data cleaning. }
 #'   \item{age_1979}{The age of the subject in 1979.}
 #'   \item{gender}{Gender of the subject, FEMALE and MALE.}
 #'   \item{race}{Race of the subject, NON-BLACK,NON-HISPANIC; HISPANIC; BLACK.}
@@ -24,17 +30,10 @@
 #'   \item{number_of_jobs}{Number of jobs that an individual has.}
[…]

### 2021-02-20 data-raw/data_preprocessing.R

@@ -4,8 +4,9 @@ library(readr)
 library(tidyr)
 library(dplyr)
 library(stringr)
-library(zoo)
-library(rstatix)
+library(MASS)
+library(broom)
+library(purrr)
 
 # READ THE DATA
 
@@ -22,7 +23,7 @@ source(here::here("data-raw/wages-high-school-demo.R"))
 
 ## tidy the date of birth data
 dob_tidy <- new_data_qnames %>%
-  select(CASEID_1979,
+  dplyr::select(CASEID_1979,
          starts_with("Q1-3_A~")) %>%
   mutate(dob_year = case_when(
     # if the years recorded in both sets match, take 79 data
@@ -45,14 +46,14 @@ dob_tidy <- new_data_qnames %>%
       (`Q1-3_A~Y_1979` == `Q1-3_A~Y_1981`) & (`Q1-3_A~M_1979` == `Q1-3_A~M_1981`) ~ FALSE,
       is.na(`Q1-3_A~M_1981`) | is.na(`Q1-3_A~Y_1981`) ~ FALSE))
 dob_tidy <- dob_tidy %>%
-  select(CASEID_1979,
+  dplyr::select(CASEID_1979,
          dob_month,
          dob_year,
          dob_conflict)
 
 ## tidy the sex and race data
 demog_tidy <- categories_qnames %>%
-  select(CASEID_1979,
+  dplyr::select(CASEID_1979,
          SAMPLE_RACE_78SCRN,
          SAMPLE_SEX_1979) %>%
   rename(gender = SAMPLE_SEX_1979,
@@ -62,7 +63,7 @@ demog_tidy <- categories_qnames %>%
 demog_education <- new_data_qnames %>%
   as_tibble() %>%
   rename(HGC_2018 = `Q3-4_2018`) %>%
-  select(CASEID_1979,
+  dplyr::select(CASEID_1979,
          starts_with("HGCREV"),
          "HGC_2012",
          "HGC_2014",
@@ -73,7 +74,7 @@ demog_education <- new_data_qnames %>%
                values_to = "grade") %>%
   separate("var", c("var", "year"), sep = -4) %>%
   filter(!is.na(grade)) %>%
-  select(-var)
+  dplyr::select(-var)
 
 ## getting the highest education completed
 highest_year <- demog_education %>%
@@ -82,7 +83,7 @@ highest_year <- demog_education %>%
   filter(hgc_i == grade) %>%
[…]

### 2021-02-20 README.Rmd
Add hex logo.
Edit the README file.

@@
 
-# yowie
+# yowie <a href='https://numbats.github.io/yowie/'><img src='man/figures/logo.png' align="right" height="138.5" /></a>
 
@@
 
-The goal of yowie is to provide longitudinal wages data sets along with several demographic variables from The National Longitudinal Survey of Youth (NLSY) from survey year 1979 to 2018.
-There are three data sets in this package:
+The goal of `yowie` is to provide longitudinal wages data sets along with several demographic variables of Americans from [the National Longitudinal Survey of Youth (NLSY79)](https://www.nlsinfo.org/content/cohorts/nlsy79) from Round 1 (the survey year 1979) to Round 28 (the survey year 2018).
+The NLSY79 is a longitudinal project held by the U.S.
+Bureau of Labor Statistics that follows the lives of a sample of American youth born between 1957-64.
+The cohort originally included 12,686 respondents ages 14-22 when first interviewed in 1979.
+There are three data sets provided in this package:
 

### 2021-02-20 R/wages_hs2020.R
Grammar and typo checking.

@@ -1,11 +1,11 @@
 #' Wages Data from the National Longitudinal Survey of Youth (NLSY79)
 #'
-#' A data set contains longitudinal data of mean hourly wages along
-#' with several demographic variables from the National Longitudinal Survey of Youth (NLSY79)
-#' held by the U.S. Bureau of Labor Statistics
-#' from Round 1 (1979 survey year) to Round 28 (2018 survey year).
-#' The cohort provided in this data set are those highest grade completed is up to 12th grade
-#' and participated at least 5 rounds of surveys.
+#' A data set contains longitudinal data of mean hourly wages along with several
+#' demographic variables of Americans from the National Longitudinal Survey of Youth
+#' (NLSY79) held by the U.S. Bureau of Labor Statistics from Round 1 (1979 survey year)
+#' to Round 28 (2018 survey year). The cohort provided in this data set is
+#' the highest grade completed, up to 12th grade, and participated in
+#' at least five rounds of surveys.
 #'
 #' @source The U.S. Bureau of Labor Statistics. (2021, January 6). *National Longitudinal Survey of Youth 1979*. https://www.nlsinfo.org/content/cohorts/nlsy79/get-data
 #'
@@ -13,12 +13,13 @@
 #' \describe{
 #'   \item{id}{A unique individual's ID number. This is the `key` of the data.}
 #'   \item{year}{The year the observation was taken. This could be the `index` of the data.}
-#'   \item{mean_hourly_wage}{The mean of the hourly wages the individual gets at each of their different jobs.
-#'                           The value could be a weighted or an arithmetic mean.
-#'                           The weighted mean is used when the information of hours of work as the weight is available.
-#'                           The mean hourly wage could also be a predicted value,
-#'                           if the original value is considered influential as the result of robust
-#'                           linear regression as the part of data cleaning. }
+#'   \item{mean_hourly_wage}{The mean of the hourly wages the individual gets at
+#'                           each of their different jobs. The value could be a
+#'                           weighted or an arithmetic mean. The weighted mean is used
+#'                           when the information of hours of work as the weight
+#'                           is available. The mean hourly wage could also be a predicted
+#'                           value if the original value is considered influential
+#'                           by the robust linear regression as part of data cleaning.}
 #'   \item{age_1979}{The age of the subject in 1979.}
[…]

### 2021-02-20 R/wages_hs_dropout.R

@@ -1,26 +1,27 @@
 #' Wages Data of High School Dropout from the National Longitudinal Survey of Youth (NLSY79)
 #'
-#' A data set contains longitudinal data of mean hourly wages along
-#' with several demographic variables from the National Longitudinal Survey of Youth (NLSY79)
-#' held by the U.S. Bureau of Labor Statistics
-#' from Round 1 (1979 survey year) to Round 28 (2018 survey year).
-#' The cohort provided in this data set are high school dropout i.e the cohort who only completed
-#' either 9th, 10th, 11th grade, or aged at least 19 years old when completed high school,
-#' and participated at least 5 rounds of surveys.
+#' A data set contains longitudinal data of mean hourly wages along with
+#' several demographic variables from the National Longitudinal Survey of Youth (NLSY79)
+#' held by the U.S. Bureau of Labor Statistics from Round 1 (1979 survey year)
+#' to Round 28 (2018 survey year). The cohort provided in this data set is high school
+#' dropouts, i.e., the cohort who only completed either 9th, 10th, 11th grade,
+#' or aged at least 19 years old when completed high school and participated
+#' in at least five rounds of surveys.
 #'
 #' @source The U.S. Bureau of Labor Statistics. (2021, January 6). *National Longitudinal Survey of Youth 1979*. https://www.nlsinfo.org/content/cohorts/nlsy79/get-data
 #'
 #' @format A data frame contains 97,087 rows and 15 variables:
 #'
-#' \describe{
+#' #' \describe{
 #'   \item{id}{A unique individual's ID number. This is the `key` of the data.}
 #'   \item{year}{The year the observation was taken. This could be the `index` of the data.}
-#'   \item{mean_hourly_wage}{The mean of the hourly wages the individual gets at each of their different jobs.
-#'                           The value could be a weighted or an arithmetic mean.
-#'                           The weighted mean is used when the information of hours of work as the weight is available.
-#'                           The mean hourly wage could also be a predicted value,
-#'                           if the original value is considered influential as the result of robust
-#'                           linear regression as the part of data cleaning. }
+#'   \item{mean_hourly_wage}{The mean of the hourly wages the individual gets at
+#'                           each of their different jobs. The value could be a
+#'                           weighted or an arithmetic mean. The weighted mean is used
+#'                           when the information of hours of work as the weight
[…]

### 2021-03-17 R/wages_hs_dropout.R
Edit the vignette and documentation of HS dropouts.

@@ -4,8 +4,8 @@
 #' several demographic variables from the National Longitudinal Survey of Youth (NLSY79)
 #' held by the U.S. Bureau of Labor Statistics from Round 1 (1979 survey year)
 #' to Round 28 (2018 survey year). The cohort provided in this data set is high school
-#' dropouts, i.e., the cohort who only completed either 9th, 10th, 11th grade,
-#' or aged at least 19 years old when completed high school and participated
+#' dropouts, i.e., the male aged between 14 and 17 years and only completed either
+#' 9th, 10th, 11th grade, or aged at least 19 years old when completed high school and participated
 #' in at least five rounds of surveys.
 #'
 #' @source The U.S. Bureau of Labor Statistics. (2021, January 6). *National Longitudinal Survey of Youth 1979*. https://www.nlsinfo.org/content/cohorts/nlsy79/get-data

### 2021-03-17 data-raw/data_preprocessing.R

@@ -322,6 +322,8 @@ wages_hs_dropout <- wages_hs2020 %>%
                      "11TH GRADE")) |
            (hgc == "12TH GRADE" &
               age_hgc >= 19)) %>%
+  filter(age_1979 <= 17,
+         gender == "MALE") %>%
   dplyr::select(-dob,
          -age_hgc)
 

### 2021-08-05 data-raw/data_preprocessing.R
update the desc

@@ -222,6 +222,7 @@ keep_me <- wages_demog_hs %>%
   count(id) %>%
   filter(n > 4)
 
+# keep the id who participated at least 5 years in the survey
 wages_demog_hs <- wages_demog_hs %>%
   filter(id %in% keep_me$id)
 
@@ -263,10 +264,10 @@ id_aug_w <- cbind(id_aug, id_w) %>%
                 w) %>%
   rename(id = `id...1`)
 
-# if the weight < 1, the mean_hourly_wage is replaced by the model's fitted/predicted value.
+# if the weight < 0.12, the mean_hourly_wage is replaced by the model's fitted/predicted value.
 # and add the flag whether the observation is predicted value or not.
-# since the fitted value is sometimes <0, and wages value could never be negative,
-# we keep the mean hourly wage value even its weight < 1.
+# since the fitted value is sometimes < 0, and wages value could never be negative,
+# we keep the mean hourly wage value even its weight < 0.12.
 
 wages_rlm_dat <- id_aug_w %>%
   mutate(wages_rlm = ifelse(w < 0.12  & .fitted >= 0, .fitted,

### 2021-08-21 R/demog_nlsy79.R
getting data to install and document

@@ -16,4 +16,6 @@
 #'  \item{yr_hgc}{The year when the highest grade completed.}
 #' }
 #'
-"demog_nlsy79"
+#' @docType data
+#' @name demog_nlsy79
+NULL

### 2021-08-21 R/wages_hs.R

@@ -38,17 +38,15 @@
 #' # data summary
 #' wages_hs
 #'
-#' # turn the data into a tsibble object
-#' library(brolgar)
-#' wages <- as_tsibble(x = wages_hs2020,
-#'                     key = id,
-#'                     index = year,
-#'                     regular = FALSE)
-#' # create the spaghetti plot of the data
 #' library(ggplot2)
-#' ggplot(wages) +
+#' library(dplyr)
+#' wages_hs_ids <- key_data(wages_hs) %>% select(id)
+#' wages_hs %>%
+#'   dplyr::filter(id %in% sample_n(wages_hs_ids, 10)$id) %>%
+#'   ggplot() +
 #'     geom_line(aes(x = year,
 #'                   y = mean_hourly_wage,
-#'                   group = id), alpha = 0.1)
-#'
-"wages_hs"
+#'                   group = id), alpha = 0.8)
+#' @docType data
+#' @name wages_hs
+NULL

### 2021-08-21 R/wages_hs_do.R

@@ -40,4 +40,6 @@
 #' library(yowie)
 #' library(tsibble)
 #' wages_hs_do %>% key_data()
-"wages_hs_do"
+#' @docType data
+#' @name wages_hs_do
+NULL

### 2021-08-23 README.Rmd
edit the documentation of data pre-processing and example in readme.

@@
 
-wages_hs2020 <- as_tsibble(x = wages_hs2020,
-                    key = id,
-                    index = year,
-                    regular = FALSE)
-
-ggplot(wages_hs2020, 
+ggplot(wages_hs, 
        aes(x = year,
@@
 ```
+
+## About the package name
+
+A [yowie](https://en.wikipedia.org/wiki/Yowie) is a mythical creature in Australian foklore akin to Big Foot or Sasquatch in north America.
+In terms of the package it can be considered an acronym: Years Of Wages to Investigate and Explore.

### 2021-08-23 data-raw/data_preprocessing.R

@@ -7,6 +7,7 @@ library(stringr)
 library(MASS)
 library(broom)
 library(purrr)
+library(tsibble)
 
 # READ THE DATA
 
@@ -294,12 +295,17 @@ wages_hs2020 <- wages_demog_hs %>%
          yr_hgc = as.integer(yr_hgc),
          number_of_jobs = as.integer(number_of_jobs))
 
+wages_hs <- as_tsibble(x = wages_hs2020,
+                       key = id,
+                       index = year,
+                       regular = FALSE)
+
 # save it to an rda object
-usethis::use_data(wages_hs2020, overwrite = TRUE)
+usethis::use_data(wages_hs, overwrite = TRUE)
 
 
 # CREATE THE DATASET FOR THE DOMEGRAPHIC INFORMATION
-demographic_nlsy79 <- full_demographics %>%
+demog_nlsy79 <- full_demographics %>%
   mutate(age_1979 = 1979 - (dob_year + 1900)) %>%
   dplyr::select(id,
          age_1979,
@@ -311,8 +317,11 @@ demographic_nlsy79 <- full_demographics %>%
   mutate(age_1979 = as.integer(age_1979),
          hgc = as.factor(hgc),
          yr_hgc = as.integer(yr_hgc))
+
+
+
 # save it to an rda object
-usethis::use_data(demographic_nlsy79, overwrite = TRUE)
+usethis::use_data(demog_nlsy79, overwrite = TRUE)
 
 # CREATE THE DATASET FOR THE HIGH SCHOOL DROP-OUT
 wages_hs_dropout <- wages_hs2020 %>%
@@ -328,8 +337,13 @@ wages_hs_dropout <- wages_hs2020 %>%
   dplyr::select(-dob,
          -age_hgc)
 
+wages_hs_do <- as_tsibble(x = wages_hs_dropout,
+                          key = id,
+                          index = year,
+                          regular = FALSE)
+
 # save it to an rda object
-usethis::use_data(wages_hs_dropout, overwrite = TRUE)
+usethis::use_data(wages_hs_do, overwrite = TRUE)
 
 
 

### 2021-08-24 R/wages_hs.R
Fix the check error and update documentation.

@@ -9,7 +9,7 @@
 #'
 #' @source The U.S. Bureau of Labor Statistics. (2021, January 6). *National Longitudinal Survey of Youth 1979*. https://www.nlsinfo.org/content/cohorts/nlsy79/get-data
 #'
-#' @format A data frame with 103,994 rows and 15 variables:
+#' @format A tsibble with 103,994 rows and 15 variables:
 #' \describe{
 #'   \item{id}{A unique individual's ID number. This is the `key` of the data.}
 #'   \item{year}{The year the observation was taken. This could be the `index` of the data.}
@@ -40,6 +40,7 @@
 #'
 #' library(ggplot2)
 #' library(dplyr)
+#' library(tsibble)
 #' wages_hs_ids <- key_data(wages_hs) %>% select(id)
 #' wages_hs %>%
 #'   dplyr::filter(id %in% sample_n(wages_hs_ids, 10)$id) %>%
@@ -49,4 +50,5 @@
 #'                   group = id), alpha = 0.8)
 #' @docType data
 #' @name wages_hs
+#' @import tsibble
 NULL

### 2021-08-24 R/wages_hs_do.R

@@ -10,7 +10,7 @@
 #'
 #' @source The U.S. Bureau of Labor Statistics. (2021, January 6). *National Longitudinal Survey of Youth 1979*. https://www.nlsinfo.org/content/cohorts/nlsy79/get-data
 #'
-#' @format A data frame contains 97,087 rows and 15 variables:
+#' @format A tsibble contains 97,087 rows and 15 variables:
 #'
 #' #' \describe{
 #'   \item{id}{A unique individual's ID number. This is the `key` of the data.}
@@ -42,4 +42,5 @@
 #' wages_hs_do %>% key_data()
 #' @docType data
 #' @name wages_hs_do
+#' @import tsibble
 NULL

### 2021-08-25 README.Rmd
Fix the check error and update documentation.
Correcting the df name in Readme.

@@
 
--  `wages_hs2020`: The wages data from the cohort whose highest grade completed is up to 12th grade.
--  `wages_hs_dropout`: The wages data of the high school dropouts.
--  `demographic_nlsy79`: The demographic data of the NLSY79 cohort.
+-  `wages_hs`: The wages data from the cohort whose highest grade completed is up to 12th grade.
+-  `wages_hs_do`: The wages data of the high school dropouts.
+-  `demog_nlsy79`: The demographic data of the NLSY79 cohort.
 
@@
 
-Here is the example of plotting some samples of subjects in `wages_hs2020` using `brolgar` (Tierney, Cook, and Prvan, 2020).
+Here is the example of plotting some samples of subjects in `wages_hs` using `brolgar` (Tierney, Cook, and Prvan, 2020).
 
-```{r}
+```{r, message=FALSE}
 library(yowie)

### 2021-09-01 README.Rmd
Modify variables and data frames.

@@
 
--  `wages_hs`: The wages data from the cohort whose highest grade completed is up to 12th grade.
+-  `wages`: The wages data from the cohort whose highest grade completed is up to 12th grade.
 -  `wages_hs_do`: The wages data of the high school dropouts.
@@
 
-ggplot(wages_hs, 
+ggplot(yowie::wages, 
        aes(x = year,
-                y = mean_hourly_wage,
+           y = wage,
            group = id)) +

### 2021-09-01 R/wages.R

@@ -13,7 +13,7 @@
 #' \describe{
 #'   \item{id}{A unique individual's ID number. This is the `key` of the data.}
 #'   \item{year}{The year the observation was taken. This could be the `index` of the data.}
-#'   \item{mean_hourly_wage}{The mean of the hourly wages the individual gets at
+#'   \item{wage}{The mean of the hourly wages the individual gets at
 #'                           each of their different jobs. The value could be a
 #'                           weighted or an arithmetic mean. The weighted mean is used
 #'                           when the information of hours of work as the weight
@@ -26,8 +26,8 @@
 #'   \item{hgc}{Highest grade completed.}
 #'   \item{hgc_i}{Integer of highest grade completed.}
 #'   \item{yr_hgc}{The year when the highest grade completed.}
-#'   \item{number_of_jobs}{Number of jobs that an individual has.}
-#'   \item{total_hours}{The number of hours the individual usually works per week.}
+#'   \item{njobs}{Number of jobs that an individual has.}
+#'   \item{hours}{The total number of hours the individual usually works per week.}
 #'   \item{is_wm}{Whether the mean hourly wage is weighted mean, using the hour work
 #'                as the weight, or regular/arithmetic mean. TRUE = is weighted mean.
 #'                FALSE = is regular mean.}
@@ -36,19 +36,19 @@
 #'
 #' @examples
 #' # data summary
-#' wages_hs
+#' wages
 #'
 #' library(ggplot2)
 #' library(dplyr)
 #' library(tsibble)
-#' wages_hs_ids <- key_data(wages_hs) %>% select(id)
-#' wages_hs %>%
-#'   dplyr::filter(id %in% sample_n(wages_hs_ids, 10)$id) %>%
+#' wages_ids <- key_data(wages) %>% select(id)
+#' wages %>%
+#'   dplyr::filter(id %in% sample_n(wages_ids, 10)$id) %>%
 #'   ggplot() +
 #'     geom_line(aes(x = year,
-#'                   y = mean_hourly_wage,
+#'                   y = wage,
 #'                   group = id), alpha = 0.8)
 #' @docType data
-#' @name wages_hs
+#' @name wages
 #' @import tsibble
 NULL

### 2021-09-01 R/wages_hs_do.R

@@ -15,7 +15,7 @@
 #' #' \describe{
 #'   \item{id}{A unique individual's ID number. This is the `key` of the data.}
 #'   \item{year}{The year the observation was taken. This could be the `index` of the data.}
-#'   \item{mean_hourly_wage}{The mean of the hourly wages the individual gets at
+#'   \item{wage}{The mean of the hourly wages the individual gets at
 #'                           each of their different jobs. The value could be a
 #'                           weighted or an arithmetic mean. The weighted mean is used
 #'                           when the information of hours of work as the weight
@@ -28,8 +28,8 @@
 #'   \item{hgc}{Highest grade completed.}
 #'   \item{hgc_i}{Integer of highest grade completed.}
 #'   \item{yr_hgc}{The year when the highest grade completed.}
-#'   \item{number_of_jobs}{Number of jobs that an individual has.}
-#'   \item{total_hours}{The number of hours the individual usually works per week.}
+#'   \item{njobs}{Number of jobs that an individual has.}
+#'   \item{hours}{The total number of hours the individual usually works per week.}
 #'   \item{is_wm}{Whether the mean hourly wage is weighted mean, using the hour work
 #'                as the weight, or regular/arithmetic mean. TRUE = is weighted mean.
 #'                FALSE = is regular mean.}

### 2021-09-01 data-raw/data_preprocessing.R

@@ -161,8 +161,10 @@ hours_wages <- left_join(rates_all,
                          hours_all,
                          by = c("id", "year", "job")) %>%
   # set the 0 value in rate_per_hour as NA
+  # set hours_work = 0 or > 84 to be NA
   mutate(rate_per_hour = ifelse(rate_per_hour == 0, NA,
-                                rate_per_hour))
+                                rate_per_hour),
+         hours_work = ifelse(hours_work == 0 | hours_work > 84, NA, hours_work))
 
 ## Calculate the mean hourly wage data
 
@@ -293,15 +295,18 @@ wages_hs2020 <- wages_demog_hs %>%
          year = as.integer(year),
          age_1979 = as.integer(age_1979),
          yr_hgc = as.integer(yr_hgc),
-         number_of_jobs = as.integer(number_of_jobs))
+         number_of_jobs = as.integer(number_of_jobs)) %>%
+  rename(wage = mean_hourly_wage,
+         njobs = number_of_jobs,
+         hours = total_hours)
 
-wages_hs <- as_tsibble(x = wages_hs2020,
+wages <- as_tsibble(x = wages_hs2020,
                        key = id,
                        index = year,
                        regular = FALSE)
 
 # save it to an rda object
-usethis::use_data(wages_hs, overwrite = TRUE)
+usethis::use_data(wages, overwrite = TRUE)
 
 
 # CREATE THE DATASET FOR THE DOMEGRAPHIC INFORMATION

### 2021-09-01 data-raw/data_preprocessing.R
Edit id type from int to factor.

@@ -291,7 +291,8 @@ wages_demog_hs <- wages_demog_hs %>%
 # rename and select the wages in tidy
 wages_hs2020 <- wages_demog_hs %>%
   dplyr::select(id, year, mean_hourly_wage, age_1979, gender, race, hgc, hgc_i, yr_hgc, number_of_jobs, total_hours, is_wm, is_pred) %>%
-  mutate(hgc = as.factor(hgc),
+  mutate(id = as.factor(id),
+         hgc = as.factor(hgc),
          year = as.integer(year),
          age_1979 = as.integer(age_1979),
          yr_hgc = as.integer(yr_hgc),
@@ -319,7 +320,8 @@ demog_nlsy79 <- full_demographics %>%
          hgc,
          hgc_i,
          yr_hgc) %>%
-  mutate(age_1979 = as.integer(age_1979),
+  mutate(id = as.factor(id),
+         age_1979 = as.integer(age_1979),
          hgc = as.factor(hgc),
          yr_hgc = as.integer(yr_hgc))
 

### 2021-09-26 data-raw/data_preprocessing.Rmd
process data vignette (check if it works on github action)

```{r setup}
library(readr)
library(tidyr)
library(dplyr)
library(stringr)
library(MASS)
library(broom)
library(purrr)
library(tsibble)
```

The downloaded data (wages-high-school-demo.dat) from NLSY website
is provided with an R file (wages-high-school-demo.R)
to read the data into R and transform the variable name and
coding the variables' value into something more sensible.
It is also come along with the wages-high-school-demo.NLSY79
Which is a tagset, that can be uploaded to the web site to recreate the data set.

```{r raw-data}
source(here::here("data-raw/wages-high-school-demo.R"))
```

# Demographic data pre-processing
## Tidy the date of birth data

```{r dob-tidy}
dob_tidy <- new_data_qnames %>%
  dplyr::select(CASEID_1979,
         starts_with("Q1-3_A~")) %>%
  mutate(dob_year = case_when(
    # if the years recorded in both sets match, take 79 data
    `Q1-3_A~Y_1979` == `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`,
    # if the year in the 81 set is missing, take the 79 data
    is.na(`Q1-3_A~Y_1981`) ~ `Q1-3_A~Y_1979`,
    # if the sets don't match for dob year, take the 79 data
    `Q1-3_A~Y_1979` != `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`),
    dob_month = case_when(
      # if months recorded in both sets match, take 79 data
      `Q1-3_A~M_1979` == `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`,
      # if month in 81 set is missing, take the 79 data
      is.na(`Q1-3_A~M_1981`) ~ `Q1-3_A~M_1979`,
      # if sets don't match for dob month, take the 79 data
      `Q1-3_A~M_1979` != `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`),
    # flag if there is a conflict between dob recorded in 79 and 81
    dob_conflict = case_when(
      (`Q1-3_A~M_1979` != `Q1-3_A~M_1981`) & !is.na(`Q1-3_A~M_1981`) ~ TRUE,
      (`Q1-3_A~Y_1979` != `Q1-3_A~Y_1981`) & !is.na(`Q1-3_A~Y_1981`) ~ TRUE,
      (`Q1-3_A~Y_1979` == `Q1-3_A~Y_1981`) & (`Q1-3_A~M_1979` == `Q1-3_A~M_1981`) ~ FALSE,
      is.na(`Q1-3_A~M_1981`) | is.na(`Q1-3_A~Y_1981`) ~ FALSE))
dob_tidy <- dob_tidy %>%
  dplyr::select(CASEID_1979,
         dob_month,
         dob_year,
         dob_conflict)
```

Tidy the sex and race data

```{r demog-tidy}
demog_tidy <- categories_qnames %>%
  dplyr::select(CASEID_1979,
         SAMPLE_RACE_78SCRN,
         SAMPLE_SEX_1979) %>%
  rename(gender = SAMPLE_SEX_1979,
         race = SAMPLE_RACE_78SCRN)
```

Tidy the grade completed in each year

```{r demog-ed}
demog_education <- new_data_qnames %>%
  as_tibble() %>%
  rename(HGC_2018 = `Q3-4_2018`) %>%
  dplyr::select(CASEID_1979,
         starts_with("HGCREV"),
         "HGC_2012",
         "HGC_2014",
         "HGC_2016",
         "HGC_2018") %>%
  pivot_longer(!CASEID_1979,
               names_to = "var",
               values_to = "grade") %>%
  separate("var", c("var", "year"), sep = -4) %>%
  filter(!is.na(grade)) %>%
  dplyr::select(-var)

# Getting the highest education completed
highest_year <- demog_education %>%
  group_by(CASEID_1979) %>%
  mutate(hgc_i = max(grade)) %>%
  filter(hgc_i == grade) %>%
  filter(year == first(year)) %>%
  rename(yr_hgc = year) %>%
  dplyr::select(CASEID_1979, yr_hgc, hgc_i) %>%
  ungroup() %>%
  mutate('hgc' = ifelse(hgc_i == 0, "NONE",
                        ifelse(hgc_i == 1, "1ST GRADE",
                               ifelse(hgc_i == 2, "2ND GRADE",
                                      ifelse(hgc_i == 3, "3RD GRADE",
                                             ifelse(hgc_i >= 4 & hgc_i <= 12, paste0(hgc_i,"TH GRADE"),
                                                    ifelse(hgc_i == 13, "1ST YEAR COL",
                                                           ifelse(hgc_i == 14, "2ND YEAR COL",
                                                                  ifelse(hgc_i == 15, "3RD YEAR COL",
                                                                         ifelse(hgc_i == 95, "UNGRADED", paste0((hgc_i - 12), "TH YEAR COL")))))))))))

## Join all of the demographic information
full_demographics <- full_join(dob_tidy, demog_tidy, by = "CASEID_1979") %>%
  full_join(highest_year, by = "CASEID_1979") %>%
  rename("id" = "CASEID_1979")

## EMPLOYMENT DATA (WAGES AND HOURS OF WORK) PREPROCESSING

## Tidying hours of work

# the list of the years using the information from QES-52A
year_A <- c(1979:1987, 1993)
# a function to get the hour of works
get_hour <- function(year){
  if(year %in% year_A){
    temp <- new_data_qnames %>%
      dplyr::select(CASEID_1979,
             starts_with("QES-52A") &
               ends_with(as.character(year)))}
  else{
    temp <- new_data_qnames %>%
      dplyr::select(CASEID_1979,
             starts_with("QES-52D") &
               ends_with(as.character(year)))}
  temp %>%
    pivot_longer(!CASEID_1979,
                 names_to = "job",
                 values_to = "hours_work") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_", substr(job, 9, 10))) %>%
    rename(id = CASEID_1979)
}
# getting the hours of work for all years
hours <- list()
for(ayear in c(1979:1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010,
               2012, 2014, 2016, 2018)) {
  hours[[ayear]] <- get_hour(ayear)
}
hours_all <- bind_rows(!!!hours)

## Tidying rates per hour data
# function to get the rates per hour
get_rate <- function(year) {
  new_data_qnames %>%
    dplyr::select(CASEID_1979,
           starts_with("HRP") &
             ends_with(as.character(year))) %>%
    pivot_longer(!CASEID_1979, names_to = "job", values_to = "rate_per_hour") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_0", substr(job, 4, 4))) %>%
    rename(id = CASEID_1979)
}
# getting the rates per hour for all years
rates <- list()
for(ayear in c(1979:1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010,
               2012, 2014, 2016, 2018)) {
  rates[[ayear]] <- get_rate(ayear)
}
rates_all <- bind_rows(!!!rates)

# join the hours of work and the rates per hour data
hours_wages <- left_join(rates_all,
                         hours_all,
                         by = c("id", "year", "job")) %>%
  # set the 0 value in rate_per_hour as NA
  # set hours_work = 0 or > 84 to be NA
  mutate(rate_per_hour = ifelse(rate_per_hour == 0, NA,
                                rate_per_hour),
         hours_work = ifelse(hours_work == 0 | hours_work > 84, NA, hours_work))

## Calculate the mean hourly wage data

# calculate number of jobs that a person has in one year
no_job <- hours_wages %>%
  filter(!is.na(rate_per_hour)) %>%
  group_by(id, year) %>%
  summarise(no_jobs = length(rate_per_hour))

# filter the observations with available rate per hour
eligible_wages <- hours_wages %>%
  filter(!is.na(rate_per_hour)) %>%
  left_join(no_job, by = c("id", "year"))

# calculate the mean_hourly_wage
mean_hourly_wage <-
  eligible_wages %>%
  group_by(id, year) %>%
  # calculate the weighted mean if the number of jobs > 1
  mutate(wages = ifelse(no_jobs == 1, rate_per_hour/100,
                        weighted.mean(rate_per_hour, hours_work, na.rm = TRUE)/100)) %>%
  # give the flag if it is the weighted mean
  # 1 for weighted mean; 0 for regular/arithmetic mean
  mutate(flag1 = ifelse(!is.na(wages) & no_jobs != 1, 1,
                        0)) %>%
  # calculate the arithmetic mean for the na
  mutate(wages = ifelse(is.na(wages), mean(rate_per_hour)/100,
                        wages)) %>%
  # create the mean hourly wage variable for one id per year i.e. summarise the job
  group_by(id, year) %>%
  summarise(wages = mean(wages),
            total_hours = sum(hours_work),
            number_of_jobs = mean(no_jobs),
            flag1 = mean(flag1)) %>%
  mutate(year = as.numeric(year)) %>%
  ungroup() %>%
  rename(mean_hourly_wage = wages) %>%
  mutate(is_wm = ifelse(flag1 == 1, TRUE,
                        FALSE)) %>%
  dplyr::select(-flag1)

## CREATING THE WAGES_HS2020 DATASET

# join the wages information and the demographic information by case id.
wages_demog <- left_join(mean_hourly_wage, full_demographics, by="id")

# calculate the years in work force and the age of the subjects in 1979
wages_demog <- wages_demog %>%
  mutate(yr_hgc = as.numeric(yr_hgc)) %>%
  mutate(years_in_workforce = year - yr_hgc) %>%
  mutate(age_1979 = 1979 - (dob_year + 1900))

# filter only the id with high school education
wages_demog_hs <- wages_demog  %>% filter(grepl("GRADE", hgc))

# calculate the number of observation
keep_me <- wages_demog_hs %>%
  count(id) %>%
  filter(n > 4)

# keep the id who participated at least 5 years in the survey
wages_demog_hs <- wages_demog_hs %>%
  filter(id %in% keep_me$id)

# nest the data by id to build a robust linear model
by_id <- wages_demog_hs %>%
  dplyr::select(id, year, mean_hourly_wage) %>%
  group_by(id) %>%
  nest()

# build a robust linear model
id_rlm <- by_id %>%
  mutate(model = map(.x = data,
                     .f = function(x){
                       rlm(mean_hourly_wage ~ year, data = x)
                     }))
# extract the property of the regression model
id_aug <- id_rlm %>%
  mutate(augmented = map(model, broom::augment)) %>%
  unnest(augmented)

# extract the weight of each observation
id_w <- id_rlm %>%
  mutate(w = map(.x = model,
                 .f = function(x){
                   x$w
                 })) %>%
  unnest(w) %>%
  dplyr::select(w)

# bind the property of each observation with their weight
id_aug_w <- cbind(id_aug, id_w) %>%
  dplyr::select(`id...1`,
                year,
                mean_hourly_wage,
                .fitted,
                .resid,
                .hat,
                .sigma,
                w) %>%
  rename(id = `id...1`)

# if the weight < 0.12, the mean_hourly_wage is replaced by the model's fitted/predicted value.
# and add the flag whether the observation is predicted value or not.
# since the fitted value is sometimes < 0, and wages value could never be negative,
# we keep the mean hourly wage value even its weight < 0.12.

wages_rlm_dat <- id_aug_w %>%
  mutate(wages_rlm = ifelse(w < 0.12  & .fitted >= 0, .fitted,
                            mean_hourly_wage)) %>%
  mutate(is_pred = ifelse(w  < 0.12 & .fitted >= 0, TRUE,
                          FALSE)) %>%
  dplyr::select(id, year, wages_rlm, is_pred)

# join back the `wages_rlm_dat` to `wages_demog_hs`

wages_demog_hs <- left_join(wages_demog_hs, wages_rlm_dat, by = c("id", "year"))

# select out the old value of mean hourly wage and change it with the wages_rlm value
wages_demog_hs <- wages_demog_hs %>%
  dplyr::select(-mean_hourly_wage) %>%
  rename(mean_hourly_wage = wages_rlm)

# rename and select the wages in tidy
wages_hs2020 <- wages_demog_hs %>%
  dplyr::select(id, year, mean_hourly_wage, age_1979, gender, race, hgc, hgc_i, yr_hgc, number_of_jobs, total_hours, is_wm, is_pred) %>%
  mutate(id = as.factor(id),
         hgc = as.factor(hgc),
         year = as.integer(year),
         age_1979 = as.integer(age_1979),
         yr_hgc = as.integer(yr_hgc),
         number_of_jobs = as.integer(number_of_jobs)) %>%
  rename(wage = mean_hourly_wage,
         njobs = number_of_jobs,
         hours = total_hours)

wages <- as_tsibble(x = wages_hs2020,
                       key = id,
                       index = year,
                       regular = FALSE)

# save it to an rda object
usethis::use_data(wages, overwrite = TRUE)

# CREATE THE DATASET FOR THE DOMEGRAPHIC INFORMATION
demog_nlsy79 <- full_demographics %>%
  mutate(age_1979 = 1979 - (dob_year + 1900)) %>%
  dplyr::select(id,
         age_1979,
         gender,
         race,
         hgc,
         hgc_i,
         yr_hgc) %>%
  mutate(id = as.factor(id),
         age_1979 = as.integer(age_1979),
         hgc = as.factor(hgc),
         yr_hgc = as.integer(yr_hgc))

```
```{r eval=FALSE}
# save it to an rda object
usethis::use_data(demog_nlsy79, overwrite = TRUE)

# CREATE THE DATASET FOR THE HIGH SCHOOL DROP-OUT
wages_hs_dropout <- wages_hs2020 %>%
  mutate(dob = 1979 - age_1979,
         age_hgc = yr_hgc - dob) %>%
  filter((hgc %in% c("9TH GRADE",
                     "10TH GRADE",
                     "11TH GRADE")) |
           (hgc == "12TH GRADE" &
              age_hgc >= 19)) %>%
  filter(age_1979 <= 17,
         gender == "MALE") %>%
  dplyr::select(-dob,
         -age_hgc)

wages_hs_do <- as_tsibble(x = wages_hs_dropout,
                          key = id,
                          index = year,
                          regular = FALSE)

```
```{r eval=FALSE}
# save it to an rda object
usethis::use_data(wages_hs_do, overwrite = TRUE)
```

### 2021-09-28 data-raw/data-preprocessing.Rmd
change pkgdown from github actions to local, added the processing of raw to input data

The data cleaning step is conducted using the suite of packages in `tidyverse`.

```{r load-pkgs}
library(tidyverse)
select <- dplyr::select
```

# Reading the data

```{r raw-data, eval=FALSE, include=FALSE}
source(here::here("data-raw/NLSY79/NLSY79.R"))
```

<details class="source">
<summary>Click here for the **source file to read data** (note this is quite long)</summary>
The code below is provided by the NLSY79 database to do the
reading and initial cleaning of data.

```{r raw-data-rmd, code=readLines(here::here("data-raw/NLSY79/NLSY79.R"))}
```

</details>
The above source code creates a data set `new_data_qnames`.
As shown below, the column names contain information on the job number
(HRP1 = job 1, HRP2 = job 2, ..., HRP5 = job 5) and the year information.

```{r untidy-data, echo=-1}
options(width=70)
new_data_qnames %>%
  select(CASEID_1979,
         starts_with("HRP") &
           ends_with(c("1979", "1980")),
         everything()) %>%
  str(list.len = 10)
```

# Demographic data pre-processing
## Tidy the date of birth data

```{r dob-tidy}
dob_tidy <- new_data_qnames %>%
  select(CASEID_1979,
                starts_with("Q1-3_A~")) %>%
  mutate(dob_year = case_when(
                      # if the years recorded in both sets match, take 79 data
                      `Q1-3_A~Y_1979` == `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`,
                      # if the year in the 81 set is missing, take the 79 data
                      is.na(`Q1-3_A~Y_1981`) ~ `Q1-3_A~Y_1979`,
                      # if the sets don't match for dob year, take the 79 data
                      `Q1-3_A~Y_1979` != `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`),
         dob_month = case_when(
                      # if months recorded in both sets match, take 79 data
                      `Q1-3_A~M_1979` == `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`,
                      # if month in 81 set is missing, take the 79 data
                      is.na(`Q1-3_A~M_1981`) ~ `Q1-3_A~M_1979`,
                      # if sets don't match for dob month, take the 79 data
                      `Q1-3_A~M_1979` != `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`),
         # flag if there is a conflict between dob recorded in 79 and 81
         dob_conflict = case_when(
                      (`Q1-3_A~M_1979` != `Q1-3_A~M_1981`) & !is.na(`Q1-3_A~M_1981`)
                      ~ TRUE,
                      (`Q1-3_A~Y_1979` != `Q1-3_A~Y_1981`) & !is.na(`Q1-3_A~Y_1981`)
                      ~ TRUE,
                      (`Q1-3_A~Y_1979` == `Q1-3_A~Y_1981`) &
                        (`Q1-3_A~M_1979` == `Q1-3_A~M_1981`) ~ FALSE,
                      is.na(`Q1-3_A~M_1981`) | is.na(`Q1-3_A~Y_1981`) ~ FALSE)) %>%
  select(CASEID_1979,
         dob_month,
         dob_year,
         dob_conflict)
```

## Tidy the sex and race data

```{r demog-tidy}
demog_tidy <- categories_qnames %>%
  select(CASEID_1979,
         SAMPLE_RACE_78SCRN,
         SAMPLE_SEX_1979) %>%
  rename(gender = SAMPLE_SEX_1979,
         race = SAMPLE_RACE_78SCRN)
```

## Tidy the grade completed in each year

```{r demog-ed}
demog_education <- new_data_qnames %>%
  as_tibble() %>%
  # in 2018, the variable's name is Q3-4_2018, instead of HGC_2018
  rename(HGC_2018 = `Q3-4_2018`) %>%
  select(CASEID_1979,
                starts_with("HGCREV"),
                "HGC_2012",
                "HGC_2014",
                "HGC_2016",
                "HGC_2018") %>%
  pivot_longer(!CASEID_1979,
               names_to = "var",
               values_to = "grade") %>%
  separate("var", c("var", "year"), sep = -4) %>%
  filter(!is.na(grade)) %>%
  select(-var)
```

### Get the highest year completed

```{r tidy-hgc}
highest_year <- demog_education %>%
  group_by(CASEID_1979) %>%
  mutate(hgc_i = max(grade)) %>%
  filter(hgc_i == grade) %>%
  filter(year == first(year)) %>%
  rename(yr_hgc = year) %>%
  select(CASEID_1979, yr_hgc, hgc_i) %>%
  ungroup() %>%
  mutate(hgc = case_when(hgc_i == 0 ~ "NONE",
                         hgc_i == 1 ~ "1ST GRADE",
                         hgc_i == 2 ~ "2ND GRADE",
                         hgc_i == 3 ~ "3RD GRADE",
                         hgc_i >= 4 & hgc_i <= 12 ~ paste0(hgc_i,"TH GRADE"),
                         hgc_i == 13 ~ "1ST YEAR COL",
                         hgc_i == 14 ~ "2ND YEAR COL",
                         hgc_i == 15 ~ "3RD YEAR COL",
                         hgc_i == 95 ~ "UNGRADED",
                         TRUE ~ paste0((hgc_i - 12), "TH YEAR COL")))

```
```{r full-demog}
full_demographics <- full_join(dob_tidy, demog_tidy, by = "CASEID_1979") %>%
  full_join(highest_year, by = "CASEID_1979") %>%
  rename("id" = "CASEID_1979") %>%
  mutate(age_1979 = 1979 - (dob_year + 1900))

```
```{r tidy-hours}
# make a list for years where we used the "QES-52A"
year_A <- c(1979:1987, 1993)
# function to get the hour of work
get_hour <- function(year) {
  if(year %in% year_A){
    temp <- new_data_qnames %>%
      select(CASEID_1979,
                    starts_with("QES-52A") &
                      ends_with(as.character(year)))}
  else{
    temp <- new_data_qnames %>%
      select(CASEID_1979,
                    starts_with("QES-52D") &
                      ends_with(as.character(year)))}
  temp %>%
    pivot_longer(!CASEID_1979,
                 names_to = "job",
                 values_to = "hours_work") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_", substr(job, 9, 10))) %>%
    rename(id = CASEID_1979)
}

# getting the hours of work of all observations
years <- c(1979:1994, seq(1996, 2018, by = 2))
hours_all <- map_dfr(years, get_hour)

```
```{r tidy-rate}
get_rate <- function(year) {
  new_data_qnames %>%
    select(CASEID_1979,
                  starts_with("HRP") &
                    ends_with(as.character(year))) %>%
    pivot_longer(!CASEID_1979, names_to = "job", values_to = "rate_per_hour") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_0", substr(job, 4, 4))) %>%
    rename(id = CASEID_1979)
}

rates_all <- map_dfr(years, get_rate)

# join hours and rates variable
hours_wages <- left_join(rates_all,
                         hours_all,
                         by = c("id", "year", "job")) %>%
  # set the 0 value in rate_per_hour as NA
  mutate(rate_per_hour = ifelse(rate_per_hour == 0, NA, rate_per_hour),
         hours_work = ifelse(hours_work == 0 | hours_work > 84, NA, hours_work))

```
```{r tidy-nojob}
# calculate number of jobs that a person has in one year
no_job <- hours_wages %>%
  filter(!is.na(rate_per_hour)) %>%
  group_by(id, year) %>%
  summarise(no_jobs = length(rate_per_hour))

# filter the observations with available rate per hour
eligible_wages <- hours_wages %>%
  filter(!is.na(rate_per_hour)) %>%
  left_join(no_job, by = c("id", "year"))

# calculate the mean_hourly_wage
# flag1 = code 1 for weighted mean
# code 0 for arithmetic mean
mean_hourly_wage <- eligible_wages %>%
  group_by(id, year) %>%
  #calculate the weighted mean if the number of jobs > 1
  mutate(wages = ifelse(no_jobs == 1, rate_per_hour/100,
                        weighted.mean(rate_per_hour, hours_work, na.rm = TRUE)/100)) %>%
  #give the flag if it the weighted mean
  mutate(flag1 = ifelse(!is.na(wages) & no_jobs != 1, 1,
                        0)) %>%
  #calculate the arithmetic mean for the na
  mutate(wages = ifelse(is.na(wages), mean(rate_per_hour)/100,
                        wages)) %>%
  group_by(id, year) %>%
  summarise(wages = mean(wages),
            total_hours = sum(hours_work),
            number_of_jobs = mean(no_jobs),
            flag1 = mean(flag1)) %>%
  mutate(year = as.numeric(year)) %>%
  ungroup() %>%
  rename(mean_hourly_wage = wages) %>%
  mutate(is_wm = ifelse(flag1 == 1, TRUE,
                        FALSE)) %>%
  select(-flag1)
head(mean_hourly_wage, n = 10)

```
```{r wages-demog-hs}
# join the wages information and the demographic information by case id.
wages_demog <- left_join(mean_hourly_wage, full_demographics, by="id")
# calculate the years in work force and the age of the subjects in 1979
wages_demog <- wages_demog %>%
  mutate(yr_hgc = as.numeric(yr_hgc)) %>%
  mutate(years_in_workforce = year - yr_hgc) %>%
  mutate(age_1979 = 1979 - (dob_year + 1900))
# filter only the id with high school education
wages_before <- wages_demog  %>% filter(grepl("GRADE", hgc))
# calculate the number of observation
keep_me <- wages_before %>%
  count(id) %>%
  filter(n > 4)
wages_before <- wages_before %>%
  filter(id %in% keep_me$id)
```

### 2021-09-28 data-raw/ORIGINAL-data_preprocessing.R

@@ -1,0 +1,362 @@
+
+
+## ---- setup
+library(readr)
+library(tidyr)
+library(dplyr)
+library(stringr)
+library(MASS)
+library(broom)
+library(purrr)
+library(tsibble)
+
+
+#' The downloaded data (wages-high-school-demo.dat) from NLSY website
+#' is provided with an R file (wages-high-school-demo.R)
+#' to read the data into R and transform the variable name and
+#' coding the variables' value into something more sensible.
+#' It is also come along with the wages-high-school-demo.NLSY79
+#' Which is a tagset, that can be uploaded to the web site to recreate the data set.
+## ---- raw-data
+source(here::here("data-raw/wages-high-school-demo.R"))
+
+#' # Demographic data pre-processing
+#' ## Tidy the date of birth data
+
+## ---- dob-tidy
+dob_tidy <- new_data_qnames %>%
+  dplyr::select(CASEID_1979,
+         starts_with("Q1-3_A~")) %>%
+  mutate(dob_year = case_when(
+    # if the years recorded in both sets match, take 79 data
+    `Q1-3_A~Y_1979` == `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`,
+    # if the year in the 81 set is missing, take the 79 data
+    is.na(`Q1-3_A~Y_1981`) ~ `Q1-3_A~Y_1979`,
+    # if the sets don't match for dob year, take the 79 data
+    `Q1-3_A~Y_1979` != `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`),
+    dob_month = case_when(
+      # if months recorded in both sets match, take 79 data
+      `Q1-3_A~M_1979` == `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`,
+      # if month in 81 set is missing, take the 79 data
+      is.na(`Q1-3_A~M_1981`) ~ `Q1-3_A~M_1979`,
+      # if sets don't match for dob month, take the 79 data
+      `Q1-3_A~M_1979` != `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`),
+    # flag if there is a conflict between dob recorded in 79 and 81
+    dob_conflict = case_when(
[…]

### 2021-09-28 data-raw/data-preprocessing.R

@@ -1,0 +1,225 @@
+
+
+#' The data cleaning step is conducted using the suite of packages in `tidyverse`.
+#+ load-pkgs
+library(tidyverse)
+select <- dplyr::select
+
+#' # Reading the data
+#+ raw-data, eval=FALSE, include=FALSE
+source(here::here("data-raw/NLSY79/NLSY79.R"))
+
+#' <details class="source">
+#' <summary>Click here for the **source file to read data** (note this is quite long)</summary>
+#' The code below is provided by the NLSY79 database to do the
+#' reading and initial cleaning of data.
+#+ raw-data-rmd, code=readLines(here::here("data-raw/NLSY79/NLSY79.R"))
+#' </details>
+
+#' The above source code creates a data set `new_data_qnames`.
+#' As shown below, the column names contain information on the job number
+#' (HRP1 = job 1, HRP2 = job 2, ..., HRP5 = job 5) and the year information.
+#+ untidy-data, echo=-1
+options(width=70)
+new_data_qnames %>%
+  select(CASEID_1979,
+         starts_with("HRP") &
+           ends_with(c("1979", "1980")),
+         everything()) %>%
+  str(list.len = 10)
+
+#' # Demographic data pre-processing
+#' ## Tidy the date of birth data
+#+ dob-tidy
+dob_tidy <- new_data_qnames %>%
+  select(CASEID_1979,
+                starts_with("Q1-3_A~")) %>%
+  mutate(dob_year = case_when(
+                      # if the years recorded in both sets match, take 79 data
+                      `Q1-3_A~Y_1979` == `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`,
+                      # if the year in the 81 set is missing, take the 79 data
+                      is.na(`Q1-3_A~Y_1981`) ~ `Q1-3_A~Y_1979`,
+                      # if the sets don't match for dob year, take the 79 data
+                      `Q1-3_A~Y_1979` != `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`),
+         dob_month = case_when(
+                      # if months recorded in both sets match, take 79 data
+                      `Q1-3_A~M_1979` == `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`,
+                      # if month in 81 set is missing, take the 79 data
[…]

### 2021-09-28 data-raw/data_preprocessing.R

@@ -1,362 +1,0 @@
-
-
-## ---- setup
-library(readr)
-library(tidyr)
-library(dplyr)
-library(stringr)
-library(MASS)
-library(broom)
-library(purrr)
-library(tsibble)
-
-
-#' The downloaded data (wages-high-school-demo.dat) from NLSY website
-#' is provided with an R file (wages-high-school-demo.R)
-#' to read the data into R and transform the variable name and
-#' coding the variables' value into something more sensible.
-#' It is also come along with the wages-high-school-demo.NLSY79
-#' Which is a tagset, that can be uploaded to the web site to recreate the data set.
-## ---- raw-data
-source(here::here("data-raw/wages-high-school-demo.R"))
-
-#' # Demographic data pre-processing
-#' ## Tidy the date of birth data
-
-## ---- dob-tidy
-dob_tidy <- new_data_qnames %>%
-  dplyr::select(CASEID_1979,
-         starts_with("Q1-3_A~")) %>%
-  mutate(dob_year = case_when(
-    # if the years recorded in both sets match, take 79 data
-    `Q1-3_A~Y_1979` == `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`,
-    # if the year in the 81 set is missing, take the 79 data
-    is.na(`Q1-3_A~Y_1981`) ~ `Q1-3_A~Y_1979`,
-    # if the sets don't match for dob year, take the 79 data
-    `Q1-3_A~Y_1979` != `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`),
-    dob_month = case_when(
-      # if months recorded in both sets match, take 79 data
-      `Q1-3_A~M_1979` == `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`,
-      # if month in 81 set is missing, take the 79 data
-      is.na(`Q1-3_A~M_1981`) ~ `Q1-3_A~M_1979`,
-      # if sets don't match for dob month, take the 79 data
-      `Q1-3_A~M_1979` != `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`),
-    # flag if there is a conflict between dob recorded in 79 and 81
-    dob_conflict = case_when(
[…]

### 2021-09-28 data-raw/data-preprocessing.Rmd
remove docs from gitignore, add docs

@@
 
-# Reading the data
+# Reading the data {#read-data}
 
@@
 </details>
+
 The above source code creates a data set `new_data_qnames`.
@@
 ## Tidy the date of birth data
+The month and year of birth is recorded in 1979 and 1981 for each individual.
+The records in 1981 are missing for some individuals so we take the month and year of
+birth from 1979 records.
+
+Where the record is present for both 1979 and 1981, we check that the
+record matches.
 
@@
 dob_tidy <- new_data_qnames %>%
-  select(CASEID_1979,
-                starts_with("Q1-3_A~")) %>%
+  select(id = CASEID_1979,
+         year_1979 = `Q1-3_A~Y_1979`,
+         year_1981 = `Q1-3_A~Y_1981`,
+         month_1979 = `Q1-3_A~M_1979`,
+         month_1981 = `Q1-3_A~M_1981`) %>%
   mutate(dob_year = case_when(
                       # if the years recorded in both sets match, take 79 data
-                      `Q1-3_A~Y_1979` == `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`,
+                      year_1979 == year_1981 ~ year_1979,
                       # if the year in the 81 set is missing, take the 79 data
-                      is.na(`Q1-3_A~Y_1981`) ~ `Q1-3_A~Y_1979`,
+                      is.na(year_1981) ~ year_1979,
                       # if the sets don't match for dob year, take the 79 data
-                      `Q1-3_A~Y_1979` != `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`),
+                      year_1979 != year_1981 ~ year_1979),
          dob_month = case_when(
                       # if months recorded in both sets match, take 79 data
-                      `Q1-3_A~M_1979` == `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`,
+                      month_1979 == month_1981 ~ month_1979,
                       # if month in 81 set is missing, take the 79 data
-                      is.na(`Q1-3_A~M_1981`) ~ `Q1-3_A~M_1979`,
+                      is.na(month_1981) ~ month_1979,
                       # if sets don't match for dob month, take the 79 data
-                      `Q1-3_A~M_1979` != `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`),
+                      month_1979 != month_1981 ~ month_1979),
          # flag if there is a conflict between dob recorded in 79 and 81
          dob_conflict = case_when(
-                      (`Q1-3_A~M_1979` != `Q1-3_A~M_1981`) & !is.na(`Q1-3_A~M_1981`)
-                      ~ TRUE,
-                      (`Q1-3_A~Y_1979` != `Q1-3_A~Y_1981`) & !is.na(`Q1-3_A~Y_1981`)
-                      ~ TRUE,
-                      (`Q1-3_A~Y_1979` == `Q1-3_A~Y_1981`) &
-                        (`Q1-3_A~M_1979` == `Q1-3_A~M_1981`) ~ FALSE,
-                      is.na(`Q1-3_A~M_1981`) | is.na(`Q1-3_A~Y_1981`) ~ FALSE)) %>%
-  select(CASEID_1979,
+                      !is.na(month_1981) && (month_1979 != month_1981) ~ TRUE,
+                      !is.na(year_1981) && (year_1979 != year_1981) ~ TRUE,
+                      (year_1979 == year_1981) & (month_1979 == month_1981) ~ FALSE,
+                      is.na(month_1981) | is.na(year_1981) ~ FALSE)) %>%
+  select(id,
          dob_month,
@@
          dob_conflict)
+
+has_dob_conflict <- any(dob_tidy$dob_conflict, na.rm = TRUE)
+```
+```{r dob-conflict-no, echo=FALSE, eval=!has_dob_conflict}
+cat("All birth month and year recorded in 1979 and 1981 match.")
+```
+```{r dob-conflict-yes, echo=has_dob_conflict, eval=has_dob_conflict}
+cat("The birth record does not match for the following individuals.")
+dob_tidy %>%
+  filter(dob_conflict)
 ```
@@
 demog_tidy <- categories_qnames %>%
-  select(CASEID_1979,
-         SAMPLE_RACE_78SCRN,
-         SAMPLE_SEX_1979) %>%
-  rename(gender = SAMPLE_SEX_1979,
-         race = SAMPLE_RACE_78SCRN)
+  select(id = CASEID_1979,
+         race = SAMPLE_RACE_78SCRN,
+         gender = SAMPLE_SEX_1979)
 ```
@@
 demog_education <- new_data_qnames %>%
-  as_tibble() %>%
   # in 2018, the variable's name is Q3-4_2018, instead of HGC_2018
   rename(HGC_2018 = `Q3-4_2018`) %>%
-  select(CASEID_1979,
+  select(id = CASEID_1979,
          starts_with("HGCREV"),
@@
          "HGC_2018") %>%
-  pivot_longer(!CASEID_1979,
+  pivot_longer(!id,
                names_to = "var",
                values_to = "grade") %>%
-  separate("var", c("var", "year"), sep = -4) %>%
+  separate("var", c("var", "year"), sep = "_") %>%
   filter(!is.na(grade)) %>%
@@
 highest_year <- demog_education %>%
-  group_by(CASEID_1979) %>%
+  group_by(id) %>%
   mutate(hgc_i = max(grade)) %>%
@@
   rename(yr_hgc = year) %>%
-  select(CASEID_1979, yr_hgc, hgc_i) %>%
+  select(id, yr_hgc, hgc_i) %>%
   ungroup() %>%
@@
                          hgc_i == 3 ~ "3RD GRADE",
-                         hgc_i >= 4 & hgc_i <= 12 ~ paste0(hgc_i,"TH GRADE"),
+                         hgc_i >= 4 & hgc_i <= 12 ~ paste0(hgc_i, "TH GRADE"),
                          hgc_i == 13 ~ "1ST YEAR COL",
@@
 ```{r full-demog}
-full_demographics <- full_join(dob_tidy, demog_tidy, by = "CASEID_1979") %>%
-  full_join(highest_year, by = "CASEID_1979") %>%
-  rename("id" = "CASEID_1979") %>%
+full_demographics <- full_join(dob_tidy, demog_tidy, by = "id") %>%
+  full_join(highest_year, by = "id") %>%
   mutate(age_1979 = 1979 - (dob_year + 1900))

### 2021-09-28 data-raw/data-preprocessing.R

@@ -5,7 +5,7 @@
 library(tidyverse)
 select <- dplyr::select
 
-#' # Reading the data
+#' # Reading the data {#read-data}
 #+ raw-data, eval=FALSE, include=FALSE
 source(here::here("data-raw/NLSY79/NLSY79.R"))
 
@@ -16,6 +16,7 @@ source(here::here("data-raw/NLSY79/NLSY79.R"))
 #+ raw-data-rmd, code=readLines(here::here("data-raw/NLSY79/NLSY79.R"))
 #' </details>
 
+#'
 #' The above source code creates a data set `new_data_qnames`.
 #' As shown below, the column names contain information on the job number
 #' (HRP1 = job 1, HRP2 = job 2, ..., HRP5 = job 5) and the year information.
@@ -30,81 +31,93 @@ new_data_qnames %>%
 
 #' # Demographic data pre-processing
 #' ## Tidy the date of birth data
+#' The month and year of birth is recorded in 1979 and 1981 for each individual.
+#' The records in 1981 are missing for some individuals so we take the month and year of
+#' birth from 1979 records.
+#'
+#' Where the record is present for both 1979 and 1981, we check that the
+#' record matches.
 #+ dob-tidy
 dob_tidy <- new_data_qnames %>%
-  select(CASEID_1979,
-                starts_with("Q1-3_A~")) %>%
+  select(id = CASEID_1979,
+         year_1979 = `Q1-3_A~Y_1979`,
+         year_1981 = `Q1-3_A~Y_1981`,
+         month_1979 = `Q1-3_A~M_1979`,
+         month_1981 = `Q1-3_A~M_1981`) %>%
   mutate(dob_year = case_when(
                       # if the years recorded in both sets match, take 79 data
-                      `Q1-3_A~Y_1979` == `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`,
+                      year_1979 == year_1981 ~ year_1979,
                       # if the year in the 81 set is missing, take the 79 data
-                      is.na(`Q1-3_A~Y_1981`) ~ `Q1-3_A~Y_1979`,
+                      is.na(year_1981) ~ year_1979,
                       # if the sets don't match for dob year, take the 79 data
-                      `Q1-3_A~Y_1979` != `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`),
+                      year_1979 != year_1981 ~ year_1979),
[…]

### 2021-09-28 data-raw/data-preprocessing.Rmd
vignette:  input to valid

@@
   filter(id %in% keep_me$id)
+
 ```
+```{r save-data, echo = FALSE, eval = FALSE}
+saveRDS(wages_before, here::here("paper/results/wages_before.rds"))
+```

### 2021-09-28 data-raw/data-valid.Rmd
vignette:  input to valid

```{r }
library(tidyverse)
library(MASS)
select <- dplyr::select
wages_before <- readRDS(here::here("paper/results/wages_before.rds"))

```
```{r rlm}

# nest the data by id to build a robust linear model
by_id <- wages_before %>%
  select(id, year, mean_hourly_wage) %>%
  group_by(id) %>%
  nest()

# build a robust linear model
id_rlm <- by_id %>%
  mutate(model = map(.x = data,
                     .f = function(x){
                       rlm(mean_hourly_wage ~ year, data = x)
                     }))
# extract the property of the regression model
id_aug <- id_rlm %>%
  mutate(augmented = map(model, broom::augment)) %>%
  unnest(augmented)

# extract the weight of each observation
id_w <- id_rlm %>%
  mutate(w = map(.x = model,
                 .f = function(x){
                   x$w
                 })) %>%
  unnest(w) %>%
  select(w)

# bind the property of each observation with their weight
id_aug_w <- cbind(id_aug, id_w) %>%
  select(`id...1`,
         year,
         mean_hourly_wage,
         .fitted,
         .resid,
         .hat,
         .sigma,
         w) %>%
  rename(id = `id...1`)

# if the weight < 0.12, the mean_hourly_wage is replaced by the model's fitted/predicted value.
# and add the flag whether the observation is predicted value or not.
# since the fitted value is sometimes <0, and wages value could never be negative,
# we keep the mean hourly wage value even its weight < 0.12.

wages_rlm_dat <- id_aug_w %>%
  mutate(wages_rlm = ifelse(w < 0.12  & .fitted >= 0, .fitted,
                            mean_hourly_wage)) %>%
  mutate(is_pred = ifelse(w < 0.12 & .fitted >= 0, TRUE,
                          FALSE)) %>%
  select(id, year, wages_rlm, is_pred)

# join back the `wages_rlm_dat` to `wages_demog_hs`

wages_after <- left_join(wages_before, wages_rlm_dat, by = c("id", "year"))

# save it to rds file so it would faster the knitting process
```
```{r save-wages-after, eval = FALSE, echo = FALSE}
saveRDS(wages_after, here::here("paper/results/wages_after.rds"))
```

### 2021-09-28 data-raw/data-preprocessing.R

@@ -235,3 +235,6 @@ keep_me <- wages_before %>%
   filter(n > 4)
 wages_before <- wages_before %>%
   filter(id %in% keep_me$id)
+
+#+ save-data, echo = FALSE, eval = FALSE
+saveRDS(wages_before, here::here("paper/results/wages_before.rds"))

### 2021-09-28 data-raw/data-valid.R

@@ -1,0 +1,65 @@
+
+library(tidyverse)
+library(MASS)
+select <- dplyr::select
+wages_before <- readRDS(here::here("paper/results/wages_before.rds"))
+
+## ---- rlm
+
+# nest the data by id to build a robust linear model
+by_id <- wages_before %>%
+  select(id, year, mean_hourly_wage) %>%
+  group_by(id) %>%
+  nest()
+
+# build a robust linear model
+id_rlm <- by_id %>%
+  mutate(model = map(.x = data,
+                     .f = function(x){
+                       rlm(mean_hourly_wage ~ year, data = x)
+                     }))
+# extract the property of the regression model
+id_aug <- id_rlm %>%
+  mutate(augmented = map(model, broom::augment)) %>%
+  unnest(augmented)
+
+# extract the weight of each observation
+id_w <- id_rlm %>%
+  mutate(w = map(.x = model,
+                 .f = function(x){
+                   x$w
+                 })) %>%
+  unnest(w) %>%
+  select(w)
+
+# bind the property of each observation with their weight
+id_aug_w <- cbind(id_aug, id_w) %>%
+  select(`id...1`,
+         year,
+         mean_hourly_wage,
+         .fitted,
+         .resid,
+         .hat,
+         .sigma,
+         w) %>%
+  rename(id = `id...1`)
+
+# if the weight < 0.12, the mean_hourly_wage is replaced by the model's fitted/predicted value.
+# and add the flag whether the observation is predicted value or not.
+# since the fitted value is sometimes <0, and wages value could never be negative,
+# we keep the mean hourly wage value even its weight < 0.12.
+
+wages_rlm_dat <- id_aug_w %>%
+  mutate(wages_rlm = ifelse(w < 0.12  & .fitted >= 0, .fitted,
+                            mean_hourly_wage)) %>%
+  mutate(is_pred = ifelse(w < 0.12 & .fitted >= 0, TRUE,
+                          FALSE)) %>%
+  select(id, year, wages_rlm, is_pred)
+
+# join back the `wages_rlm_dat` to `wages_demog_hs`
+
+wages_after <- left_join(wages_before, wages_rlm_dat, by = c("id", "year"))
+
+# save it to rds file so it would faster the knitting process
+#+ save-wages-after, eval = FALSE, echo = FALSE
+saveRDS(wages_after, here::here("paper/results/wages_after.rds"))

### 2021-09-28 data-raw/eda.R

@@ -1,0 +1,198 @@
+## ---- sw_wages
+
+sw <- brolgar::wages
+
+sw_wages <- sw %>%
+  ggplot(aes(x = xp,
+             y = ln_wages)) +
+  geom_line(aes(group = id), alpha = 0.1) +
+  geom_smooth(se = FALSE, colour = "pink") +
+  ggtitle("A original") +
+  theme_bw() +
+  ylab("Hourly wage ($, natural log)") +
+  xlab("Experience in years") +
+  #theme(plot.title = element_text(size = 10)) +
+  ylim(0, 5)
+
+sw_wages_mod <- sw %>%
+  as_tibble() %>%
+  mutate(hgc = ifelse(high_grade < 9, "8TH", "12TH")) %>%
+  mutate(race = case_when(black == 1 ~ "black",
+                          hispanic == 1 ~ "hispanic",
+                          TRUE ~ "white")) %>%
+  mutate(race = factor(race)) %>%
+  ggplot(aes(x = xp,
+             y = ln_wages,
+             linetype = hgc)) +
+  geom_smooth(method = "lm", se = FALSE,
+              colour = "black") +
+  ggtitle("A original") +
+  scale_linetype("") +
+  theme_bw() +
+  ylab("Hourly wage ($, natural log)") +
+  xlab("Experience in years") +
+  theme(legend.position = "bottom",
+        legend.direction = "horizontal")
+
+## ---- do_refreshed
+
+# singer-willet did not stated how they calculate experience
+# in database, there is work experience topic
+# (https://www.nlsinfo.org/content/cohorts/nlsy79/topical-guide/employment/work-experience)
+# but there is no variable explicitly mention about experience.
+# in this case, we approximate it with year-1979
+
+# filter do data up to 1994 since the textbook data only covers that period.
+
+do <- yowie::wages_hs_do %>%
+  mutate(lnwage = log(wage)) #%>%
+  #filter(year < 1995)
+
+do_ref <- do %>%
+  ggplot(aes(x = year,
+             y = lnwage)) +
+  geom_line(aes(group = id), alpha = 0.1) +
+  geom_smooth(se = FALSE, colour = "pink") +
+  ggtitle("B refreshed data") +
+  theme_bw() +
+  ylab("Hourly wage ($, natural log)") +
+  xlab("Year of data collection") +
+  theme(plot.title = element_text(size = 10)) +
+  ylim(0, 5)
+
+do_ref_mod <- do %>%
[…]

### 2021-09-28 data-raw/ida.R

@@ -1,0 +1,363 @@
+## ---- age-table
+age_table <- full_demographics %>%
+  group_by(age_1979) %>%
+  count(age_1979)
+
+age_table %>%
+  mutate(n = scales::comma(n)) %>%
+  kable(
+      caption = "The frequency table of the age at the start of the survey (?CHECK) in the full NSLY79 data",
+      col.names = c("Age", "Number of individuals"),
+      booktabs = TRUE,
+      linesep = "",
+      align = "r") %>%
+  kable_styling()
+
+## ---- gender-race-table
+#' @param x a tabyl with frequency table in "core" attribute
+freq_formatting <- function(x) {
+  attr(x, "core") %>%
+    adorn_totals(c("row", "col")) %>%
+    mutate(across(where(is.numeric), ~scales::comma(.x, 1)))
+}
+
+gender_race_table <- full_demographics %>%
+  tabyl(gender, race) %>%
+  adorn_totals(c("row", "col")) %>%
+  adorn_percentages("row") %>%
+  adorn_pct_formatting(digits = 2) %>%
+  adorn_ns(., position = "front", ns = freq_formatting(.)) %>%
+  mutate(gender = str_to_title(gender))
+
+kable(gender_race_table,
+      caption = "The contingency table for gender and race for the full NLSY79 data.",
+      col.names = c("Gender", "Hispanic", "Black", "Non-Black, Non-Hispanic", "Total"),
+      booktabs = TRUE,
+      linesep = "",
+      align = "lrrrr") %>%
+  kable_classic() %>%
+  row_spec(2, hline_after = TRUE) %>%
+  add_header_above(c(" " = 1, "Race" = 3, " " = 1))
+
+## ---- summarytable
+kable(as.array(summary(wages_before$mean_hourly_wage)),
+      caption = "Summary statistics of the hourly wages from the high school dropout subcohort of NLSY79 data",
+      col.names = c("Statistics", "Hourly wage ($)"),
+      booktabs = TRUE,
+      linesep = "") %>%
+  kable_styling()
+
+## ---- sample-plot
+wages_before_tsibble <- as_tsibble(x = wages_before,
+                                     key = id,
+                                     index = year,
+                                     regular = FALSE)
+
+set.seed(20210225)
+ggplot(wages_before_tsibble,
+       aes(x = year,
[…]

### 2021-09-28 app/app.Rmd
paper + app folder added, vignette updated

---
title: "Customize Your Threshold"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
library(shiny)
library(shinyWidgets)
```

```{r data}
wages_hs <- readRDS(here::here("app/wages_hs.rds")) %>%
    dplyr::select(-`X1.x`, -`X1.y`)
```

Inputs {.sidebar}
-----------------------------------------------------------------------
<br/>

<style>
div.blue { background-color:#FADBD8; border-radius: 10px; padding: 10px;}
</style>
<div class = "blue">

```{r}
p(icon("info-circle", lib = "font-awesome"), strong("About"))
```

The wages data in `yowie` package uses the weight of robust linear model to detect anomalies and imputes those anomalies with its fitted values. The threshold of weight used is 0.12 to maintain the natural variability of the data. Here, you can customize your own threshold.

</div>

<br/>

```{r}
sliderInput("th", "Threshold:",
            min = 0.01,
            max = 1,
            value = 5)
```

<br/>

Row {data-height=400}
-------------------------------------

### Spaghetti plots

```{r}

thres_data_all <- reactive({
  temp_data <- wages_hs %>%
    mutate(wages_rlm = ifelse(w < input$th  & .fitted >= 0, .fitted,
                                      mean_hourly_wage)) %>%
    mutate(is_pred = ifelse(w < input$th & .fitted >= 0, TRUE, FALSE))

  wages_compare <- temp_data %>%
    select(id, year, mean_hourly_wage, wages_rlm) %>%
    rename(original = mean_hourly_wage,
           imputed = wages_rlm) %>%
    pivot_longer(c(-id, -year), names_to = "type", values_to = "wages") %>%
    mutate(type = factor(type, levels = c("original", "imputed")))
  })

renderPlot({
  ggplot(thres_data_all()) +
    geom_line(aes(x = year,
                  y = wages,
                  group = id),
              alpha = 0.1) +
    theme_bw() +
    facet_wrap(~type, scales = "free_y") +
    theme(strip.text = element_text(size = 12))
  })
```

### Summary Statistics

```{r}
max_ori <- max(wages_hs$mean_hourly_wage)
min_ori <- round(min(wages_hs$mean_hourly_wage), 2)
med_ori <- round(median(wages_hs$mean_hourly_wage), 2)
mean_ori <- round(mean(wages_hs$mean_hourly_wage), 2)
```

```{r}
pred <- reactive({
  temp_data <- wages_hs %>%
    mutate(wages_rlm = ifelse(w < input$th  & .fitted >= 0, .fitted,
                                      mean_hourly_wage)) %>%
    mutate(is_pred = ifelse(w < input$th & .fitted >= 0, TRUE, FALSE))
  })
```

<style>
div.blue { background-color:#FADBD8; border-radius: 10px; padding: 10px;}
</style>
<div class = "blue">

```{r}
p(icon("info-circle", lib = "font-awesome"), "Summary statistics of wages in imputed data")
renderText({paste0("Minimum = ", round(min(pred()$wages_rlm), 2), " (Original = 0.01)")})
renderText({paste0("Median = ", round(median(pred()$wages_rlm), 2), " (Original = 7.2)")})
renderText({paste0("Mean = ", round(mean(pred()$wages_rlm), 2), " (Original = 11.87)")})
renderText({paste0("Maximum = ", round(max(pred()$wages_rlm), 2), " (Original = 33400)")})
```

</div>

### Box

```{r}
renderValueBox({
  n_pred <- round(nrow(filter(pred(), is_pred == TRUE))/nrow(pred())*100,2)
  valueBox(
    value = n_pred,
    caption = "percentage of imputed observations",
    icon = "fa-area-chart"
  )
})

```

Row {data-height=650}
-------------------------------------

### Individual plots

```{r}
actionButton("resample", "Click here to sample individuals to display")
```

```{r}

compare_data <- eventReactive(input$resample, {
  sample_id <- sample(unique(wages_hs$id), 16)
  sample <- subset(wages_hs, id %in% sample_id)
})

thres_data <- reactive({
  temp_data <- compare_data() %>%
    mutate(wages_rlm = ifelse(w < input$th  & .fitted >= 0, .fitted,
                                      mean_hourly_wage)) %>%
    mutate(is_pred = ifelse(w < input$th & .fitted >= 0, TRUE, FALSE))

  wages_compare <- temp_data %>%
    select(id, year, mean_hourly_wage, wages_rlm) %>%
    rename(original = mean_hourly_wage,
           imputed = wages_rlm) %>%
    pivot_longer(c(-id, -year), names_to = "type", values_to = "wages") %>%
    mutate(type = factor(type, levels = c("original", "imputed")))
  })

renderPlot({
  ggplot(thres_data()) +
    geom_line(aes(x = year,
                  y = wages,
                  colour = type,
                  linetype = type),
              alpha = 1,
              size = 1) +
    geom_point(aes(x = year,
                   y = wages,
                   colour = type),
               alpha = 0.5,
               size = 1) +
    scale_color_manual(values = c("#229954", "#DC7633")) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 10, size = 5),
          legend.position = "bottom",
          legend.text = element_text(size = 12),
          legend.key.size = unit(6,"line")) +
    facet_wrap(~id, scales = "free_y")
  })

```

### 2021-09-28 paper/paper-html/paper-html.Rmd
paper + app folder added, vignette updated

---
title: "The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth"
author:
  - name: Dewi Amaliah
    thanks: ""
    department: Department of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: dama0007@student.monash.edu
  - name: Dianne Cook
    department: Department of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: dicook@monash.edu
  - name: Emi Tanaka
    department: Department of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: emi.tanaka@monash.edu
  - name: Kate Hyde
    thanks: ""
    department: Department of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: ""
  - name: Nicholas Tierney
    thanks: ""
    department: ""
    affiliation: ""
    location: ""
    email: ""
bibliography: ../references.bib
biblio-style: unsrt
keyword: Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression
abstract: >
  NSLY79 is a prominent open data source that has been an important support for educational purposes and multidisciplinary research on longitudinal data. Subsets of this data can be found in numerous textbooks and research articles. However, the steps and decisions taken to get from the original data to these subsets is never clearly articulated. This article describes our journey when trying to re-create a textbook example data set from the original database. Ultimately, the goal is to provide the resources so that refreshing data subsets can be done on a regular basis, especially when updates are made to the database. Thus, this paper demonstrates the process of initial data analysis (IDA) -- tidying, cleaning, and documenting the process -- to make data available for text book examples or research. A brief exploratory analysis of the resulting data is also conducted. Three new data sets, and the code to produce them are provided in an accompanying open source R package, called `yowie`. As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users. 
output: 
  bookdown::html_document2:
    theme: paper
---

```{r, include = FALSE}
knitr::opts_chunk$set(dev = "png", fig.ext = "png")
```

```{r, child = "../paper.Rmd"}
```

### 2021-09-28 paper/paper-jsdse/paper-jsdse.Rmd
paper + app folder added, vignette updated

---
authors:
  - name: Dewi Amaliah
    thanks: "Corresponding author, dlamaleeah@gmail.com"
    department: Dept of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: dlamaleeah@gmail.com
  - name: Dianne Cook
    department: Dept of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: dicook@monash.edu
  - name: Emi Tanaka
    department: Dept of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: emi.tanaka@monash.edu
  - name: Kate Hyde
    thanks: ""
    department: Dept of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: ""
  - name: Nicholas Tierney
    thanks: ""
    department: Telethon Kids Institute
    affiliation: Perth, WA
    location: "Perth, WA"
    email: ""
bibliography: ../references.bib
biblio-style: unsrt
#preamble: >
output: 
  bookdown::pdf_document2:
    template: TAS-template-for-rmd.tex
    keep_tex: true
date: "`r Sys.Date()`"
always_allow_html: true
---

```{r, include = FALSE}
paper <- here::here("paper/paper.Rmd")
meta <- rmarkdown::yaml_front_matter(paper)
```

---
title: "`r meta$title`"
abstract: >
  `r meta$abstract`
keywords: `r meta$keywords`
---

```{r, child = "../paper.Rmd"}
```

### 2021-09-28 paper/paper-jsdse/paper-jsdse.tex
paper + app folder added, vignette updated

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{tcolorbox}
\usepackage{fontawesome}
\usepackage{booktabs}
\usepackage{url} % not crucial - just used below for the URL
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% need this for pandoc %
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

%-------------------------------------------------------------------
% R MARKDOWN
%-------------------------------------------------------------------
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

\begin{document}

%\bibliographystyle{natbib}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth}
  \author{
            Dewi Amaliah\thanks{Corresponding author, \href{mailto:dlamaleeah@gmail.com}{\nolinkurl{dlamaleeah@gmail.com}}}\\
      Dept of Econometrics and Business Statistics, Monash University, \\
            Dianne Cook\\
      Dept of Econometrics and Business Statistics, Monash University, \\
            Emi Tanaka\\
      Dept of Econometrics and Business Statistics, Monash University, \\
            Kate Hyde\\
      Dept of Econometrics and Business Statistics, Monash University, \\
            Nicholas Tierney\\
      Telethon Kids Institute, Perth, WA, }
        \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Title}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
The National Longitudinal Survey of Youth (NLSY79) is a prominent open data source that has been important for educational purposes and multidisciplinary research on longitudinal data. Subsets of this data can be found in numerous textbooks and research articles. However, the steps and decisions taken to get from the raw data to the textbook data is never clearly articulated. This article describes our journey when trying to re-create a textbook example data set from the original database, with the goal being to refresh the textbook data more regularly. Thus, this paper demonstrates the process -- extracting, tidying, cleaning, and exploring with documentation -- to make refreshed data available for education or research. Three new data sets, and the code to produce them are provided in an accompanying open source R package, called \texttt{yowie}. As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
\end{abstract}

\noindent%
{\it Keywords:}  Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!

\hypertarget{intro}{%
\section{Introduction}\label{intro}}

``Open data'' is data that is freely accessible, modifiable, and shareable by anyone for any purpose (Open Knowledge Foundation 2021). This type of data can be useful as example data in statistical textbooks and for research purposes. However, open data is often referred to as what we might call ``wild data'' because it requires substantial cleaning and tidying to tame it into textbook shape. Huebner, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative and essential for the integrity of downstream statistical analyses and model building (Huebner et al. 2020).
Data cleaning can be considered to be a part of what is called ``initial data analysis'' (IDA) (Chatfield 1985). In IDA one would also explore the data, especially to check if the data is consistent with assumptions required for modeling. This is also related to exploratory data analysis (EDA), coined by Tukey (1977) with a focus on learning from data. EDA can be considered to encompass IDA. Dasu and Johnson (2003) say that data cleaning and exploration, without naming it as IDA, is a difficult task and consumes 80\% of the data mining task.

Despite it's importance, this IDA stage is often undervalued and neglected (Chatfield 1985). There are few research papers that document the data cleaning (Wickham 2014). Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way and is only shared among restricted parties (Huebner et al. 2020).

This paper demonstrates the steps of IDA, and documents the process, for a prominent open data source, National Longitudinal Survey of Youth 1979, henceforth referred to NLSY79. This data has been playing an important role in research in various disciplines, including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century (Pergamit et al. 2001). In addition, this is considered a carefully designed longitudinal survey with high retention rates, making it suitable for life course research (Pergamit et al. 2001; Cooksey 2017). According to Cooksey (2017), thousands of articles, and hundreds of book chapters and monographs have utilized this data. Moreover, the NLSY79 is considered the most widely used and most important cohort in the survey data (Pergamit et al. 2001).

Singer and Willett (2003) used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling of wages on workforce experience, with covariates education and race. Our aim is to refresh this textbook data with data from 1994 through to the latest data reported in 2018. Here, we investigate the process of getting from the raw NLSY79 data to a textbook data set as similar as possible to that provided by Singer and Willett (2003). We should also note that race is a variable in the original data set, and for compatibility it is also provided with the refreshed data, for the purposes of studying racism, not race (Fullilove 1998),

Our process of cleaning builds heavily on the tidyverse approach (H. Wickham, Averick, et al. 2019). The data is first organised into ``tidy data'' (Wickham 2014) and then further wrangled using the data pipeline and split-apply-combine approach (Wickham 2011). The resulting data is provided in a new R package called \texttt{yowie} which includes the code so that the process is reproducible, and could be used to further refresh the data as new records are made available in the NLSY79 database.

This paper is structured in the following way. Section \ref{database} describes the NLSY79 data source. Section \ref{cleaning} presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and IDA to find and repair anomalies. Our final subset is compared the old textbook subset in Section \ref{compare}. Finally, Section \ref{summary} summarizes the contribution and makes recommendations for the NLSY79 data curators.

\hypertarget{database}{%
\section{The NLSY79}\label{database}}

\hypertarget{database-1}{%
\subsection{Database}\label{database-1}}

The NLSY79 is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth born between 1957-1964 (The U.S. Bureau of Labor Statistics, n.d.). The cohort originally included 12,686 respondents aged 14-22 when first interviewed in 1979. It comprised of Blacks, Hispanics, economically disadvantaged non-Black non-Hispanics, and youth in the military. In 1984 and 1990, two sub-samples were dropped from the interview; the dropped subjects are the 1,079 members of the military sample and 1,643 members of the economically disadvantaged non-Black non-Hispanics, respectively. Hence, 9,964 respondents remain in the eligible samples. The surveys were conducted annually from 1979 to 1994 and biennially thereafter. Data are currently available from Round 1 (1979 survey year) to Round 28 (2018 survey year).

Although the main focus area of the NLSY is labor and employment, the NLSY also covers several other topics, including education; training and achievement; household, geography, and contextual variables; dating, marriage, cohabitation; sexual activity, pregnancy, and fertility; children; income, assets and program participation; health; attitudes and expectations; and crime and substance use.

There are two ways to conduct the interview of the NLSY79, which are face-to-face interviews or by telephone. In recent survey years, more than 90 percent of respondents were interviewed by telephone (Cooksey 2017).

\hypertarget{target}{%
\subsection{Target data}\label{target}}

The NLSY79 data used in Singer and Willett (2003). contains the longitudinal measurements on yearly mean hourly wages with years of workforce experience, and demographic variables education and race, from 1979 through to 1994. In addition, the cohort is restricted to male high-school dropouts who first participated in the study at age 14-17 years. Thus the target data set is to collect the same variables for the extended time frame of 1994 through to 2018, the most recent year reported. .

\hypertarget{cleaning}{%
\section{Data cleaning}\label{cleaning}}

van der Loo and de Jonge (2018) describe the notion of a ``statistical value chain'' where the production stages of data cleaning process are earmarked as raw data (data from the data source), input data (data organised with correct type and identified variables) and valid data (data that faithfully represent the variables). In this section, we outline the steps to download the raw data (Section \ref{getdata}) and then tidying the raw data into input data, specifically for wages data (Section \ref{tidywages}), demographic variables (Section \ref{tidydemog}) and the employment variables (Section \ref{tidyemp}), so that the resulting input data can be used downstream for validating the data as described in Section \ref{ida}.

\hypertarget{getdata}{%
\subsection{Getting the data}\label{getdata}}

The NLSY79 data contains a large scope of variables but for our aim, the scope required is limited to demographic profiles and wages data. More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}, clicked on the direct link to NSLY79 data and navigated as described below.

\begin{tcolorbox}[label="box:data", title=Navigating the data source to download the raw data]

\faDatabase~NLSY79 (\url{https://www.nlsinfo.org/investigator/pages/search?s=NLSY79})\\
\vspace{1mm}
\faCheck~The CASEID will be always be selected.\\
\vspace{1mm}
\faCheck~The 3 recommended demographic variable (sample ID, race and sex) were selected.\\
\vspace{1mm}
For the remaining variables, we went to the ``Variable Search'' tab and select variables as follows

\begin{itemize}
\item[$\triangleright$] Education, Training and Achievement Scores
\begin{itemize}
\item[$\triangleright$] Education $\triangleright$ Summary measures $\triangleright$ All schools $\triangleright$ By year $\triangleright$ Highest grade completed
\begin{itemize}
\item[\faCheck] All 80 variables in Highest grade completed were selected.
\end{itemize}
\end{itemize}
\item[$\triangleright$] Employment
\begin{itemize}
\item[$\triangleright$] Summary measures $\triangleright$ By job
\begin{itemize}
\item[$\triangleright$] Hours worked  
\begin{itemize}
\item[\faCheck] All 442 primary variables in Hours worked were selected.
\end{itemize}
\item[$\triangleright$] Hourly wages
\begin{itemize}
\item[\faCheck] All 156 variables in Hourly wages were selected.
\end{itemize}
\end{itemize}
\end{itemize}
\item[$\triangleright$] Household, Geography and Contextual Variables
\begin{itemize}
\item[$\triangleright$] Context $\triangleright$ Summary measures $\triangleright$ Basic demographics $\triangleright$ Date of birth
\begin{itemize}
\item[\faCheck] All 4 variables in Date of birth were selected. 
\end{itemize}
\end{itemize}
\end{itemize}
\begin{itemize}
\item[\faCloudDownload] To download all 686 variables selected, we then navigate to the tab "Save / Download" then select the tab "Advanced Download". We select the R Source code, Codebook of selected variables, Short Description File and Comma-delimited datafile of selected variables with Reference Number as column headers. We name the filename "NLSY79" and press the download button.
\end{itemize}

\end{tcolorbox}

The downloaded data set comes as a zip file, containing the following set of files:

\begin{itemize}
\tightlist
\item
  \texttt{NLSY79.csv}: csv format of the response data.
\item
  \texttt{NLSY79.dat}: dat format of the response data.
\item
  \texttt{NLSY79.NLSY79}: Tagset of variables that can be uploaded to the website to recreate the data set.
\item
  \texttt{NLSY79.R}: R script for reading the data into R and converting the variables' names and label into something more sensible. We utilize this code to do an initial data tidying which produces two data set, \texttt{categories\_qnames} (where the observations are stored in categorical/interval values) and \texttt{new\_data\_qnames} (the observations are stored in integer form).
\end{itemize}

\hypertarget{tidywages}{%
\subsubsection{Tidying the wages data}\label{tidywages}}

The raw data, \texttt{new\_data\_qnames}, is organised such that each row corresponds to an individual. As respondents can have multiple jobs at specific years, the column names, such as \texttt{HRP1\_1979}, \texttt{HRP2\_1979}, \texttt{HRP1\_1980} and \texttt{HRP2\_1980}, contain the information about the job number up to 5 (\texttt{HRP1} = job 1, \texttt{HRP2} = job 2) and the year. The raw data consequently has a large number of columns (686 to be specific). The values in the cell under the variables that begin with \texttt{HRP} correspond to the hourly wage in dollars. A glimpse of this data output is shown below.

\begin{verbatim}
#> 'data.frame':    12686 obs. of  686 variables:
#>  $ CASEID_1979                  : int  1 2 3 4 5 6 7 8 9 10 ...
#>  $ HRP1_1979                    : int  328 385 365 NA 310 NA NA NA 214 NA ...
#>  $ HRP2_1979                    : int  NA NA NA NA 375 NA NA NA NA NA ...
#>  $ HRP3_1979                    : int  NA NA 275 NA NA NA NA NA NA NA ...
#>  $ HRP4_1979                    : int  NA NA NA NA NA 250 NA NA NA NA ...
#>  $ HRP5_1979                    : int  NA NA NA NA NA NA NA NA NA NA ...
#>  $ HRP1_1980                    : int  NA 457 397 NA 333 275 300 394 200 318 ...
#>  $ HRP2_1980                    : int  NA NA 367 NA NA NA NA NA NA NA ...
#>  $ HRP3_1980                    : int  NA NA 380 NA NA NA 290 NA NA NA ...
#>  $ HRP4_1980                    : int  NA NA NA NA NA NA NA NA NA NA ...
#>   [list output truncated]
\end{verbatim}

According to Wickham (2014), tidy data sets comply with three rules: (i) each variable forms a column, (ii) each observation forms a row, and (iii) each type of observational unit forms a table. The raw data does not comply these rules, thus we re-arrange and wrangle the data into tidy data form, columns corresponding to individual ID, year, job number, wage in dollars and the demographic variables. This is done using the \texttt{tidyverse} suite of packages (H. Wickham, Averick, et al. 2019), \texttt{tidyr} (H. Wickham 2020b) to pivot the data into long form, \texttt{dplyr} (H. Wickham, François, et al. 2020), and \texttt{stringr} (H. Wickham 2019) for creating new variables, and levels of factors by text wrangling. The long form of the data makes it possible to do these data transformations efficiently, and it is an intermediate step towards the final target data.

\hypertarget{tidydemog}{%
\subsubsection{Tidying demographic variables}\label{tidydemog}}

\texttt{Age\ in\ 1979}, \texttt{gender}, \texttt{race}, \texttt{highest\ grade\ completed}, and the \texttt{year\ when\ the\ highest\ grade\ completed} are the variables that we want to use in the cleaned data set.

Age in 1979 are derived from year and month of birth variables in the raw data. The variables have two versions, which are the 1979 version and the 1981 version. We only used the 1979 data and did a consistency check of those years and flag the inconsistent observations.

For \texttt{gender} and \texttt{race}, we only rename the column names. It is worth noting that using these two variable needs special attention. Gender, as reported in the data, only has two categories, which is recognised today as inadequate. Gender is not binary. Further, race is as reported in the database. When doing analysis with this variable, one should keep in mind that the purpose is to study racism rather than race.

The next step is tidying the \texttt{highest\ grade\ completed} variable. This variable came with several version in each year. We choose the revised May data because it seemed to have less missing and presumably has been checked. However, there is no revised May data for 2012, 2014, 2016, and 2018. Thus, we just use the ordinary May data for these years. The code for tidying this variable is provided in the supplementary materials.

Further, \texttt{highest\ grade\ completed} is measured and could be updated in each period of the survey. Hence, we only used the highest grade completed ever and derived the year when it is completed. The code for tidying this variable is also available in the supplementary materials.

Finally, we get all of the demographic profiles of the NLSY79 cohort. We then save this data as \texttt{demog\_nlsy79}.

\hypertarget{tidyemp}{%
\subsubsection{Tidying employment variables}\label{tidyemp}}

The employment data comprises of three variables, i.e., \texttt{total\ hours\ of\ work\ per\ week}, \texttt{number\ of\ jobs\ that\ an\ individual\ has}, and \texttt{mean\ hourly\ wage}.

\texttt{hours\ worked\ per\ week} initially has only one version per job, no choice from 1979 to 1987 (QES-52A). From 1988 onward, when we had more options, we chose the variable for total hours, including time spent working from home (QES-52D). However, in 1993, this variable did not have all the five D variables (the first one and the last one were missing), so we used QES-52A variable instead. In addition, 2008 only had jobs 1-4 for the QES-52D variable (whereas the other years had 1-5), so we use these.

The same way was also deployed to tidy the rate of wage by year and by ID. The difference is that the hourly rate has only one version of each year. The hours of work and the hourly rate are then joined to calculate the number of jobs that a respondent has and their mean hourly wage. Some observations have 0 in their hourly rate and work hours, which is considered an invalid value. Thus, these observations are set to be NA. Besides, there are some observations with unusual hours of work. Here, we also censored the observations with \textgreater{} 84 hours of work a week to be NA.

Since our ultimate goal is to calculate the mean hourly wage, the number of jobs is calculated based on the availability of the \texttt{rate\_per\_hour} information. For example, the number of jobs of ID 1, based on \texttt{hours\_work}, is 2. However, since the information of hourly rate of \texttt{job\_02} is not available, the number of jobs is considered as 1.

Further, we calculated the mean hourly wage for each ID in each year using a weighted mean with the hours of work as the weight. However, there is a lot of missing value in \texttt{hours\_work} variable. In that case, we only calculated the mean hourly wage based on the arithmetic/regular mean method. Hence, we created a new variable to flag whether the mean hourly wage is weighted or regular. Additionally, if an ID only had one job, we directly used their hourly wages information and flagged it as arithmetic mean.

\begin{verbatim}
#> # A tibble: 10 x 6
#>       id  year mean_hourly_wage total_hours number_of_jobs is_wm
#>    <int> <dbl>            <dbl>       <int>          <dbl> <lgl>
#>  1     1  1979             3.28          38              1 FALSE
#>  2     1  1981             3.61          NA              1 FALSE
#>  3     2  1979             3.85          35              1 FALSE
#>  4     2  1980             4.57          NA              1 FALSE
#>  5     2  1981             5.14          NA              1 FALSE
#>  6     2  1982             5.71          35              1 FALSE
#>  7     2  1983             5.71          NA              1 FALSE
#>  8     2  1984             5.14          NA              1 FALSE
#>  9     2  1985             7.71          NA              1 FALSE
#> 10     2  1986             7.69          NA              1 FALSE
\end{verbatim}

The employment and demographic variables are then joined. We also filtered the data to only have the cohort who completed the education up to 12th grade and participated at least five rounds in the survey and save it to an object called \texttt{wages}.

\hypertarget{ida}{%
\subsection{Initial data analysis}\label{ida}}

According to Huebner, Vach, and Cessie (2016), initial data analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis. Moreover, Chatfield (1985) argues that the two main objectives of IDA are data description, which is to assess the structure and the quality of the data, and model formulation without any formal statistical inference.

In this paper, we conduct an IDA or a preliminary data analysis to assess the consistency of the data with the cohort information that the NLSY provides. In addition, we also aim to find the anomaly in the wages values using this approach. We mainly use graphical summaries to do the IDA using \texttt{ggplot2} (Wickham 2016) and \texttt{brolgar} (Tierney, Cook, and Prvan 2020).

As stated previously, the respondents' ages ranged from 12 to 22 when first interviewed in 1979. Hence, we validate whether all of the respondents were in this range. Additionally, the \href{https://www.nlsinfo.org/content/cohorts/nlsy79/intro-to-the-sample/nlsy79-sample-introduction}{NLSY} also provides the number of the survey cohort by their gender (6,403 males and 6,283 females) and race (7,510 Non-Black/Non-Hispanic; 3,174 Black; 2,002 Hispanic). To validate this, we used the \texttt{demog\_nlsy79}, i.e., the data with the survey years 1979 sample. Table \ref{tab:age-table} and Table \ref{tab:gender-race-table} suggest that the demographic data we had is consistent with the sample information in the database.

\begin{table}

\caption{\label{tab:age-table}The frequency table of the age at the start of the survey (?CHECK) in the full NSLY79 data}
\centering
\begin{tabular}[t]{rr}
\toprule
Age & Number of individuals\\
\midrule
15 & 1,265\\
16 & 1,550\\
17 & 1,600\\
18 & 1,530\\
19 & 1,662\\
20 & 1,722\\
21 & 1,677\\
22 & 1,680\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}

\caption{\label{tab:gender-race-table}The contingency table for gender and race for the full NLSY79 data.}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Race} & \multicolumn{1}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){2-4}
Gender & Hispanic & Black & Non-Black, Non-Hispanic & Total\\
\midrule
Male & 1,000 (15.62\%) & 1,613 (25.19\%) & 3,790 (59.19\%) & 6,403 (100.00\%)\\
Female & 1,002 (15.95\%) & 1,561 (24.84\%) & 3,720 (59.21\%) & 6,283 (100.00\%)\\
\midrule
Total & 2,002 (15.78\%) & 3,174 (25.02\%) & 7,510 (59.20\%) & 12,686 (100.00\%)\\
\bottomrule
\end{tabular}
\end{table}

The next step is that explore the mean hourly wage data. In this case, we only explore the wages data of up to \(12^{th}\) grade. Here, we use visualisation techniques to do IDA.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figures/sample-plot-1}

}

\caption{Longitudinal profiles of wages for a random sample of 20 individuals in the refreshed data. Most, but not all, individuals here experienced increasing wages over time, and several have experienced considerable fluctuation in wages. Some individuals are only measured for a short period.}\label{fig:sample-plot}
\end{figure}

We take 36 samples randomly from the data and plot them, as shown in Figure \ref{fig:sample-plot}. It implies that these respondents have a lot of variability in wages, for example, the IDs in panel numbers 5, 7, and 11. The plot also implies that the samples have a different pattern of mean hourly wages. Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9. However, when checking the summary plots (Figure \ref{fig:feature-plot}), we found that some observations had exceptionally high wages. Some of them, for example respondents in Figure \ref {fig:feature-plot} C had experienced unusual wages only in certain year.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figures/feature-plot-1}

}

\caption{Summary plots to check the cleaned data reveal more cleaning is necessary: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual (id=39) with an unusual wage relative to their years of data (C). Some values of hourly wages are unbelievable, and some individuals have extemely unusual wages in some years.}\label{fig:feature-plot}
\end{figure}

The anomalies are also found in the total hours of work, where some observations work for 420 hours a week in total. According to Pergamit et al. (2001), one of the flaws of the NLSY79 employment data is that the NLSY79 collects the information of the working hours since the last interview. Thus, it might be challenging for the respondents to track the within-job hours' changes between survey years, especially for the respondents with fluctuated working hours or seasonal jobs. It even has been more challenging since 1994, where the respondents had to recall two years periods. This shortcoming might also contribute to the fluctuation of one's wages data.

\hypertarget{replacing-extreme-values}{%
\subsubsection{Replacing extreme values}\label{replacing-extreme-values}}

As part of the IDA, which is the model formulation, we build a robust linear regression model to treat the extreme values in the data. Robust linear regression yields an estimation robust to the influence of noise or contamination (Koller 2016). It also aims to detect the contamination by weighting each observation based on how ``well-behaved'' they are, known as robustness weight. Observations with lower robustness weight are suggested as an outlier by this method (Koller 2016).

Since we work with longitudinal data, we build the model for each ID instead of the overall data. The robust mixed model could be the best model to be employed in this case. However, this method is too computationally and memory expensive, especially for a large data set, like the NLSY79 data. Thus, the model for each ID is built utilizing the \texttt{nest} and \texttt{map} function from \texttt{tidyr} (H. Wickham 2020b) and \texttt{purrr} (Henry and Wickham 2020), respectively.

We build the model using the \texttt{rlm} function from \texttt{MASS} package (Venables and Ripley 2002). We set the \texttt{mean\_hourly\_wage} and \texttt{year} as the dependent and predictor, respectively. Furthermore, we use M-Estimation with Huber weighting, where the observation with a slight residual gets a weight of 1, while the larger the residual, the smaller the weight (less than 1) (UCLA: Statistical Consulting Group 2021). However, the challenging part of detecting the anomaly using the robustness weight is determining the weight threshold in which the observations are considered outliers. Moreover, it should be noted that not all the outliers are due to an error. Instead, it might be that one had reasonably increasing or decreasing wages in a particular period.

To minimize the risk of mistakenly identifying an outlier as an ``erroneous outlier'', we simulate some thresholds and study how they affect the data. We find that 0.12 is the most reasonable value to be the threshold to minimize that drawback's risk because it still captures the sensible spikes in the data. In other words, we keep maintaining the natural variability of the wages while minimizing anomalies because of the error in the data recording. After deciding the threshold, we impute the observations whose weights are less than 0.12 with the models' predicted value. We then flag those observations in a new variable called \texttt{is\_pred}.

Figure \ref{fig:compare-plot} shows the mean hourly wage before and after the extreme values are replaced. It implies that the fluctuation can still be observed in the data after the treatment. However, the large spikes, which are considered ``erroneous outliers'', are already eliminated from the data. Hence, the model produces a data set with a more reasonable degree of fluctuation.

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{figures/compare-plot-1}

}

\caption{Comparison between the original (black dots) and the corrected (solid grey) mean hourly wage for a sample of individuals. A robust linear model prediction was used to correct mean hourly wages value. We can see that some extreme spikes, corresponding to implausible wages, have been replaced with values more similar to wages in neighboring years, but otherwise the profiles are not changed. Some spikes might remain when wage vaues are plausible.}\label{fig:compare-plot}
\end{figure}

Further, \ref{fig:fixed-feature-plot} A shows that after eliminating the extreme values, the highest value has decreased to be around \$350. The spikes are still observed but not as extreme as the original data set. In Figure \ref{fig:fixed-feature-plot} B, we plot the three features of mean hourly wages, namely the minimum, median, and maximum value. We still can see some extreme values in maximum wages, but consider it as a natural variability of the data. In \ref{fig:fixed-feature-plot} C, we can see that after imputing the extreme value in ID=39, we can see how the the wages change over the years more clearly.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figures/fixed-feature-plot-1}

}

\caption{Re-make of the summary plots of the fully processed data suggest that it is now in a reasonable state: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual with an unusual wage relative to their years of data (C). }\label{fig:fixed-feature-plot}
\end{figure}

Finally, we save the imputed data and set the appropriate data type for the variables.

\hypertarget{recap}{%
\subsection{Recap}\label{recap}}

Figure \ref{fig:flow-chart} summarizes the steps taken to go from raw to input to valid data (van der Loo and de Jonge 2018) to create a refreshed wages data set.

\begin{figure}

{\centering \includegraphics[width=0.85\linewidth]{/Users/deeamaliah/Documents/summer_project/summer-wages-refresh/paper/figures/flowchart}

}

\caption{The stages of data cleaning from the raw data to get three datasets contained yowie. \# of individuals means the number of respondents included in each stage, while \# of observations means the number of rows in the data. The color represented the stage of data cleaning in statistical value chain (van der Loo \& de Jonge, 2018). Pink, blue, and green represents the raw, input, and valid data, respectively.}\label{fig:flow-chart}
\end{figure}

\hypertarget{compare}{%
\section{Comparison of refreshed with the original data}\label{compare}}

Now is the time to see how close we have come to the original textbook data. The original set contains wages of high school dropouts wages (Singer and Willett 2003) from 1979 through to 1994. The refreshed data set for comparison is also wages on high school drop outs, over 1979 to 2018. The original set is available in the R package \texttt{brolgar} and the refreshed data is available in the R package \texttt{yowie}.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figures/plotting-sw-do-1}

}

\caption{Comparison of original textbook example (A) with refreshed data (B). The original data was inflation-adjusted to 1990 prices and the individual's time of collection was converted to a length of experience in the workforce, which make it difficult to precisely compare the two sets.}\label{fig:plotting-sw-do}
\end{figure}

There are two aspects of the original data that make a direct comparison difficult. The time variable provided is ``experience in the workforce''. It is not clear how this is calculated. One would expect that there is a record of the day the individual first started a job, and this is used to adjust the year of collection. We have not been able to find this information in the database. Secondly, wages were inflation-adjusted to 1990 prices, which is not done for the refreshed data.

Figure \ref{fig:plotting-sw-do} shows a comparison of the two sets. (A direct ID matching is not possible.) There are 888 individuals in the original and 1,188 individuals in the refreshed data. This suggests some individuals were removed in the original data.
In addition, in the original set the racial break down was 246 black, 204 hispanic and 438 white, while in the refreshed data there are 346 black, 219 hispanic and 623 white, so the proportions have not been exactly reproduced. On education, in the refreshed data there are very few individuals with less than a 12th grade education, which varies from the original. In the original data there were 366 individuals with up to 8th grade and 522 9th-12th grade, as compared to 967 with 12th grade in the refreshed data.

\hypertarget{summary}{%
\section{Summary}\label{summary}}

This paper has described the stages to take a particular open data set and make it a textbook data set, ready for the classroom or research. In the first stage, we showed the steps performed to get the data from the NLSY79 database. The data format needed converting to tidy format, and this was described. After that, an initial data analysis was conducted to investigate and screen the quality of the data. We found and fixed the anomalous observations in wages using a robust linear regression model. Finally a comparison is made between the original and refreshed data sets.

The data cleaning process is documented and the code has been made available. These provide the opportunity to again refresh the textbook data as updated data is published in the NLSY79 database.
Determining the appropriate robustness weight to threshold the anomalous observations is documented, and a \texttt{shiny} (Chang et al. 2020) app is provided to assist.
The current subset is made available in a new R package, called yowie.

Various difficulties were encountered in trying to refresh the data, which include:

\begin{itemize}
\tightlist
\item
  determining which records should be downloaded from the database.
\item
  there are many errors in the data, e.g.~hourly wages greater than \$30,000 per hour, and there was no information on how these were handled in the original data.
\item
  there is no explicit variable in the database recording high school dropout, which means we needed to compare date of 12th grade with the individual's age.
\item
  calculating experience in the workforce, would require knowing the time of first job within the first year the individual was recorded.
\end{itemize}

Ultimately, the refreshed data is reasonably similar to the original, but unsatisfactorily far from it. The last step required would be to inflation-adjust wages, but this is better to do with each wave of new data added, so that it is relative to the last date in the data.

Finally, for the data providers we recommend that a validation system with clear rules is added on data entry, and that alternative output formats, such as a tidy format would help users make better use of the resource. The problem with many of the wages records is that there are implausible values, or confusion on how to record wages for multiple jobs. These values could be validated with simple checks at data entry. Providing an open data resource also is accompanied with the responsibility that the data, especially data that is as valuable as this, is reliable. Users need to be able to trust the data.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

We would like to thank Aarathy Babu for the insight and discussion during the writing of this paper.

The entire analysis is conducted using \texttt{R} (R Core Team 2020) in \texttt{rstudio} using these packages: \texttt{tidyverse} (H. Wickham, Averick, et al. 2019), \texttt{ggplot2} (Wickham 2016), \texttt{dplyr} (H. Wickham, François, et al. 2020), \texttt{readr} (H. Wickham and Hester 2020), \texttt{tidyr} (H. Wickham 2020b), \texttt{stringr} (H. Wickham 2019), \texttt{purrr} (Henry and Wickham 2020), \texttt{broom} (Robinson, Hayes, and Couch 2021), \texttt{brolgar} (Tierney, Cook, and Prvan 2020), \texttt{patchwork} (Pedersen 2020), \texttt{kableExtra} (Zhu 2019), \texttt{MASS} (Venables and Ripley 2002), \texttt{janitor} (Firke 2020), \texttt{DiagrammeR} (Iannone 2020), \texttt{rsvg} (Ooms 2020), \texttt{webshot} (Chang 2019), \texttt{mgcv} (Wood 2003), \texttt{tsibble} (Wang, Cook, and Hyndman 2020), and \texttt{modelr} (H. Wickham 2020a). The paper are generated using \texttt{knitr} (Xie 2014), \texttt{rmarkdown} (Xie, Dervieux, and Riederer 2020), and \texttt{rticles} (Allaire et al. 2021).

\hypertarget{supplementary-materials}{%
\section{Supplementary Materials}\label{supplementary-materials}}

\begin{itemize}
\item
  \textbf{Codes} : R script to reproduce data tidying and cleaning are available in this \href{https://github.com/numbats/yowie/blob/master/data-raw/data_preprocessing.R}{Github Repository}.
\item
  \textbf{R Package \texttt{yowie}}:\texttt{yowie} is a data container R package that contains 3 datasets, namely the high school mean hourly wage data, high school dropouts mean hourly wage data, and demographic data of the NLSY79 cohort. This package could be accessed \href{https://github.com/numbats/yowie}{here}.
\item
  \textbf{shiny app}: An web interactive \texttt{shiny} app to run a simulation to customize the weight threshold. This app could be accessed \href{https://github.com/numbats/wages-refresh/tree/main/app}{here}.
\end{itemize}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-rticles}{}%
Allaire, JJ, Yihui Xie, R Foundation, Hadley Wickham, Journal of Statistical Software, Ramnath Vaidyanathan, Association for Computing Machinery, et al. 2021. \emph{Rticles: Article Formats for R Markdown}. \url{https://CRAN.R-project.org/package=rticles}.

\leavevmode\hypertarget{ref-webshot}{}%
Chang, Winston. 2019. \emph{Webshot: Take Screenshots of Web Pages}. \url{https://CRAN.R-project.org/package=webshot}.

\leavevmode\hypertarget{ref-shiny}{}%
Chang, Winston, Joe Cheng, JJ Allaire, Yihui Xie, and Jonathan McPherson. 2020. \emph{Shiny: Web Application Framework for R}. \url{https://CRAN.R-project.org/package=shiny}.

\leavevmode\hypertarget{ref-Chatfield1985TIEo}{}%
Chatfield, C. 1985. ``The Initial Examination of Data.'' \emph{Journal of the Royal Statistical Society. Series A. General} 148 (3): 214--53.

\leavevmode\hypertarget{ref-eliznlsy}{}%
Cooksey, Elizabeth C. 2017. ``Using the National Longitudinal Surveys of Youth (Nlsy) to Conduct Life Course Analyses.'' In \emph{Handbook of Life Course Health Development}, edited by Richard M. Lerner Neal Halfon Christoper B. Forrest, 561--77. Cham: Springer. \url{https://doi.org/https://doi.org/10.1007/978-3-319-47143-3_23}.

\leavevmode\hypertarget{ref-DasuTamraparni2003Edma}{}%
Dasu, Tamraparni, and Theodore Johnson. 2003. \emph{Exploratory Data Mining and Data Cleaning}. Wiley Series in Probability and Statistics. Hoboken: WILEY.

\leavevmode\hypertarget{ref-janitor}{}%
Firke, Sam. 2020. \emph{Janitor: Simple Tools for Examining and Cleaning Dirty Data}. \url{https://CRAN.R-project.org/package=janitor}.

\leavevmode\hypertarget{ref-racismnotrace}{}%
Fullilove, M. T. 1998. ``Comment: Abandoning "Race" as a Variable in Public Health Research--an Idea Whose Time Has Come.'' \emph{American Journal of Public Health} 88 (9): 1297--8.

\leavevmode\hypertarget{ref-purrr}{}%
Henry, Lionel, and Hadley Wickham. 2020. \emph{Purrr: Functional Programming Tools}. \url{https://CRAN.R-project.org/package=purrr}.

\leavevmode\hypertarget{ref-HuebnerMariannePhD2016Asat}{}%
Huebner, Marianne, Werner Vach, and Saskia le Cessie. 2016. ``A Systematic Approach to Initial Data Analysis Is Good Research Practice.'' \emph{The Journal of Thoracic and Cardiovascular Surgery} 151 (1): 25--27.

\leavevmode\hypertarget{ref-HuebnerMarianne2020Haar}{}%
Huebner, Marianne, Werner Vach, Saskia le Cessie, Carsten Oliver Schmidt, and Lara Lusa. 2020. ``Hidden Analyses: A Review of Reporting Practice and Recommendations for More Transparent Reporting of Initial Data Analyses.'' \emph{BMC Medical Research Methodology} 20 (1): 61--61.

\leavevmode\hypertarget{ref-DiagrammeR}{}%
Iannone, Richard. 2020. \emph{DiagrammeR: Graph/Network Visualization}. \url{https://CRAN.R-project.org/package=DiagrammeR}.

\leavevmode\hypertarget{ref-KollerManuel2016rARP}{}%
Koller, Manuel. 2016. ``Robustlmm: An R Package for Robust Estimation of Linear Mixed-Effects Models.'' \emph{Journal of Statistical Software} 75 (6): 1--24.

\leavevmode\hypertarget{ref-rsvg}{}%
Ooms, Jeroen. 2020. \emph{Rsvg: Render Svg Images into Pdf, Png, Postscript, or Bitmap Arrays}. \url{https://CRAN.R-project.org/package=rsvg}.

\leavevmode\hypertarget{ref-opendata}{}%
Open Knowledge Foundation. 2021. ``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.'' 2021. \url{http://opendefinition.org/od/2.1/en/}.

\leavevmode\hypertarget{ref-patchwork}{}%
Pedersen, Thomas Lin. 2020. \emph{Patchwork: The Composer of Plots}. \url{https://CRAN.R-project.org/package=patchwork}.

\leavevmode\hypertarget{ref-MichaelRPergamit2001DWTN}{}%
Pergamit, Michael R., Charles R. Pierret, Donna S. Rothstein, and Jonathan R. Veum. 2001. ``Data Watch: The National Longitudinal Surveys.'' \emph{The Journal of Economic Perspectives} 15 (2): 239--53.

\leavevmode\hypertarget{ref-R}{}%
R Core Team. 2020. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.

\leavevmode\hypertarget{ref-broom}{}%
Robinson, David, Alex Hayes, and Simon Couch. 2021. \emph{Broom: Convert Statistical Objects into Tidy Tibbles}. \url{https://CRAN.R-project.org/package=broom}.

\leavevmode\hypertarget{ref-SingerJudithD2003Alda}{}%
Singer, Judith D, and John B Willett. 2003. \emph{Applied Longitudinal Data Analysis: Modeling Change and Event Occurrence}. Oxford u.a: Oxford Univ. Pr.

\leavevmode\hypertarget{ref-nlsy79}{}%
The U.S. Bureau of Labor Statistics. n.d. ``National Longitudinal Survey of Youth 1979.'' Available at \url{https://www.nlsinfo.org/content/cohorts/nlsy79} (2021/25/02).

\leavevmode\hypertarget{ref-brolgar}{}%
Tierney, Nicholas, Di Cook, and Tania Prvan. 2020. \emph{Brolgar: BRowse over Longitudinal Data Graphically and Analytically in R}. \url{https://github.com/njtierney/brolgar}.

\leavevmode\hypertarget{ref-tukey}{}%
Tukey, John W. (John Wilder). 1977. \emph{Exploratory Data Analysis}. Addison-Wesley Series in Behavioral Science. Reading, Mass.: Addison-Wesley Pub. Co.

\leavevmode\hypertarget{ref-rlm}{}%
UCLA: Statistical Consulting Group. 2021. ``Robust Regression \textbar{} R Data Analysis Examples.'' February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.

\leavevmode\hypertarget{ref-LooMarkvander2018Sdcw}{}%
van der Loo, Mark, and Edwin de Jonge. 2018. \emph{Statistical Data Cleaning with Applications in R}.

\leavevmode\hypertarget{ref-mass}{}%
Venables, W. N., and B. D. Ripley. 2002. \emph{Modern Applied Statistics with S}. Fourth. New York: Springer. \url{http://www.stats.ox.ac.uk/pub/MASS4}.

\leavevmode\hypertarget{ref-tsibble}{}%
Wang, Earo, Dianne Cook, and Rob J Hyndman. 2020. ``A New Tidy Data Structure to Support Exploration and Modeling of Temporal Data.'' \emph{Journal of Computational and Graphical Statistics} 29 (3): 466--78. \url{https://doi.org/10.1080/10618600.2019.1695624}.

\leavevmode\hypertarget{ref-plyr}{}%
Wickham, Hadley. 2011. ``The Split-Apply-Combine Strategy for Data Analysis.'' \emph{Journal of Statistical Software, Articles} 40 (1): 1--29. \url{https://doi.org/10.18637/jss.v040.i01}.

\leavevmode\hypertarget{ref-WickhamHadley2014TD}{}%
---------. 2014. ``Tidy Data.'' \emph{Journal of Statistical Software} 59 (10): 1--23.

\leavevmode\hypertarget{ref-ggplot2}{}%
---------. 2016. \emph{Ggplot2: Elegant Graphics for Data Analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.

\leavevmode\hypertarget{ref-stringr}{}%
---------. 2019. \emph{Stringr: Simple, Consistent Wrappers for Common String Operations}. \url{https://CRAN.R-project.org/package=stringr}.

\leavevmode\hypertarget{ref-modelr}{}%
---------. 2020a. \emph{Modelr: Modelling Functions That Work with the Pipe}. \url{https://CRAN.R-project.org/package=modelr}.

\leavevmode\hypertarget{ref-tidyr}{}%
---------. 2020b. \emph{Tidyr: Tidy Messy Data}. \url{https://CRAN.R-project.org/package=tidyr}.

\leavevmode\hypertarget{ref-tidyverse}{}%
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019. ``Welcome to the tidyverse.'' \emph{Journal of Open Source Software} 4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\leavevmode\hypertarget{ref-dplyr}{}%
Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2020. \emph{Dplyr: A Grammar of Data Manipulation}. \url{https://CRAN.R-project.org/package=dplyr}.

\leavevmode\hypertarget{ref-readr}{}%
Wickham, Hadley, and Jim Hester. 2020. \emph{Readr: Read Rectangular Text Data}. \url{https://CRAN.R-project.org/package=readr}.

\leavevmode\hypertarget{ref-mgcv}{}%
Wood, S. N. 2003. ``Thin-Plate Regression Splines.'' \emph{Journal of the Royal Statistical Society (B)} 65 (1): 95--114.

\leavevmode\hypertarget{ref-knitr}{}%
Xie, Yihui. 2014. ``Knitr: A Comprehensive Tool for Reproducible Research in R.'' In \emph{Implementing Reproducible Computational Research}, edited by Victoria Stodden, Friedrich Leisch, and Roger D. Peng. Chapman; Hall/CRC. \url{http://www.crcpress.com/product/isbn/9781466561595}.

\leavevmode\hypertarget{ref-rmarkdown}{}%
Xie, Yihui, Christophe Dervieux, and Emily Riederer. 2020. \emph{R Markdown Cookbook}. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://bookdown.org/yihui/rmarkdown-cookbook}.

\leavevmode\hypertarget{ref-kableExtra}{}%
Zhu, Hao. 2019. \emph{KableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.

\bibliographystyle{Chicago}

\end{document}

### 2021-09-28 paper/paper.Rmd
paper + app folder added, vignette updated

---
title: "The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth"
authors:
  - name: Dewi Amaliah
    thanks: ""
    department: Department of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: dama0007@student.monash.edu
  - name: Dianne Cook
    department: Department of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: dicook@monash.edu
  - name: Emi Tanaka
    department: Department of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: emi.tanaka@monash.edu
  - name: Kate Hyde
    thanks: ""
    department: Department of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: ""
  - name: Nicholas Tierney
    thanks: ""
    department: ""
    affiliation: ""
    location: ""
    email: ""
abstract: > 
 The National Longitudinal Survey of Youth (NLSY79) is a prominent open data source that has been important for educational purposes and multidisciplinary research on longitudinal data. Subsets of this data can be found in numerous textbooks and research articles. However, the steps and decisions taken to get from the raw data to the textbook data is never clearly articulated. This article describes our journey when trying to re-create a textbook example data set from the original database, with the goal being to refresh the textbook data more regularly. Thus, this paper demonstrates the process -- extracting, tidying, cleaning, and exploring with documentation -- to make refreshed data available for education or research. Three new data sets, and the code to produce them are provided in an accompanying open source R package, called `yowie`. As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users. 
bibliography: references.bib
biblio-style: unsrt
#preamble: >
header-includes:
  - \usepackage{tcolorbox}
  - \usepackage{fontawesome}
  - \usepackage{color, colortbl}
output: 
  #rticles::arxiv_article:
    #keep_tex: false
  bookdown::pdf_book:
    base_format: rticles::arxiv_article
date: "24/02/2021"
always_allow_html: true
keywords: "Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression"
---

```{r setup, echo = FALSE, cache = FALSE, include = FALSE}
library(knitr)
opts_chunk$set(echo = FALSE, 
               warning = FALSE, 
               message = FALSE,
               cache = TRUE, 
               cache.path = "cache/",
               fig.retina = 2,
               fig.path = "figures/",
               comment = "#>",
               fig.align = 'center')
read_chunk(here::here("data-raw/data-preprocessing.R"))
read_chunk(here::here("data-raw/ida.R"))
read_chunk(here::here("data-raw/eda.R"))
```

```{r load-pkgs}
```

```{r theme, eval = FALSE}
theme_set(theme_bw(base_size = 18) +
            theme(plot.background = element_rect(fill = 'transparent', colour = NA), axis.line.y = element_line(color = "black", linetype = "solid"),
                  plot.title.position = "plot",
                  plot.title = element_text(size = 24),
                  panel.background  = element_rect(fill = 'transparent', colour = NA),
                  legend.background = element_rect(fill = 'transparent', colour = NA),
                  legend.key        = element_rect(fill = 'transparent', colour = NA)
                  ) )
```

# Introduction {#intro}

"Open data" is data that is freely accessible, modifiable, and shareable by anyone for any purpose [@opendata]. This type of data can be useful as example data in statistical textbooks and for research purposes. However, open data is often referred to as what we might call "wild data" because it requires substantial cleaning and tidying to tame it into textbook shape. @HuebnerMariannePhD2016Asat emphasize that making the data cleaning process accountable and transparent is imperative and essential for the integrity of downstream statistical analyses and model building [@HuebnerMarianne2020Haar]. 
Data cleaning can be considered to be a part of what is called "initial data analysis" (IDA) [@Chatfield1985TIEo]. In IDA one would also explore the data, especially to check if the data is consistent with assumptions required for modeling. This is also related to exploratory data analysis (EDA), coined by @tukey with a focus on learning from data. EDA can be considered to encompass IDA. @DasuTamraparni2003Edma say that data cleaning and exploration, without naming it as IDA, is a difficult task and consumes 80% of the data mining task.

Despite it's importance, this IDA stage is often undervalued and neglected [@Chatfield1985TIEo]. There are few research papers that document the data cleaning [@WickhamHadley2014TD]. Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way and is only shared among restricted parties [@HuebnerMarianne2020Haar].

This paper demonstrates the steps of IDA, and documents the process, for a prominent open data source, National Longitudinal Survey of Youth 1979, henceforth referred to NLSY79. This data has been playing an important role in research in various disciplines, including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century [@MichaelRPergamit2001DWTN]. In addition, this is considered a carefully designed longitudinal survey with high retention rates, making it suitable for life course research [@MichaelRPergamit2001DWTN; @eliznlsy]. According to @eliznlsy, thousands of articles, and hundreds of book chapters and monographs have utilized this data. Moreover, the NLSY79 is considered the most widely used and most important cohort in the survey data [@MichaelRPergamit2001DWTN].

@SingerJudithD2003Alda used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling of wages on workforce experience, with covariates education and race. Our aim is to refresh this textbook data with data from 1994 through to the latest data reported in 2018. Here, we investigate the process of getting from the raw NLSY79 data to a textbook data set as similar as possible to that provided by @SingerJudithD2003Alda. We should also note that race is a variable in the original data set, and for compatibility it is also provided with the refreshed data, for the purposes of studying racism, not race [@racismnotrace],  <!--However, we cannot create the exact same data set as published in their book since we do not have the information of what age threshold they used to determine the high school dropouts.-->

Our process of cleaning builds heavily on the tidyverse approach [@tidyverse]. The data is first organised into "tidy data" [@WickhamHadley2014TD] and then further wrangled using the data pipeline and split-apply-combine approach [@plyr]. The resulting data is provided in a new R package called `yowie` which includes the code so that the process is reproducible, and could be used to further refresh the data as new records are made available in the NLSY79 database.

This paper is structured in the following way. Section \@ref(database) describes the NLSY79 data source. Section \@ref(cleaning) presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and IDA to find and repair anomalies. Our final subset is compared the old textbook subset in Section \@ref(compare). Finally, Section \@ref(summary) summarizes the contribution and makes recommendations for the NLSY79 data curators.

# The NLSY79 {#database}

## Database

The NLSY79 is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth born between 1957-1964 [@nlsy79]. The cohort originally included 12,686 respondents aged 14-22 when first interviewed in 1979. It comprised of Blacks, Hispanics, economically disadvantaged non-Black non-Hispanics, and youth in the military. In 1984 and 1990, two sub-samples were dropped from the interview; the dropped subjects are the 1,079 members of the military sample and 1,643 members of the economically disadvantaged non-Black non-Hispanics, respectively. Hence, 9,964 respondents remain in the eligible samples. The surveys were conducted annually from 1979 to 1994 and biennially thereafter. Data are currently available from Round 1 (1979 survey year) to Round 28 (2018 survey year).

Although the main focus area of the NLSY is labor and employment, the NLSY also covers several other topics, including education; training and achievement; household, geography, and contextual variables; dating, marriage, cohabitation; sexual activity, pregnancy, and fertility; children; income, assets and program participation; health; attitudes and expectations; and crime and substance use.

There are two ways to conduct the interview of the NLSY79, which are face-to-face interviews or by telephone. In recent survey years, more than 90 percent of respondents were interviewed by telephone [@eliznlsy].

## Target data {#target}

The NLSY79 data used in @SingerJudithD2003Alda. contains the longitudinal measurements on yearly mean hourly wages with years of workforce experience, and demographic variables education and race, from 1979 through to 1994. In addition, the cohort is restricted to male high-school dropouts who first participated in the study at age 14-17 years. Thus the target data set is to collect the same variables for the extended time frame of 1994 through to 2018, the most recent year reported. .

<!-- Since no specific criteria on high school dropouts was defined
in @SingerJudithD2003Alda, we define high school dropouts as respondents who only completed either 9th to 11th grade, or who completed 12th when their age is at least 19 years old (meaning that the individual may have dropped out of high school but returned at a later time to complete high school education). XXX This should probably go further down in the flow. -->

# Data cleaning {#cleaning}

@LooMarkvander2018Sdcw describe the notion of a "statistical value chain" where the production stages of data cleaning process are earmarked as raw data (data from the data source), input data (data organised with correct type and identified variables) and valid data (data that faithfully represent the variables).  In this section, we outline the steps to download the raw data (Section \@ref(getdata)) and then tidying the raw data into input data, specifically for wages data (Section \@ref(tidywages)), demographic variables (Section \@ref(tidydemog)) and the employment variables (Section \@ref(tidyemp)), so that the resulting input data can be used downstream for validating the data as described in Section \@ref(ida).

## Getting the data {#getdata}

The NLSY79 data contains a large scope of variables but for our aim, the scope required is limited to demographic profiles and wages data. More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}, clicked on the direct link to NSLY79 data and navigated as described below.

::: {.tcolorbox data-latex="[label=\"box:data\", title=Navigating the data source to download the raw data]" }
\faDatabase\ NLSY79 (\url{https://www.nlsinfo.org/investigator/pages/search?s=NLSY79})  
\vspace{1mm}
\faCheck\ The CASEID will be always be selected.  
\vspace{1mm}
\faCheck\ The 3 recommended demographic variable (sample ID, race and sex) were selected.  
\vspace{1mm}
For the remaining variables, we went to the "Variable Search" tab and select variables as follows
\begin{itemize}
\item[$\triangleright$] Education, Training and Achievement Scores
\begin{itemize}
\item[$\triangleright$] Education $\triangleright$ Summary measures $\triangleright$ All schools $\triangleright$ By year $\triangleright$ Highest grade completed
\begin{itemize}
\item[\faCheck] All 80 variables in Highest grade completed were selected.
\end{itemize}
\end{itemize}
\item[$\triangleright$] Employment
\begin{itemize}
\item[$\triangleright$] Summary measures $\triangleright$ By job
\begin{itemize}
\item[$\triangleright$] Hours worked  
\begin{itemize}
\item[\faCheck] All 442 primary variables in Hours worked were selected.
\end{itemize}
\item[$\triangleright$] Hourly wages
\begin{itemize}
\item[\faCheck] All 156 variables in Hourly wages were selected.
\end{itemize}
\end{itemize}
\end{itemize}
\item[$\triangleright$] Household, Geography and Contextual Variables
\begin{itemize}
\item[$\triangleright$] Context $\triangleright$ Summary measures $\triangleright$ Basic demographics $\triangleright$ Date of birth
\begin{itemize}
\item[\faCheck] All 4 variables in Date of birth were selected. 
\end{itemize}
\end{itemize}
\end{itemize}
\begin{itemize}
\item[\faCloudDownload] To download all 686 variables selected, we then navigate to the tab "Save / Download" then select the tab "Advanced Download". We select the R Source code, Codebook of selected variables, Short Description File and Comma-delimited datafile of selected variables with Reference Number as column headers. We name the filename "NLSY79" and press the download button.
\end{itemize}
:::

The downloaded data set comes as a zip file, containing the following set of files:

- `NLSY79.csv`: csv format of the response data.
- `NLSY79.dat`: dat format of the response data.
- `NLSY79.NLSY79`: Tagset of variables that can be uploaded to the website to recreate the data set.
- `NLSY79.R`: R script for reading the data into R and converting the variables' names and label into something more sensible. We utilize this code to do an initial data tidying which produces two data set,  `categories_qnames` (where the observations are stored in categorical/interval values) and `new_data_qnames` (the observations are stored in integer form).

```{r raw-data}
```

### Tidying the wages data {#tidywages}

The raw data, `new_data_qnames`, is organised such that each row corresponds to an individual. As respondents can have multiple jobs at specific years, the column names, such as `HRP1_1979`, `HRP2_1979`, `HRP1_1980` and `HRP2_1980`, contain the information about the job number up to 5 (`HRP1` = job 1, `HRP2` = job 2) and the year. The raw data consequently has a large number of columns (686 to be specific). The values in the cell under the variables that begin with `HRP` correspond to the hourly wage in dollars. A glimpse of this data output is shown below.

```{r untidy-data, fig.cap="The structure of the NLSY79 raw data on wages for up to five jobs in any year."}
```

According to @WickhamHadley2014TD, tidy data sets comply with three rules: (i) each variable forms a column, (ii) each observation forms a row, and (iii) each type of observational unit forms a table. The raw data does not comply these rules, thus we re-arrange and wrangle the data into tidy data form, columns corresponding to individual ID, year, job number, wage in dollars and the demographic variables. This is done using the `tidyverse` suite of packages [@tidyverse], `tidyr` [@tidyr] to pivot the data into long form, `dplyr` [@dplyr], and `stringr` [@stringr] for creating new variables, and levels of factors by text wrangling. The long form of the data makes it possible to do these data transformations efficiently, and it is an intermediate step towards the final target data.

<!-- required by making use of the `tidyverse` suite of packages [@tidyverse], which include the main set of packages needed, namely `tidyr` [@tidyr], `dplyr` [@dplyr], and `stringr` [@stringr]. The tidy data form, where we have a variable that indicates the individual ID, year, job number, wage in dollars and demographic variables, is important for downstream analysis, such as visualisation and modelling that will be performed later.

[REMOVE] Unfortunately, the `new_data_qnames` did not meet these requirements in the way that data for each year and job for the same respondent is stored as one variable. Hence, the data contains a huge number of columns (686 columns). The example of the untidy data set is displayed in Table \ref{tab:untidy-data}. The table intended to display the hourly rate of each respondent by job (HRP1 to HRP5) and by year (1979 and 1980). The table implies that the column headers are values of the year and job, not variable names. Consequently, the data should be tidied and wrangled first to extract the demographic and employment variables we want to put in the final data set. We mainly used `tidyr` [@tidyr], `dplyr` [@dplyr], and `stringr` [@stringr] to do this job.-->

### Tidying demographic variables {#tidydemog}

`Age in 1979`, `gender`, `race`, `highest grade completed`, and the `year when the highest grade completed` are the variables that we want to use in the cleaned data set.

Age in 1979 are derived from year and month of birth variables in the raw data. The variables have two versions, which are the 1979 version and the 1981 version. We only used the 1979 data and did a consistency check of those years and flag the inconsistent observations.

```{r dob-tidy}
```

For `gender` and `race`, we only rename the column names. It is worth noting that using these two variable needs special attention. Gender, as reported in the data, only has two categories, which is recognised today as inadequate. Gender is not binary. Further, race is as reported in the database. When doing analysis with this variable, one should keep in mind that the purpose is to study racism rather than race.

```{r demog-tidy}
```

The next step is tidying the `highest grade completed` variable. This variable came with several version in each year. We choose the revised May data because it seemed to have less missing and presumably has been checked. However, there is no revised May data for 2012, 2014, 2016, and 2018. Thus, we just use the ordinary May data for these years. The code for tidying this variable is provided in the supplementary materials.

```{r tidy-grade}
```

Further, `highest grade completed` is measured and could be updated in each period of the survey. Hence, we only used the highest grade completed ever and derived the year when it is completed. The code for tidying this variable is also available in the supplementary materials.

```{r tidy-hgc}
```

Finally, we get all of the demographic profiles of the NLSY79 cohort. We then save this data as `demog_nlsy79`.

```{r full-demog}
```

### Tidying employment variables {#tidyemp}

The employment data comprises of three variables, i.e., `total hours of work per week`, `number of jobs that an individual has`, and `mean hourly wage`.

`hours worked per week` initially has only one version per job, no choice from 1979 to 1987 (QES-52A). From 1988 onward, when we had more options, we chose the variable for total hours, including time spent working from home (QES-52D). However, in 1993, this variable did not have all the five D variables (the first one and the last one were missing), so we used QES-52A variable instead. In addition, 2008 only had jobs 1-4 for the QES-52D variable (whereas the other years had 1-5), so we use these.

```{r tidy-hours}
```

The same way was also deployed to tidy the rate of wage by year and by ID. The difference is that the hourly rate has only one version of each year. The hours of work and the hourly rate are then joined to calculate the number of jobs that a respondent has and their mean hourly wage. Some observations have 0 in their hourly rate and work hours, which is considered an invalid value. Thus, these observations are set to be NA. Besides, there are some observations with unusual hours of work. Here, we also censored the observations with > 84 hours of work a week to be NA.

```{r tidy-rate}
```

Since our ultimate goal is to calculate the mean hourly wage, the number of jobs is calculated based on the availability of the `rate_per_hour` information. For example, the number of jobs of ID 1, based on `hours_work`, is 2. However, since the information of hourly rate of `job_02` is not available, the number of jobs is considered as 1.

Further, we calculated the mean hourly wage for each ID in each year using a weighted mean with the hours of work as the weight. However, there is a lot of missing value in `hours_work` variable. In that case, we only calculated the mean hourly wage based on the arithmetic/regular mean method. Hence, we created a new variable to flag whether the mean hourly wage is weighted or regular. Additionally, if an ID only had one job, we directly used their hourly wages information and flagged it as arithmetic mean.

```{r tidy-nojob}
```

The employment and demographic variables are then joined. We also filtered the data to only have the cohort who completed the education up to 12th grade and participated at least five rounds in the survey and save it to an object called `wages`.

```{r wages-demog-hs}
```

## Initial data analysis {#ida}

According to @HuebnerMariannePhD2016Asat, initial data analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis. Moreover, @Chatfield1985TIEo argues that the two main objectives of IDA are data description, which is to assess the structure and the quality of the data, and model formulation without any formal statistical inference.

In this paper, we conduct an IDA or a preliminary data analysis to assess the consistency of the data with the cohort information that the NLSY provides. In addition, we also aim to find the anomaly in the wages values using this approach. We mainly use graphical summaries to do the IDA using `ggplot2` [@ggplot2] and `brolgar` [@brolgar].

As stated previously, the respondents' ages ranged from 12 to 22 when first interviewed in 1979. Hence, we validate whether all of the respondents were in this range. Additionally, the [NLSY](https://www.nlsinfo.org/content/cohorts/nlsy79/intro-to-the-sample/nlsy79-sample-introduction) also provides the number of the survey cohort by their gender (6,403 males and 6,283 females) and race (7,510 Non-Black/Non-Hispanic; 3,174 Black; 2,002 Hispanic). To validate this, we used the `demog_nlsy79`, i.e., the data with the survey years 1979 sample. Table \ref{tab:age-table} and Table \ref{tab:gender-race-table} suggest that the demographic data we had is consistent with the sample information in the database.

```{r age-table}
```

```{r gender-race-table}
```

The next step is that explore the mean hourly wage data. In this case, we only explore the wages data of up to $12^{th}$ grade.  Here, we use visualisation techniques to do IDA.

```{r summarytable, eval=FALSE}
```

```{r sample-plot, echo = FALSE, fig.cap="Longitudinal profiles of wages for a random sample of 20 individuals in the refreshed data. Most, but not all, individuals here experienced increasing wages over time, and several have experienced considerable fluctuation in wages. Some individuals are only measured for a short period.", fig.height=8, fig.width=10, out.width="100%"}
```

We take 36 samples randomly from the data and plot them, as shown in Figure \ref{fig:sample-plot}. It implies that these respondents have a lot of variability in wages, for example, the IDs in panel numbers 5, 7, and 11. The plot also implies that the samples have a different pattern of mean hourly wages. Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9. However, when checking the summary plots (Figure \ref{fig:feature-plot}), we found that some observations had exceptionally high wages. Some of them, for example respondents in Figure \ref {fig:feature-plot} C had experienced unusual wages only in certain year.

```{r feature-plot, fig.cap = "Summary plots to check the cleaned data reveal more cleaning is necessary: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual (id=39) with an unusual wage relative to their years of data (C). Some values of hourly wages are unbelievable, and some individuals have extemely unusual wages in some years.", fig.width=10, fig.height=4, out.width="100%"}
```

```{r high-wages}
```

```{r high-wages, fig.cap= "6 out of 45 IDs with extremely high mean hourly wage. Most of the IDs only have one point of high wage.", eval=FALSE}
```

The anomalies are also found in the total hours of work, where some observations work for 420 hours a week in total. According to @MichaelRPergamit2001DWTN, one of the flaws of the NLSY79 employment data is that the NLSY79 collects the information of the working hours since the last interview. Thus, it might be challenging for the respondents to track the within-job hours' changes between survey years, especially for the respondents with fluctuated working hours or seasonal jobs. It even has been more challenging since 1994, where the respondents had to recall two years periods. This shortcoming might also contribute to the fluctuation of one's wages data.

### Replacing extreme values

As part of the IDA, which is the model formulation, we build a robust linear regression model to treat the extreme values in the data. Robust linear regression yields an estimation robust to the influence of noise or contamination [@KollerManuel2016rARP]. It also aims to detect the contamination by weighting each observation based on how "well-behaved" they are, known as robustness weight. Observations with lower robustness weight are suggested as an outlier by this method [@KollerManuel2016rARP].

Since we work with longitudinal data, we build the model for each ID instead of the overall data. The robust mixed model could be the best model to be employed in this case. However, this method is too computationally and memory expensive, especially for a large data set, like the NLSY79 data. Thus, the model for each ID is built utilizing the `nest` and `map` function from `tidyr` [@tidyr] and `purrr` [@purrr], respectively.

We build the model using the `rlm` function from `MASS` package [@mass]. We set the `mean_hourly_wage` and `year` as the dependent and predictor, respectively. Furthermore, we use M-Estimation with Huber weighting, where the observation with a slight residual gets a weight of 1, while the larger the residual, the smaller the weight (less than 1) [@rlm]. However, the challenging part of detecting the anomaly using the robustness weight is determining the weight threshold in which the observations are considered outliers. Moreover, it should be noted that not all the outliers are due to an error. Instead, it might be that one had reasonably increasing or decreasing wages in a particular period.

To minimize the risk of mistakenly identifying an outlier as an "erroneous outlier", we simulate some thresholds and study how they affect the data. We find that 0.12 is the most reasonable value to be the threshold to minimize that drawback's risk because it still captures the sensible spikes in the data. In other words, we keep maintaining the natural variability of the wages while minimizing anomalies because of the error in the data recording. After deciding the threshold, we impute the observations whose weights are less than 0.12 with the models' predicted value. We then flag those observations in a new variable called `is_pred`.

```{r rlm, eval=FALSE}
```

Figure \ref{fig:compare-plot} shows the mean hourly wage before and after the extreme values are replaced. It implies that the fluctuation can still be observed in the data after the treatment. However, the large spikes, which are considered "erroneous outliers", are already eliminated from the data. Hence, the model produces a data set with a more reasonable degree of fluctuation.

```{r compare-data}
```

```{r compare-plot, fig.cap="Comparison between the original (black dots) and the corrected (solid grey) mean hourly wage for a sample of individuals. A robust linear model prediction was used to correct mean hourly wages value. We can see that some extreme spikes, corresponding to implausible wages, have been replaced with values more similar to wages in neighboring years, but otherwise the profiles are not changed. Some spikes might remain when wage vaues are plausible.", fig.height=8, fig.width=10, out.width="90%"}
```

Further, \ref{fig:fixed-feature-plot} A shows that after eliminating the extreme values, the highest value has decreased to be around $350. The spikes are still observed but not as extreme as the original data set. In Figure \ref{fig:fixed-feature-plot} B, we plot the three features of mean hourly wages, namely the minimum, median, and maximum value. We still can see some extreme values in maximum wages, but consider it as a natural variability of the data. In \ref{fig:fixed-feature-plot} C, we can see that after imputing the extreme value in ID=39, we can see how the the wages change over the years more clearly.

```{r fixed-feature-plot, fig.cap = "Re-make of the summary plots of the fully processed data suggest that it is now in a reasonable state: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual with an unusual wage relative to their years of data (C). ", fig.width=10, fig.height=4, out.width="100%"}
```

Finally, we save the imputed data and set the appropriate data type for the variables.

```{r save-data}
```

```{r nrow}
```

## Recap

Figure \ref{fig:flow-chart} summarizes the steps taken to go from raw to input to valid data [@LooMarkvander2018Sdcw] to create a refreshed wages data set.

<!--To summarise, the  a workflow from getting the data from the NSLY79 database until a ready-analysed data. Referring to @LooMarkvander2018Sdcw, this section shows  part of statistical value chain, which is data cleaning from raw, input, and valid data. The raw demographic and employment data from the database are not in a tidy format. Hence, we perform data wrangling to get the input data. We inspect the input data to find eligible observations to the target data. One of the target data is wages data of respondents who completed up to 12th grade and participated at least in 5 rounds of survey. This needs us to join the extracted demographic and employment data.

Further, using IDA, we inspect on the existence of anomalies/extreme values in the data using the weight from robust linear model. We then replace these extreme values using their predicted value. 
From this stage forward, we assume that we already have valid data. We also subset the data to get the wages of high school dropouts. We then make these three data sets and the cleaning workflow documentation publicly available through an R data container package called `yowie`. A visualisation of these process is displayed in  Figure \ref{fig:flow-chart}.-->

```{r flow-chart, fig.cap="The stages of data cleaning from the raw data to get three datasets contained yowie. \\# of individuals means the number of respondents included in each stage, while \\# of observations means the number of rows in the data. The color represents the stage of data cleaning in statistical value chain (van der Loo \\& de Jonge, 2018). Pink, blue, and green represents the raw, input, and valid data, respectively.", out.width="85%"}
```

# Comparison of refreshed with the original data {#compare}

```{r sw_wages}
```

```{r do_refreshed}
```

```{r compare-sw-do}
```

Now is the time to see how close we have come to the original textbook data. The original set contains wages of high school dropouts wages [@SingerJudithD2003Alda] from 1979 through to 1994. The refreshed data set for comparison is also wages on high school drop outs, over 1979 to 2018. The original set is available in the R package `brolgar` and the refreshed data is available in the R package `yowie`.

```{r plotting-sw-do, fig.cap = "Comparison of original textbook example (A) with refreshed data (B). The original data was inflation-adjusted to 1990 prices and the individual's time of collection was converted to a length of experience in the workforce, which make it difficult to precisely compare the two sets.", fig.width = 9, fig.height = 5, out.width = "100%"}
```

There are two aspects of the original data that make a direct comparison difficult. The time variable provided is "experience in the workforce". It is not clear how this is calculated. One would expect that there is a record of the day the individual first started a job, and this is used to adjust the year of collection. We have not been able to find this information in the database. Secondly, wages were inflation-adjusted to 1990 prices, which is not done for the refreshed data.

Figure \ref{fig:plotting-sw-do} shows a comparison of the two sets. (A direct ID matching is not possible.) There are 888 individuals in the original and 1,188 individuals in the refreshed data. This suggests some individuals were removed in the original data.
In addition, in the original set the racial break down was 246 black, 204 hispanic and 438 white, while in the refreshed data there are 346 black, 219 hispanic and 623 white, so the proportions have not been exactly reproduced. On education, in the refreshed data there are very few individuals with less than a 12th grade education, which varies from the original. In the original data there were 366 individuals with up to 8th grade and 522 9th-12th grade, as compared to 967 with 12th grade in the refreshed data.

```{r summaries, eval=FALSE}
```

<!-- Hence, in this section, we will compare how close the `wages_hs_do`, henceforth refreshed data, to the textbook data. It is also worth noting that the textbook data is reused from @MurnaneRichardJ1999DMDB’s paper. Hence, it is not directly extracted from the NLSY79 database.  @SingerJudithD2003Alda do not state the time frame covered by the data. However, from @MurnaneRichardJ1999DMDB’s paper, we find that the textbook data cover the survey period from 1979-1994. Hence, for comparison purposes, we filter the refreshed data only until 1994.

In the absence of highs chool dropouts criteria in the textbook data, we create a clearer criteria of high school dropouts as mentioned in Section \@ref(#target). Consequently, we do not have the same ID of high-school dropouts as in the textbook data. The textbook data has `r nrow(sw_id)` IDs included, while the refreshed data has `r nrow(do_id)` IDs. Furthermore, there are only `r nrow(sw_do)` IDs included both in the textbook and the refreshed data. After observing @MurnaneRichardJ1999DMDB’s paper, it is mentioned that the data is randomly drawn from the low income subsample of the NLSY. This is probably the reason why, even using some criteria, we cannot match the dropouts in the textbook and the refreshed data.

Further, the temporal variable used in the textbook data is experience, instead of year. The only explanation we found regarding this variable is that it is derived from the years of IDs’ experience since their first job. @SingerJudithD2003Alda stated that this variable reflects the specific moment in each ID labor force history associated with each wage observation. 
Since it is not clearly articulated, we decide to keep the temporal variable in the refreshed data as year. Accordingly, in this comparison, we only compare the same ID but not the same temporal variable.

Regarding the wages, the textbook data express the wages as inflation adjusted with 1990 dollars as constant [@SingerJudithD2003Alda]. However, it is not stated how exactly the wages are adjusted. Note that in the refreshed data we publish in `yowie`, we do not adjust the data with inflation because the time frame covered ranged from 1979-2018. Hence, we allow flexibility to the data user to adjust, not adjust,  or adjust it with different base years. For comparison purposes, we adjust the wages with the average Consumer Price Index For All Urban Consumers (CPI-U) we got from @fed. -->

<!-- Figure \ref{fig:plotting-sw-do} shows the comparison between the textbook and the refreshed data. Even though it has different temporal variables and possibly different CPI indicators used in inflation adjustment, we can still capture some of the patterns in the refreshed data.

From this comparison, we learn that we cannot make an apple to apple comparison and produce the exact same data as the textbook due to the absence of clearly articulated data extraction and derivation process. Hence, the takeaway is that it is imperative to disclose how the data derived from the initial source to make data analysis reproducible. This is why in this paper, we show the practice of reproducible and responsible data cleaning workflow to make sure the longitudinal data from a continuous survey can be refreshed.-->

# Summary {#summary}

This paper has described the stages to take a particular open data set and make it a textbook data set, ready for the classroom or research. In the first stage, we showed the steps performed to get the data from the NLSY79 database. The data format needed converting to tidy format, and this was described. After that, an initial data analysis was conducted to investigate and screen the quality of the data. We found and fixed the anomalous observations in wages using a robust linear regression model. Finally a comparison is made between the original and refreshed data sets.

The data cleaning process is documented and the code has been made available. These provide the opportunity to again refresh the textbook data as updated data is published in the NLSY79 database.
Determining the appropriate robustness weight to threshold the anomalous observations is documented, and a `shiny`  [@shiny] app is provided to assist. 
The current subset is made available in a new R package, called yowie.

Various difficulties were encountered in trying to refresh the data, which include:

- determining which records should be downloaded from the database.
- there are many errors in the data, e.g. hourly wages greater than $30,000 per hour. Further for original data, wages greater than 75 dollars were set to be missing. However, in recent context, this value might be too low to be set as the maximum threshold. 
- there is no explicit variable in the database recording high school dropout, which means we needed to compare date of 12th grade with the individual's age.
- calculating experience in the workforce, would require knowing the time of first job within the first year the individual was recorded.

Ultimately, the refreshed data is reasonably similar to the original, but unsatisfactorily far from it. The last step required would be to inflation-adjust wages, but this is better to do with each wave of new data added, so that it is relative to the last date in the data.

Finally, for the data providers we recommend that a validation system with clear rules is added on data entry, and that alternative output formats, such as a tidy format would help users make better use of the resource. The problem with many of the wages records is that there are implausible values, or confusion on how to record wages for multiple jobs. These values could be validated with simple checks at data entry. Providing an open data resource also is accompanied with the responsibility that the data, especially data that is as valuable as this, is reliable. Users need to be able to trust the data.

<!--this paper implies that data providers should design a database that is able to produce tidy data sets. A data provider should also check for data anomalies before the data publishing or at least provides a set of rules or threshold values. For example, in this case, is the threshold of reasonable wages. This will greatly support the data users to validate and set the same understanding of which data are considered outliers. Moreover, providing validation rules would facilitate any established data validation tool, such as `validate` [@validate] package. In this case, we cannot use this handy validation package due to the absence of validation rules. -->

# Acknowledgements

We would like to thank Aarathy Babu for the insight and discussion during the writing of this paper.

The entire analysis is conducted using `R` [@R] in `rstudio` using these packages: `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tidyr` [@tidyr], `stringr` [@stringr], `purrr` [@purrr], `broom` [@broom], `brolgar` [@brolgar], `patchwork` [@patchwork], `kableExtra` [@kableExtra], `MASS` [@mass], `janitor` [@janitor], `DiagrammeR` [@DiagrammeR], `rsvg` [@rsvg], `webshot` [@webshot], `mgcv` [@mgcv], `tsibble` [@tsibble], and `modelr` [@modelr]. The paper are generated using `knitr` [@knitr], `rmarkdown` [@rmarkdown], and `rticles` [@rticles].

# Supplementary Materials

- **Codes** : R script to reproduce data tidying and cleaning are available in this [page](https://numbats.github.io/yowie/articles/process-data.html).

- **R Package `yowie`**:`yowie` is a data container R package that contains 3 datasets, namely the high school mean hourly wage data, high school dropouts mean hourly wage data, and demographic data of the NLSY79 cohort. This package could be accessed [here](https://github.com/numbats/yowie).

- **shiny app**: An web interactive `shiny` app to customize the weight threshold. This app could be accessed [here](https://github.com/numbats/wages-refresh/tree/main/app).

# References

### 2021-09-28 paper/paper.tex
paper + app folder added, vignette updated

--- 2021-09-28 paper/paper-jsdse/paper-jsdse.tex
@@
-\documentclass[12pt]{article}
-\usepackage{amsmath}
-\usepackage{graphicx,psfrag,epsf}
-\usepackage{enumerate}
-\usepackage{natbib}
-\usepackage{tcolorbox}
-\usepackage{fontawesome}
-\usepackage{booktabs}
-\usepackage{url} % not crucial - just used below for the URL
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\documentclass{article}
 
-%\pdfminorversion=4
-% NOTE: To produce blinded version, replace "0" with "1" below.
-\newcommand{\blind}{0}
+\usepackage{arxiv}
 
-% need this for pandoc %
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
+\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+\usepackage{lipsum}
+\usepackage{graphicx}
+
+\title{The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth}
+
+\author{
+    Dewi Amaliah
+   \\
+    Department of Econometrics and Business Statistics \\
+    Monash University \\
+  Clayton, VIC 3800 \\
+  \texttt{\href{mailto:dama0007@student.monash.edu}{\nolinkurl{dama0007@student.monash.edu}}} \\
+   \And
+    Dianne Cook
+   \\
+    Department of Econometrics and Business Statistics \\
+    Monash University \\
+  Clayton, VIC 3800 \\
+  \texttt{\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}} \\
+   \And
+    Emi Tanaka
+   \\
+    Department of Econometrics and Business Statistics \\
+    Monash University \\
+  Clayton, VIC 3800 \\
+  \texttt{\href{mailto:emi.tanaka@monash.edu}{\nolinkurl{emi.tanaka@monash.edu}}} \\
+   \And
+    Kate Hyde
+   \\
+    Department of Econometrics and Business Statistics \\
+    Monash University \\
+  Clayton, VIC 3800 \\
+  \texttt{} \\
+   \And
+    Nicholas Tierney
+   \\
+     \\
+   \\
+  \texttt{} \\
+  }
+
+% Pandoc citation processing
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
 \newlength{\cslhangindent}
 \setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
 \newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
@@
  {}
-\usepackage{calc} % for \widthof, \maxof
+\usepackage{calc} % for calculating minipage widths
 \newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
-% DON'T change margins - should be 1 inch all around.
-\addtolength{\oddsidemargin}{-.5in}%
-\addtolength{\evensidemargin}{-.5in}%
-\addtolength{\textwidth}{1in}%
-\addtolength{\textheight}{1.3in}%
-\addtolength{\topmargin}{-.8in}%
-
-%-------------------------------------------------------------------
-% R MARKDOWN
-%-------------------------------------------------------------------
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-
-\def\spacingset#1{\renewcommand{\baselinestretch}%
-{#1}\small\normalsize} \spacingset{1}
+\usepackage{tcolorbox}
+\usepackage{fontawesome}
+\usepackage{color, colortbl}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 
 \begin{document}
+\maketitle
 
-%\bibliographystyle{natbib}
+\def\tightlist{}
 
-\def\spacingset#1{\renewcommand{\baselinestretch}%
-{#1}\small\normalsize} \spacingset{1}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\if0\blind
-{
-  \title{\bf The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth}
-  \author{
-            Dewi Amaliah\thanks{Corresponding author, \href{mailto:dlamaleeah@gmail.com}{\nolinkurl{dlamaleeah@gmail.com}}}\\
-      Dept of Econometrics and Business Statistics, Monash University, \\
-            Dianne Cook\\
-      Dept of Econometrics and Business Statistics, Monash University, \\
-            Emi Tanaka\\
-      Dept of Econometrics and Business Statistics, Monash University, \\
-            Kate Hyde\\
-      Dept of Econometrics and Business Statistics, Monash University, \\
-            Nicholas Tierney\\
-      Telethon Kids Institute, Perth, WA, }
-        \maketitle
-} \fi
-
-\if1\blind
-{
-  \bigskip
-  \bigskip
-  \bigskip
-  \begin{center}
-    {\LARGE\bf Title}
-\end{center}
-  \medskip
-} \fi
-
-\bigskip
 \begin{abstract}
@@
 
-\noindent%
-{\it Keywords:}  Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression
-\vfill
-
-\newpage
-\spacingset{1.45} % DON'T change the spacing!
+\keywords{
+    Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression
+  }
 
@@
 However, open data is often referred to as what we might call ``wild data'' because it requires substantial cleaning and tidying to tame it into textbook shape.
-Huebner, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative and essential for the integrity of downstream statistical analyses and model building (Huebner et al.
+Huebner Marianne, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative and essential for the integrity of downstream statistical analyses and model building (M.
+Huebner et al.
 2020).
@@
 There are few research papers that document the data cleaning (Wickham 2014).
-Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way and is only shared among restricted parties (Huebner et al.
+Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way and is only shared among restricted parties (M.
+Huebner et al.
 2020).
@@
 
-Our process of cleaning builds heavily on the tidyverse approach (H.
-Wickham, Averick, et al.
+Our process of cleaning builds heavily on the tidyverse approach (Wickham et al.
 2019).
@@
 The raw data does not comply these rules, thus we re-arrange and wrangle the data into tidy data form, columns corresponding to individual ID, year, job number, wage in dollars and the demographic variables.
-This is done using the \texttt{tidyverse} suite of packages (H.
-Wickham, Averick, et al.
-2019), \texttt{tidyr} (H.
-Wickham 2020b) to pivot the data into long form, \texttt{dplyr} (H.
-Wickham, François, et al.
-2020), and \texttt{stringr} (H.
-Wickham 2019) for creating new variables, and levels of factors by text wrangling.
+This is done using the \texttt{tidyverse} suite of packages (Wickham et al.
+2019), \texttt{tidyr} (Wickham 2020b) to pivot the data into long form, \texttt{dplyr} (Wickham et al.
+2020), and \texttt{stringr} (Wickham 2019) for creating new variables, and levels of factors by text wrangling.
 The long form of the data makes it possible to do these data transformations efficiently, and it is an intermediate step towards the final target data.
@@
 
-According to Huebner, Vach, and Cessie (2016), initial data analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
+According to Huebner Marianne, Vach, and Cessie (2016), initial data analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
 Moreover, Chatfield (1985) argues that the two main objectives of IDA are data description, which is to assess the structure and the quality of the data, and model formulation without any formal statistical inference.
@@
 However, this method is too computationally and memory expensive, especially for a large data set, like the NLSY79 data.
-Thus, the model for each ID is built utilizing the \texttt{nest} and \texttt{map} function from \texttt{tidyr} (H.
-Wickham 2020b) and \texttt{purrr} (Henry and Wickham 2020), respectively.
+Thus, the model for each ID is built utilizing the \texttt{nest} and \texttt{map} function from \texttt{tidyr} (Wickham 2020b) and \texttt{purrr} (Henry and Wickham 2020), respectively.
 
@@
 
-To minimize the risk of mistakenly identifying an outlier as an ``erroneous outlier'', we simulate some thresholds and study how they affect the data.
+To minimize the risk of mistakenly identifying an outlier as an ``erroneous outlier,'' we simulate some thresholds and study how they affect the data.
 We find that 0.12 is the most reasonable value to be the threshold to minimize that drawback's risk because it still captures the sensible spikes in the data.
@@
 It implies that the fluctuation can still be observed in the data after the treatment.
-However, the large spikes, which are considered ``erroneous outliers'', are already eliminated from the data.
+However, the large spikes, which are considered ``erroneous outliers,'' are already eliminated from the data.
 Hence, the model produces a data set with a more reasonable degree of fluctuation.
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/Users/deeamaliah/Documents/summer_project/summer-wages-refresh/paper/figures/flowchart}
+{\centering \includegraphics[width=0.85\linewidth]{/Users/etan0038/Dropbox/repo/summer-wages-refresh/paper/figures/flowchart}
 
@@
 There are two aspects of the original data that make a direct comparison difficult.
-The time variable provided is ``experience in the workforce''.
-It is not clear how this is calculated.
+The time variable provided is ``experience in the workforce.'' It is not clear how this is calculated.
 One would expect that there is a record of the day the individual first started a job, and this is used to adjust the year of collection.
@@
 
-The entire analysis is conducted using \texttt{R} (R Core Team 2020) in \texttt{rstudio} using these packages: \texttt{tidyverse} (H.
-Wickham, Averick, et al.
-2019), \texttt{ggplot2} (Wickham 2016), \texttt{dplyr} (H.
-Wickham, François, et al.
-2020), \texttt{readr} (H.
-Wickham and Hester 2020), \texttt{tidyr} (H.
-Wickham 2020b), \texttt{stringr} (H.
-Wickham 2019), \texttt{purrr} (Henry and Wickham 2020), \texttt{broom} (Robinson, Hayes, and Couch 2021), \texttt{brolgar} (Tierney, Cook, and Prvan 2020), \texttt{patchwork} (Pedersen 2020), \texttt{kableExtra} (Zhu 2019), \texttt{MASS} (Venables and Ripley 2002), \texttt{janitor} (Firke 2020), \texttt{DiagrammeR} (Iannone 2020), \texttt{rsvg} (Ooms 2020), \texttt{webshot} (Chang 2019), \texttt{mgcv} (Wood 2003), \texttt{tsibble} (Wang, Cook, and Hyndman 2020), and \texttt{modelr} (H.
-Wickham 2020a).
+The entire analysis is conducted using \texttt{R} (R Core Team 2020) in \texttt{rstudio} using these packages: \texttt{tidyverse} (Wickham et al.
+2019), \texttt{ggplot2} (Wickham 2016), \texttt{dplyr} (Wickham et al.
+2020), \texttt{readr} (Wickham and Hester 2020), \texttt{tidyr} (Wickham 2020b), \texttt{stringr} (Wickham 2019), \texttt{purrr} (Henry and Wickham 2020), \texttt{broom} (Robinson, Hayes, and Couch 2021), \texttt{brolgar} (Tierney, Cook, and Prvan 2020), \texttt{patchwork} (Pedersen 2020), \texttt{kableExtra} (Zhu 2019), \texttt{MASS} (Venables and Ripley 2002), \texttt{janitor} (Firke 2020), \texttt{DiagrammeR} (Iannone 2020), \texttt{rsvg} (Ooms 2020), \texttt{webshot} (Chang 2019), \texttt{mgcv} (Wood 2003), \texttt{tsibble} (Wang, Cook, and Hyndman 2020), and \texttt{modelr} (Wickham 2020a).
 The paper are generated using \texttt{knitr} (Xie 2014), \texttt{rmarkdown} (Xie, Dervieux, and Riederer 2020), and \texttt{rticles} (Allaire et al.
@@
 \hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
 \leavevmode\hypertarget{ref-rticles}{}%
@@
 2021.
-\emph{Rticles: Article Formats for R Markdown}.
+\emph{Rticles: Article Formats for r Markdown}.
 \url{https://CRAN.R-project.org/package=rticles}.
@@
 \leavevmode\hypertarget{ref-shiny}{}%
-Chang, Winston, Joe Cheng, JJ Allaire, Yihui Xie, and Jonathan McPherson. 2020. \emph{Shiny: Web Application Framework for R}. \url{https://CRAN.R-project.org/package=shiny}.
+Chang, Winston, Joe Cheng, JJ Allaire, Yihui Xie, and Jonathan McPherson. 2020. \emph{Shiny: Web Application Framework for r}. \url{https://CRAN.R-project.org/package=shiny}.
 
 \leavevmode\hypertarget{ref-Chatfield1985TIEo}{}%
-Chatfield, C. 1985. ``The Initial Examination of Data.'' \emph{Journal of the Royal Statistical Society. Series A. General} 148 (3): 214--53.
+Chatfield, C. 1985. {``The Initial Examination of Data.''} \emph{Journal of the Royal Statistical Society. Series A. General} 148 (3): 214--53.
 
@@
 2017.
-``Using the National Longitudinal Surveys of Youth (Nlsy) to Conduct Life Course Analyses.'' In \emph{Handbook of Life Course Health Development}, edited by Richard M.
+{``Using the National Longitudinal Surveys of Youth (NLSY) to Conduct Life Course Analyses.''} In \emph{Handbook of Life Course Health Development}, edited by Richard M.
 Lerner Neal Halfon Christoper B.
@@
 Cham: Springer.
-\url{https://doi.org/https://doi.org/10.1007/978-3-319-47143-3_23}.
+https://doi.org/\url{https://doi.org/10.1007/978-3-319-47143-3_23}.
 
@@
 \leavevmode\hypertarget{ref-racismnotrace}{}%
-Fullilove, M. T. 1998. ``Comment: Abandoning "Race" as a Variable in Public Health Research--an Idea Whose Time Has Come.'' \emph{American Journal of Public Health} 88 (9): 1297--8.
+Fullilove, M. T. 1998. {``Comment: Abandoning "Race" as a Variable in Public Health Research--an Idea Whose Time Has Come.''} \emph{American Journal of Public Health} 88 (9): 1297--98.
 
@@
 2016.
-``A Systematic Approach to Initial Data Analysis Is Good Research Practice.'' \emph{The Journal of Thoracic and Cardiovascular Surgery} 151 (1): 25--27.
+{``A Systematic Approach to Initial Data Analysis Is Good Research Practice.''} \emph{The Journal of Thoracic and Cardiovascular Surgery} 151 (1): 25--27.
 
@@
 2020.
-``Hidden Analyses: A Review of Reporting Practice and Recommendations for More Transparent Reporting of Initial Data Analyses.'' \emph{BMC Medical Research Methodology} 20 (1): 61--61.
+{``Hidden Analyses: A Review of Reporting Practice and Recommendations for More Transparent Reporting of Initial Data Analyses.''} \emph{BMC Medical Research Methodology} 20 (1): 61--61.
 
@@
 \leavevmode\hypertarget{ref-KollerManuel2016rARP}{}%
-Koller, Manuel. 2016. ``Robustlmm: An R Package for Robust Estimation of Linear Mixed-Effects Models.'' \emph{Journal of Statistical Software} 75 (6): 1--24.
+Koller, Manuel. 2016. {``Robustlmm: An r Package for Robust Estimation of Linear Mixed-Effects Models.''} \emph{Journal of Statistical Software} 75 (6): 1--24.
 
 \leavevmode\hypertarget{ref-rsvg}{}%
-Ooms, Jeroen. 2020. \emph{Rsvg: Render Svg Images into Pdf, Png, Postscript, or Bitmap Arrays}. \url{https://CRAN.R-project.org/package=rsvg}.
+Ooms, Jeroen. 2020. \emph{Rsvg: Render SVG Images into PDF, PNG, PostScript, or Bitmap Arrays}. \url{https://CRAN.R-project.org/package=rsvg}.
 
 \leavevmode\hypertarget{ref-opendata}{}%
-Open Knowledge Foundation. 2021. ``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.'' 2021. \url{http://opendefinition.org/od/2.1/en/}.
+Open Knowledge Foundation. 2021. {``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.''} 2021. \url{http://opendefinition.org/od/2.1/en/}.
 
@@
 2001.
-``Data Watch: The National Longitudinal Surveys.'' \emph{The Journal of Economic Perspectives} 15 (2): 239--53.
+{``Data Watch: The National Longitudinal Surveys.''} \emph{The Journal of Economic Perspectives} 15 (2): 239--53.
 
@@
 \leavevmode\hypertarget{ref-nlsy79}{}%
-The U.S. Bureau of Labor Statistics. n.d. ``National Longitudinal Survey of Youth 1979.'' Available at \url{https://www.nlsinfo.org/content/cohorts/nlsy79} (2021/25/02).
+The U.S. Bureau of Labor Statistics. n.d. {``National Longitudinal Survey of Youth 1979.''} Available at \url{https://www.nlsinfo.org/content/cohorts/nlsy79} (2021/25/02).
 
 \leavevmode\hypertarget{ref-brolgar}{}%
-Tierney, Nicholas, Di Cook, and Tania Prvan. 2020. \emph{Brolgar: BRowse over Longitudinal Data Graphically and Analytically in R}. \url{https://github.com/njtierney/brolgar}.
+Tierney, Nicholas, Di Cook, and Tania Prvan. 2020. \emph{Brolgar: BRowse over Longitudinal Data Graphically and Analytically in r}. \url{https://github.com/njtierney/brolgar}.
 
@@
 \leavevmode\hypertarget{ref-rlm}{}%
-UCLA: Statistical Consulting Group. 2021. ``Robust Regression \textbar{} R Data Analysis Examples.'' February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.
+UCLA: Statistical Consulting Group. 2021. {``Robust Regression \textbar{} r Data Analysis Examples.''} February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.
 
 \leavevmode\hypertarget{ref-LooMarkvander2018Sdcw}{}%
-van der Loo, Mark, and Edwin de Jonge. 2018. \emph{Statistical Data Cleaning with Applications in R}.
+van der Loo, Mark, and Edwin de Jonge. 2018. \emph{Statistical Data Cleaning with Applications in r}.
 
 \leavevmode\hypertarget{ref-mass}{}%
-Venables, W. N., and B. D. Ripley. 2002. \emph{Modern Applied Statistics with S}. Fourth. New York: Springer. \url{http://www.stats.ox.ac.uk/pub/MASS4}.
+Venables, W. N., and B. D. Ripley. 2002. \emph{Modern Applied Statistics with s}. Fourth. New York: Springer. \url{http://www.stats.ox.ac.uk/pub/MASS4}.
 
@@
 2020.
-``A New Tidy Data Structure to Support Exploration and Modeling of Temporal Data.'' \emph{Journal of Computational and Graphical Statistics} 29 (3): 466--78.
+{``A New Tidy Data Structure to Support Exploration and Modeling of Temporal Data.''} \emph{Journal of Computational and Graphical Statistics} 29 (3): 466--78.
 \url{https://doi.org/10.1080/10618600.2019.1695624}.
@@
 \leavevmode\hypertarget{ref-plyr}{}%
-Wickham, Hadley. 2011. ``The Split-Apply-Combine Strategy for Data Analysis.'' \emph{Journal of Statistical Software, Articles} 40 (1): 1--29. \url{https://doi.org/10.18637/jss.v040.i01}.
+Wickham, Hadley. 2011. {``The Split-Apply-Combine Strategy for Data Analysis.''} \emph{Journal of Statistical Software, Articles} 40 (1): 1--29. \url{https://doi.org/10.18637/jss.v040.i01}.
 
 \leavevmode\hypertarget{ref-WickhamHadley2014TD}{}%
----------. 2014. ``Tidy Data.'' \emph{Journal of Statistical Software} 59 (10): 1--23.
+---------. 2014. {``Tidy Data.''} \emph{Journal of Statistical Software} 59 (10): 1--23.
 
@@
 2019.
-``Welcome to the tidyverse.'' \emph{Journal of Open Source Software} 4 (43): 1686.
+{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software} 4 (43): 1686.
 \url{https://doi.org/10.21105/joss.01686}.
@@
 \leavevmode\hypertarget{ref-mgcv}{}%
-Wood, S. N. 2003. ``Thin-Plate Regression Splines.'' \emph{Journal of the Royal Statistical Society (B)} 65 (1): 95--114.
+Wood, S. N. 2003. {``Thin-Plate Regression Splines.''} \emph{Journal of the Royal Statistical Society (B)} 65 (1): 95--114.
 
@@
 2014.
-``Knitr: A Comprehensive Tool for Reproducible Research in R.'' In \emph{Implementing Reproducible Computational Research}, edited by Victoria Stodden, Friedrich Leisch, and Roger D.
+{``Knitr: A Comprehensive Tool for Reproducible Research in {R}.''} In \emph{Implementing Reproducible Computational Research}, edited by Victoria Stodden, Friedrich Leisch, and Roger D.
 Peng.
@@
 \leavevmode\hypertarget{ref-kableExtra}{}%
-Zhu, Hao. 2019. \emph{KableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.
+Zhu, Hao. 2019. \emph{kableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.
 
-\bibliographystyle{Chicago}
+\end{CSLReferences}
+
+\bibliographystyle{unsrt}
+\bibliography{references.bib}
 

### 2021-09-28 data-raw/data-preprocessing.R

@@ -1,26 +1,15 @@
 
 
-#' The data cleaning step is conducted using the suite of packages in `tidyverse`.
-#+ load-pkgs
+## ---- load-pkgs
 library(tidyverse)
 select <- dplyr::select
 
-#' # Reading the data {#read-data}
-#+ raw-data, eval=FALSE, include=FALSE
+## ---- read-data
 source(here::here("data-raw/NLSY79/NLSY79.R"))
 
-#' <details class="source">
-#' <summary>Click here for the **source file to read data** (note this is quite long)</summary>
-#' The code below is provided by the NLSY79 database to do the
-#' reading and initial cleaning of data.
-#+ raw-data-rmd, code=readLines(here::here("data-raw/NLSY79/NLSY79.R"))
-#' </details>
 
-#'
-#' The above source code creates a data set `new_data_qnames`.
-#' As shown below, the column names contain information on the job number
-#' (HRP1 = job 1, HRP2 = job 2, ..., HRP5 = job 5) and the year information.
-#+ untidy-data, echo=-1
+
+## ---- untidy-data
 options(width=70)
 new_data_qnames %>%
   select(CASEID_1979,
@@ -29,15 +18,8 @@ new_data_qnames %>%
          everything()) %>%
   str(list.len = 10)
 
-#' # Demographic data pre-processing
-#' ## Tidy the date of birth data
-#' The month and year of birth is recorded in 1979 and 1981 for each individual.
-#' The records in 1981 are missing for some individuals so we take the month and year of
-#' birth from 1979 records.
-#'
-#' Where the record is present for both 1979 and 1981, we check that the
-#' record matches.
-#+ dob-tidy
+
+## ---- dob-tidy
 dob_tidy <- new_data_qnames %>%
   select(id = CASEID_1979,
          year_1979 = `Q1-3_A~Y_1979`,
@@ -70,23 +52,16 @@ dob_tidy <- new_data_qnames %>%
          dob_conflict)
 
 has_dob_conflict <- any(dob_tidy$dob_conflict, na.rm = TRUE)
-#+ dob-conflict-no, echo=FALSE, eval=!has_dob_conflict
-cat("All birth month and year recorded in 1979 and 1981 match.")
-#+ dob-conflict-yes, echo=has_dob_conflict, eval=has_dob_conflict
[…]

### 2021-09-28 paper/paper-jsdse/paper-jsdse.tex
fixed to be able to compile

@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/Users/deeamaliah/Documents/summer_project/summer-wages-refresh/paper/figures/flowchart}
+{\centering \includegraphics[width=0.85\linewidth]{/Users/etan0038/yowie/paper/figures/flowchart}
 
@@
 \# of individuals means the number of respondents included in each stage, while \# of observations means the number of rows in the data.
-The color represented the stage of data cleaning in statistical value chain (van der Loo \& de Jonge, 2018).
+The color represents the stage of data cleaning in statistical value chain (van der Loo \& de Jonge, 2018).
 Pink, blue, and green represents the raw, input, and valid data, respectively.}\label{fig:flow-chart}
@@
 \item
-  there are many errors in the data, e.g.~hourly wages greater than \$30,000 per hour, and there was no information on how these were handled in the original data.
+  there are many errors in the data, e.g.~hourly wages greater than \$30,000 per hour.
+Further for original data, wages greater than 75 dollars were set to be missing.
+However, in recent context, this value might be too low to be set as the maximum threshold.
 \item
@@
 \item
-  \textbf{Codes} : R script to reproduce data tidying and cleaning are available in this \href{https://github.com/numbats/yowie/blob/master/data-raw/data_preprocessing.R}{Github Repository}.
+  \textbf{Codes} : R script to reproduce data tidying and cleaning are available in this \href{https://numbats.github.io/yowie/articles/process-data.html}{page}.
 \item
@@
 \item
-  \textbf{shiny app}: An web interactive \texttt{shiny} app to run a simulation to customize the weight threshold.
-This app could be accessed \href{https://github.com/numbats/wages-refresh/tree/main/app}{here}.
+  \textbf{shiny app}: An web interactive \texttt{shiny} app to customize the weight threshold. This app could be accessed \href{https://github.com/numbats/wages-refresh/tree/main/app}{here}.
 \end{itemize}
@@
 \leavevmode\hypertarget{ref-kableExtra}{}%
-Zhu, Hao. 2019. \emph{KableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.
+Zhu, Hao. 2019. \emph{kableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.
+
+\end{CSLReferences}
 

### 2021-09-28 paper/paper.tex
fixed to be able to compile

@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/Users/etan0038/Dropbox/repo/summer-wages-refresh/paper/figures/flowchart}
+{\centering \includegraphics[width=0.85\linewidth]{/Users/etan0038/yowie/paper/figures/flowchart}
 

### 2021-09-28 data-raw/data-preprocessing.R

@@ -4,10 +4,8 @@
 library(tidyverse)
 select <- dplyr::select
 
-## ---- read-data
+## ---- raw-data
 source(here::here("data-raw/NLSY79/NLSY79.R"))
-
-
 
 ## ---- untidy-data
 options(width=70)

### 2021-09-28 data-raw/ida.R

@@ -1,3 +1,10 @@
+## ---- load-pkgs2
+library(kableExtra)
+library(janitor)
+library(tsibble)
+library(brolgar)
+library(patchwork)
+
 ## ---- age-table
 age_table <- full_demographics %>%
   group_by(age_1979) %>%

### 2021-09-28 data-raw/data-valid.R
data valid vig fixed

@@ -1,4 +1,5 @@
 
+## ---- load-pkgs-valid
 library(tidyverse)
 library(MASS)
 select <- dplyr::select
@@ -14,35 +15,18 @@ by_id <- wages_before %>%
 
 # build a robust linear model
 id_rlm <- by_id %>%
-  mutate(model = map(.x = data,
-                     .f = function(x){
-                       rlm(mean_hourly_wage ~ year, data = x)
-                     }))
-# extract the property of the regression model
-id_aug <- id_rlm %>%
-  mutate(augmented = map(model, broom::augment)) %>%
-  unnest(augmented)
+  mutate(model = map(data, ~rlm(mean_hourly_wage ~ year, data = .x)))
 
-# extract the weight of each observation
-id_w <- id_rlm %>%
-  mutate(w = map(.x = model,
-                 .f = function(x){
-                   x$w
-                 })) %>%
-  unnest(w) %>%
-  select(w)
+# extract the properties of regression model and weight of each observation
 
-# bind the property of each observation with their weight
-id_aug_w <- cbind(id_aug, id_w) %>%
-  select(`id...1`,
-         year,
-         mean_hourly_wage,
-         .fitted,
-         .resid,
-         .hat,
-         .sigma,
-         w) %>%
-  rename(id = `id...1`)
+id_aug_w <- id_rlm %>%
+  mutate(.fitted = map(model, fitted),
+         .resid = map(model, resid),
+         .hat = map(model, hatvalues),
+         .sigma = map(model, sigma),
+         w = map(model, ~.x$w)) %>%
+  select(-model) %>%
+  unnest(data:w)
 
 # if the weight < 0.12, the mean_hourly_wage is replaced by the model's fitted/predicted value.
 # and add the flag whether the observation is predicted value or not.
@@ -60,6 +44,6 @@ wages_rlm_dat <- id_aug_w %>%
 
 wages_after <- left_join(wages_before, wages_rlm_dat, by = c("id", "year"))
 
+## ---- save-wages-after
 # save it to rds file so it would faster the knitting process
-#+ save-wages-after, eval = FALSE, echo = FALSE
 saveRDS(wages_after, here::here("paper/results/wages_after.rds"))

### 2021-09-28 paper/paper-jsdse/paper-jsdse.tex
tidying demographic data description updated

@@
 Thus the target data set is to collect the same variables for the extended time frame of 1994 through to 2018, the most recent year reported.
-.
 
@@
 \item[\faCloudDownload] To download all 686 variables selected, we then navigate to the tab "Save / Download" then select the tab "Advanced Download".
-We select the R Source code, Codebook of selected variables, Short Description File and Comma-delimited datafile of selected variables with Reference Number as column headers.
+We select the R Source code and Comma-delimited datafile of selected variables with Reference Number as column headers.
 We name the filename "NLSY79" and press the download button.
+There are also options to get control or dictionary files for SAS, SPSS and STATA.
 \end{itemize}
@@
 \item
-  \texttt{NLSY79.csv}: csv format of the response data.
+  \texttt{NLSY79.csv}: csv format of the response data,
 \item
-  \texttt{NLSY79.dat}: dat format of the response data.
+  \texttt{NLSY79.dat}: dat format of the response data,
 \item
-  \texttt{NLSY79.NLSY79}: Tagset of variables that can be uploaded to the website to recreate the data set.
+  \texttt{NLSY79.NLSY79}: Tagset of variables that can be uploaded to the website to recreate the data set, and
 \item
   \texttt{NLSY79.R}: R script for reading the data into R and converting the variables' names and label into something more sensible.
-We utilize this code to do an initial data tidying which produces two data set, \texttt{categories\_qnames} (where the observations are stored in categorical/interval values) and \texttt{new\_data\_qnames} (the observations are stored in integer form).
 \end{itemize}
 
-\hypertarget{tidywages}{%
-\subsubsection{Tidying the wages data}\label{tidywages}}
+We only alter the file path in \texttt{NLSY79.R} and run the script without any other alteration.
+This results in an initial processing of the raw data into two data set, \texttt{categories\_qnames} (where the observations are stored in categorical/interval values) and \texttt{new\_data\_qnames} (the observations are stored in integer form).
 
@@
 The long form of the data makes it possible to do these data transformations efficiently, and it is an intermediate step towards the final target data.
+The code for tidying the data are demonstrated at \url{https://numbats.github.io/yowie/articles/raw-to-input-data.html} but also described in the subsequent subsections.
 
@@
 
-\texttt{Age\ in\ 1979}, \texttt{gender}, \texttt{race}, \texttt{highest\ grade\ completed}, and the \texttt{year\ when\ the\ highest\ grade\ completed} are the variables that we want to use in the cleaned data set.
+In our final target data, we wish to include the demographic variables with variable names specified in brackets: gender (\texttt{gender}), race (\texttt{race}), age (\texttt{age}), highest grade completed (\texttt{hgc}) and the corresponding year this grade was completed (\texttt{yr\_hgc}).
 
-Age in 1979 are derived from year and month of birth variables in the raw data.
-The variables have two versions, which are the 1979 version and the 1981 version.
-We only used the 1979 data and did a consistency check of those years and flag the inconsistent observations.
-
-For \texttt{gender} and \texttt{race}, we only rename the column names.
+For gender and race, we only rename the column names.
 It is worth noting that using these two variable needs special attention.
@@
 
-The next step is tidying the \texttt{highest\ grade\ completed} variable.
+The \texttt{new\_data\_qnames} contains the variables \texttt{Q1-3\_A\textasciitilde{}Y\_1979} and \texttt{Q1-3\_A\textasciitilde{}Y\_1981} which records two versions of the birth year of the respondent; this is also the case for the record of birth month (\texttt{Q1-3\_A\textasciitilde{}M\_1979} and \texttt{Q1-3\_A\textasciitilde{}M\_1981}).
+The record contains two versions of birth year and birth month as the survey recorded this in 1979 and 1981.
+We checked for consistency for the two versions and found no discrepancy where the responses were recorded in both 1979 and 1981.
+The age was then calculated using the birth year.
+
+The next step is tidying the highest grade completed (\texttt{hgc}) variable.
 This variable came with several version in each year.
@@
 
-Further, \texttt{highest\ grade\ completed} is measured and could be updated in each period of the survey.
+Furthermore, \texttt{hgc} is measured and could be updated in each period of the survey.
 Hence, we only used the highest grade completed ever and derived the year when it is completed.
-The code for tidying this variable is also available in the supplementary materials.
+The code for deriving this variable is also available at \url{https://numbats.github.io/yowie/articles/raw-to-input-data.html}.
 

### 2021-09-28 paper/paper.Rmd
fixed to be able to compile
tidying demographic data description updated

@@
 
-```{r load-pkgs}
+```{r ref.label=c("load-pkgs", "load-pkgs2"), cache = FALSE}
 ```
@@
 
-- `NLSY79.csv`: csv format of the response data.
-- `NLSY79.dat`: dat format of the response data.
-- `NLSY79.NLSY79`: Tagset of variables that can be uploaded to the website to recreate the data set.
+- `NLSY79.csv`: csv format of the response data,
+- `NLSY79.dat`: dat format of the response data,
+- `NLSY79.NLSY79`: Tagset of variables that can be uploaded to the website to recreate the data set, and
 - `NLSY79.R`: R script for reading the data into R and converting the variables' names and label into something more sensible.
-We utilize this code to do an initial data tidying which produces two data set,  `categories_qnames` (where the observations are stored in categorical/interval values) and `new_data_qnames` (the observations are stored in integer form).
+
+We only alter the file path in `NLSY79.R` and run the script without any other alteration.
+This results in an initial processing of the raw data into two data set,  `categories_qnames` (where the observations are stored in categorical/interval values) and `new_data_qnames` (the observations are stored in integer form).
 
@@
 ```
-
-### Tidying the wages data {#tidywages}
 
@@
 
-```{r untidy-data, fig.cap="The structure of the NLSY79 raw data on wages for up to five jobs in any year."}
+```{r untidy-data}
 ```
@@
 
-`Age in 1979`, `gender`, `race`, `highest grade completed`, and the `year when the highest grade completed` are the variables that we want to use in the cleaned data set.
+In our final target data, we wish to include the demographic variables with variable names specified in brackets: gender (`gender`), race (`race`), age (`age`), highest grade completed (`hgc`) and the corresponding year this grade was completed (`yr_hgc`).
 
-Age in 1979 are derived from year and month of birth variables in the raw data.
-The variables have two versions, which are the 1979 version and the 1981 version.
-We only used the 1979 data and did a consistency check of those years and flag the inconsistent observations.
-
-```{r dob-tidy}
-```
-
-For `gender` and `race`, we only rename the column names.
+For gender and race, we only rename the column names.
 It is worth noting that using these two variable needs special attention.
@@
 
+The `new_data_qnames` contains the variables `Q1-3_A~Y_1979` and `Q1-3_A~Y_1981` which records two versions of the birth year of the respondent; this is also the case for the record of birth month (`Q1-3_A~M_1979` and `Q1-3_A~M_1981`).
+The record contains two versions of birth year and birth month as the survey recorded this in 1979 and 1981.
+We checked for consistency for the two versions and found no discrepancy where the responses were recorded in both 1979 and 1981.
+The age was then calculated using the birth year.
+
+```{r dob-tidy}
+```
 ```{r demog-tidy}
 ```
+```{r demog-ed}
+```
 
-The next step is tidying the `highest grade completed` variable.
+The next step is tidying the highest grade completed (`hgc`) variable.
 This variable came with several version in each year.
@@
 
-Further, `highest grade completed` is measured and could be updated in each period of the survey.
+Furthermore, `hgc` is measured and could be updated in each period of the survey.
 Hence, we only used the highest grade completed ever and derived the year when it is completed.
-The code for tidying this variable is also available in the supplementary materials.
+The code for deriving this variable is also available at \url{https://numbats.github.io/yowie/articles/raw-to-input-data.html}.
 

### 2021-09-28 data-raw/ida.R

@@ -91,7 +91,7 @@ spag <- wages_before %>%
   geom_line(alpha = 0.5) +
   scale_x_continuous("Year",
                      breaks = seq(1980, 2020, 10),
-                     labels = c("80", "90", "00", "10", "20"),
+                     #labels = c("80", "90", "00", "10", "20"),
                      minor_breaks = seq(1980, 2020, 5)) +
   ggtitle("A") +
   theme_bw() +
@@ -137,7 +137,7 @@ plot_high <- ggplot(filter(wages_high2, id == 39)) +
   annotate("text", x = 2010, y = 1000, label = "id: 39") +
   scale_x_continuous("Year",
                      breaks = seq(1980, 2020, 10),
-                     labels = c("80", "90", "00", "10", "20"),
+                     #labels = c("80", "90", "00", "10", "20"),
                      minor_breaks = seq(1980, 2020, 5)) +
   ylab("Hourly wage ($)") +
   theme_bw() +

### 2021-09-29 paper/paper.Rmd
Modify the url in the paper.
Modify keywords according to jsdse guidelines.

@@
 always_allow_html: true
-keywords: "Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression"
+keywords: "Data cleaning; Data tidying; Reproducible workflow; Longitudinal data; NLSY79; Initial data analysis;"
 ---
@@
 
-- **shiny app**: An web interactive `shiny` app to customize the weight threshold. This app could be accessed [here](https://github.com/numbats/wages-refresh/tree/main/app).
+- **shiny app**: An web interactive `shiny` app to customize the weight threshold. This app could be accessed [here](https://ebsmonash.shinyapps.io/yowie_app/).
 

### 2021-09-29 paper/paper.Rmd
hgc_i explanation added in paper, the intermediate data.frame objects printed in raw2input vignette

--- 2021-09-28 paper/paper.Rmd
@@
 
-In our final target data, we wish to include the demographic variables with variable names specified in brackets: gender (`gender`), race (`race`), age (`age`), highest grade completed (`hgc`) and the corresponding year this grade was completed (`yr_hgc`).
+In our final target data, we wish to include the demographic variables with variable names specified in brackets: gender (`gender`), race (`race`), age (`age_1979`), highest grade completed (`hgc`), highest grade completed in terms of years, e.g.
+9th grade = 9, 3rd year college = 15, (`hgc_i`) and the corresponding year this grade was completed (`yr_hgc`).
 
@@
 We choose the revised May data because it seemed to have less missing and presumably has been checked.
-However, there is no revised May data for 2012, 2014, 2016, and 2018.
-Thus, we just use the ordinary May data for these years.
+However, there is no revised May data for 2012, 2014, 2016, and 2018, thus, we use the ordinary May data for these years.
 The code for tidying this variable is provided in the supplementary materials.

### 2021-09-29 data-raw/data-preprocessing.R

@@ -80,10 +80,10 @@ demog_education <- new_data_qnames %>%
 # Get the highest year completed
 highest_year <- demog_education %>%
   group_by(id) %>%
-  mutate(hgc_i = max(grade)) %>%
-  filter(hgc_i == grade) %>%
-  filter(year == first(year)) %>%
-  rename(yr_hgc = year) %>%
+  filter(grade == max(grade),
+         year == min(year)) %>%
+  rename(yr_hgc = year,
+         hgc_i = grade) %>%
   select(id, yr_hgc, hgc_i) %>%
   ungroup() %>%
   mutate(hgc = case_when(hgc_i == 0 ~ "NONE",
@@ -189,9 +189,9 @@ mean_hourly_wage <- eligible_wages %>%
   mutate(year = as.numeric(year)) %>%
   ungroup() %>%
   rename(mean_hourly_wage = wages) %>%
-  mutate(is_wm = ifelse(flag1 == 1, TRUE,
-                        FALSE)) %>%
+  mutate(is_wm = flag1 == 1) %>%
   select(-flag1)
+
 head(mean_hourly_wage, n = 10)
 
 ## ---- wages-demog-hs

### 2021-09-29 app/app.Rmd

@@
 ```{r data}
-wages_hs <- readRDS(here::here("app/wages_hs.rds")) %>%
+wages_hs <- readRDS(here::here("wages_hs.rds")) %>%
     dplyr::select(-`X1.x`, -`X1.y`)
@@
 ```{r}
-sliderInput("th", "Threshold:",
+sliderInput("th", "Slide to set the threshold",
             min = 0.01,
@@
 
-Row {data-height=400}
+<style>
+div.blue { background-color:#FADBD8; border-radius: 10px; padding: 10px;}
+</style>
+<div class = "blue">
+
+```{r}
+p(icon("info-circle", lib = "font-awesome"), strong("More"))
+```
+
+Have any issue or question? 
+Please put it [here](https://github.com/numbats/yowie/issues)
+
+</div>
+
+Row {data-height=300}
 -------------------------------------
@@
 ```{r}
+p(icon("info-circle", lib = "font-awesome"), "The imputed plot shows how the data changes after the threshold is customized.
+You can then compare it with the original plot.
+The following is the summary statistics of the imputed data.")
+```
+
+```{r}
 max_ori <- max(wages_hs$mean_hourly_wage)
@@
 ```{r}
-p(icon("info-circle", lib = "font-awesome"), "Summary statistics of wages in imputed data")
 renderText({paste0("Minimum = ", round(min(pred()$wages_rlm), 2), " (Original = 0.01)")})

### 2021-09-29 paper/paper.Rmd
shiny app part expanded in paper

@@
 
-- **shiny app**: An web interactive `shiny` app to customize the weight threshold. This app could be accessed [here](https://ebsmonash.shinyapps.io/yowie_app/).
+- **shiny app**: An interactive `shiny` web app to visualise the effect of selecting different weight threshold for substituting the wages data to its predicted value from a fit of the robust linear regression model.
+This app could be accessed [here](https://ebsmonash.shinyapps.io/yowie_app/) with the source code provided [here](https://github.com/numbats/yowie/tree/master/app).
 

### 2021-09-29 paper/paper.Rmd
remove packages not used in acknowledgement

@@
 
-The entire analysis is conducted using `R` [@R] in `rstudio` using these packages: `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tidyr` [@tidyr], `stringr` [@stringr], `purrr` [@purrr], `broom` [@broom], `brolgar` [@brolgar], `patchwork` [@patchwork], `kableExtra` [@kableExtra], `MASS` [@mass], `janitor` [@janitor], `DiagrammeR` [@DiagrammeR], `rsvg` [@rsvg], `webshot` [@webshot], `mgcv` [@mgcv], `tsibble` [@tsibble], and `modelr` [@modelr].
-The paper are generated using `knitr` [@knitr], `rmarkdown` [@rmarkdown], and `rticles` [@rticles].
+The entire analysis is conducted using `R` [@R] in RStudio IDE using these packages: `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tidyr` [@tidyr], `stringr` [@stringr], `purrr` [@purrr], `brolgar` [@brolgar], `patchwork` [@patchwork], `kableExtra` [@kableExtra], `MASS` [@mass], `janitor` [@janitor], and `tsibble` [@tsibble].
+The paper are generated using `knitr` [@knitr] and `rmarkdown` [@rmarkdown].
 

### 2021-09-29 paper/paper.Rmd
figures in paper updated so uses tag instead of title

@@
 ```{r feature-plot, fig.cap = "Summary plots to check the cleaned data reveal more cleaning is necessary: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual (id=39) with an unusual wage relative to their years of data (C).
-Some values of hourly wages are unbelievable, and some individuals have extemely unusual wages in some years.", fig.width=10, fig.height=4, out.width="100%"}
+Some values of hourly wages are unbelievable, and some individuals have extemely unusual wages in some years.", fig.width=10, fig.height=4, out.width="100%", cache = FALSE}
 ```
@@
 
-### Replacing extreme values
+### Replacing extreme values {#censor}
 
@@
 
-Further, \ref{fig:fixed-feature-plot} A shows that after eliminating the extreme values, the highest value has decreased to be around $350.
+Further, Figure \ref{fig:fixed-feature-plot} A shows that after eliminating the extreme values, the highest value has decreased to be around $350.
 The spikes are still observed but not as extreme as the original data set.
@@
 We still can see some extreme values in maximum wages, but consider it as a natural variability of the data.
-In \ref{fig:fixed-feature-plot} C, we can see that after imputing the extreme value in ID=39, we can see how the the wages change over the years more clearly.
+In Figure \ref{fig:fixed-feature-plot} C, we can see that after imputing the extreme value in ID=39, we can see how the the wages change over the years more clearly.
 
 ```{r fixed-feature-plot, fig.cap = "Re-make of the summary plots of the fully processed data suggest that it is now in a reasonable state: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual with an unusual wage relative to their years of data (C).
-", fig.width=10, fig.height=4, out.width="100%"}
+", fig.width=10, fig.height=4, out.width="100%", cache = FALSE}
 ```
@@
 ```{r plotting-sw-do, fig.cap = "Comparison of original textbook example (A) with refreshed data (B).
-The original data was inflation-adjusted to 1990 prices and the individual's time of collection was converted to a length of experience in the workforce, which make it difficult to precisely compare the two sets.", fig.width = 9, fig.height = 5, out.width = "100%"}
+The original data was inflation-adjusted to 1990 prices and the individual's time of collection was converted to a length of experience in the workforce, which make it difficult to precisely compare the two sets.", fig.width = 9, fig.height = 5, out.width = "100%", cache = FALSE}
 ```
@@
 Secondly, wages were inflation-adjusted to 1990 prices, which is not done for the refreshed data.
+
+The treatment of unlikely wages differ in the refreshed data.
+In the original data by @SingerJudithD2003Alda, wages greater than $75 are set to be missing.
+However, in recent context, this value is too low to be set as the maximum threshold.
+We opted to use the weights from a robust linear regression to determine what should be set as missing value as described in Section \@ref(censor).
 
@@
 Determining the appropriate robustness weight to threshold the anomalous observations is documented, and a `shiny`  [@shiny] app is provided to assist. 
-The current subset is made available in a new R package, called yowie.
+The current subset is made available in a new R package, called `yowie`.
 
@@
 - determining which records should be downloaded from the database.
-- there are many errors in the data, e.g.
-hourly wages greater than $30,000 per hour.
-Further for original data, wages greater than 75 dollars were set to be missing.
-However, in recent context, this value might be too low to be set as the maximum threshold.
+- there are many errors in the data, e.g. hourly wages greater than $30,000 per hour.
 - there is no explicit variable in the database recording high school dropout, which means we needed to compare date of 12th grade with the individual's age.

### 2021-09-29 paper/paper.tex
figures in paper updated so uses tag instead of title

@@
 \keywords{
-    Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression
+    Data cleaning; Data tidying; Reproducible workflow; Longitudinal data; NLSY79; Initial data analysis;
   }
@@
 This is done using the \texttt{tidyverse} suite of packages (Wickham et al.
-2019), \texttt{tidyr} (Wickham 2020b) to pivot the data into long form, \texttt{dplyr} (Wickham et al.
+2019), \texttt{tidyr} (Wickham 2020) to pivot the data into long form, \texttt{dplyr} (Wickham et al.
 2020), and \texttt{stringr} (Wickham 2019) for creating new variables, and levels of factors by text wrangling.
@@
 
-In our final target data, we wish to include the demographic variables with variable names specified in brackets: gender (\texttt{gender}), race (\texttt{race}), age (\texttt{age}), highest grade completed (\texttt{hgc}) and the corresponding year this grade was completed (\texttt{yr\_hgc}).
+In our final target data, we wish to include the demographic variables with variable names specified in brackets: gender (\texttt{gender}), race (\texttt{race}), age (\texttt{age\_1979}), highest grade completed (\texttt{hgc}), highest grade completed in terms of years, e.g.~9th grade = 9, 3rd year college = 15, (\texttt{hgc\_i}) and the corresponding year this grade was completed (\texttt{yr\_hgc}).
 
@@
 
-\hypertarget{replacing-extreme-values}{%
-\subsubsection{Replacing extreme values}\label{replacing-extreme-values}}
+\hypertarget{censor}{%
+\subsubsection{Replacing extreme values}\label{censor}}
 
@@
 However, this method is too computationally and memory expensive, especially for a large data set, like the NLSY79 data.
-Thus, the model for each ID is built utilizing the \texttt{nest} and \texttt{map} function from \texttt{tidyr} (Wickham 2020b) and \texttt{purrr} (Henry and Wickham 2020), respectively.
+Thus, the model for each ID is built utilizing the \texttt{nest} and \texttt{map} function from \texttt{tidyr} (Wickham 2020) and \texttt{purrr} (Henry and Wickham 2020), respectively.
 
@@
 
-Further, \ref{fig:fixed-feature-plot} A shows that after eliminating the extreme values, the highest value has decreased to be around \$350.
+Further, Figure \ref{fig:fixed-feature-plot} A shows that after eliminating the extreme values, the highest value has decreased to be around \$350.
 The spikes are still observed but not as extreme as the original data set.
@@
 We still can see some extreme values in maximum wages, but consider it as a natural variability of the data.
-In \ref{fig:fixed-feature-plot} C, we can see that after imputing the extreme value in ID=39, we can see how the the wages change over the years more clearly.
+In Figure \ref{fig:fixed-feature-plot} C, we can see that after imputing the extreme value in ID=39, we can see how the the wages change over the years more clearly.
 
@@
 
+The treatment of unlikely wages differ in the refreshed data.
+In the original data by Singer and Willett (2003), wages greater than \$75 are set to be missing.
+However, in recent context, this value is too low to be set as the maximum threshold.
+We opted to use the weights from a robust linear regression to determine what should be set as missing value as described in Section \ref{censor}.
+
 Figure \ref{fig:plotting-sw-do} shows a comparison of the two sets.
@@
 Determining the appropriate robustness weight to threshold the anomalous observations is documented, and a \texttt{shiny} (Chang et al. 2020) app is provided to assist.
-The current subset is made available in a new R package, called yowie.
+The current subset is made available in a new R package, called \texttt{yowie}.
 
@@
   there are many errors in the data, e.g.~hourly wages greater than \$30,000 per hour.
-Further for original data, wages greater than 75 dollars were set to be missing.
-However, in recent context, this value might be too low to be set as the maximum threshold.
 \item
@@
 
-The entire analysis is conducted using \texttt{R} (R Core Team 2020) in \texttt{rstudio} using these packages: \texttt{tidyverse} (Wickham et al.
+The entire analysis is conducted using \texttt{R} (R Core Team 2020) in RStudio IDE using these packages: \texttt{tidyverse} (Wickham et al.
 2019), \texttt{ggplot2} (Wickham 2016), \texttt{dplyr} (Wickham et al.
-2020), \texttt{readr} (Wickham and Hester 2020), \texttt{tidyr} (Wickham 2020b), \texttt{stringr} (Wickham 2019), \texttt{purrr} (Henry and Wickham 2020), \texttt{broom} (Robinson, Hayes, and Couch 2021), \texttt{brolgar} (Tierney, Cook, and Prvan 2020), \texttt{patchwork} (Pedersen 2020), \texttt{kableExtra} (Zhu 2019), \texttt{MASS} (Venables and Ripley 2002), \texttt{janitor} (Firke 2020), \texttt{DiagrammeR} (Iannone 2020), \texttt{rsvg} (Ooms 2020), \texttt{webshot} (Chang 2019), \texttt{mgcv} (Wood 2003), \texttt{tsibble} (Wang, Cook, and Hyndman 2020), and \texttt{modelr} (Wickham 2020a).
-The paper are generated using \texttt{knitr} (Xie 2014), \texttt{rmarkdown} (Xie, Dervieux, and Riederer 2020), and \texttt{rticles} (Allaire et al.
-2021).
+2020), \texttt{readr} (Wickham and Hester 2020), \texttt{tidyr} (Wickham 2020), \texttt{stringr} (Wickham 2019), \texttt{purrr} (Henry and Wickham 2020), \texttt{brolgar} (Tierney, Cook, and Prvan 2020), \texttt{patchwork} (Pedersen 2020), \texttt{kableExtra} (Zhu 2019), \texttt{MASS} (Venables and Ripley 2002), \texttt{janitor} (Firke 2020), and \texttt{tsibble} (Wang, Cook, and Hyndman 2020).
+The paper are generated using \texttt{knitr} (Xie 2014) and \texttt{rmarkdown} (Xie, Dervieux, and Riederer 2020).
 
@@
 \item
-  \textbf{shiny app}: An web interactive \texttt{shiny} app to customize the weight threshold. This app could be accessed \href{https://github.com/numbats/wages-refresh/tree/main/app}{here}.
+  \textbf{shiny app}: An interactive \texttt{shiny} web app to visualise the effect of selecting different weight threshold for substituting the wages data to its predicted value from a fit of the robust linear regression model.
+This app could be accessed \href{https://ebsmonash.shinyapps.io/yowie_app/}{here} with the source code provided \href{https://github.com/numbats/yowie/tree/master/app}{here}.
 \end{itemize}
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-rticles}{}%
-Allaire, JJ, Yihui Xie, R Foundation, Hadley Wickham, Journal of Statistical Software, Ramnath Vaidyanathan, Association for Computing Machinery, et al.
-2021.
-\emph{Rticles: Article Formats for r Markdown}.
-\url{https://CRAN.R-project.org/package=rticles}.
-
-\leavevmode\hypertarget{ref-webshot}{}%
-Chang, Winston. 2019. \emph{Webshot: Take Screenshots of Web Pages}. \url{https://CRAN.R-project.org/package=webshot}.
-
 \leavevmode\hypertarget{ref-shiny}{}%
@@
 
-\leavevmode\hypertarget{ref-DiagrammeR}{}%
-Iannone, Richard. 2020. \emph{DiagrammeR: Graph/Network Visualization}. \url{https://CRAN.R-project.org/package=DiagrammeR}.
-
 \leavevmode\hypertarget{ref-KollerManuel2016rARP}{}%
 Koller, Manuel. 2016. {``Robustlmm: An r Package for Robust Estimation of Linear Mixed-Effects Models.''} \emph{Journal of Statistical Software} 75 (6): 1--24.
-
-\leavevmode\hypertarget{ref-rsvg}{}%
-Ooms, Jeroen. 2020. \emph{Rsvg: Render SVG Images into PDF, PNG, PostScript, or Bitmap Arrays}. \url{https://CRAN.R-project.org/package=rsvg}.
 
@@
 R Core Team. 2020. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.
-
-\leavevmode\hypertarget{ref-broom}{}%
-Robinson, David, Alex Hayes, and Simon Couch. 2021. \emph{Broom: Convert Statistical Objects into Tidy Tibbles}. \url{https://CRAN.R-project.org/package=broom}.
 
@@
 
-\leavevmode\hypertarget{ref-modelr}{}%
----------. 2020a. \emph{Modelr: Modelling Functions That Work with the Pipe}. \url{https://CRAN.R-project.org/package=modelr}.
-
 \leavevmode\hypertarget{ref-tidyr}{}%
----------. 2020b. \emph{Tidyr: Tidy Messy Data}. \url{https://CRAN.R-project.org/package=tidyr}.
+---------. 2020. \emph{Tidyr: Tidy Messy Data}. \url{https://CRAN.R-project.org/package=tidyr}.
 
@@
 Wickham, Hadley, and Jim Hester. 2020. \emph{Readr: Read Rectangular Text Data}. \url{https://CRAN.R-project.org/package=readr}.
-
-\leavevmode\hypertarget{ref-mgcv}{}%
-Wood, S. N. 2003. {``Thin-Plate Regression Splines.''} \emph{Journal of the Royal Statistical Society (B)} 65 (1): 95--114.
 

### 2021-09-29 data-raw/eda.R

@@ -7,7 +7,7 @@ sw_wages <- sw %>%
              y = ln_wages)) +
   geom_line(aes(group = id), alpha = 0.1) +
   geom_smooth(se = FALSE, colour = "pink") +
-  ggtitle("A original") +
+  labs(tag = "(A)", title = "Original") +
   theme_bw() +
   ylab("Hourly wage ($, natural log)") +
   xlab("Experience in years") +
@@ -26,7 +26,7 @@ sw_wages_mod <- sw %>%
              linetype = hgc)) +
   geom_smooth(method = "lm", se = FALSE,
               colour = "black") +
-  ggtitle("A original") +
+  labs(tag = "(A)", title = "Original") +
   scale_linetype("") +
   theme_bw() +
   ylab("Hourly wage ($, natural log)") +
@@ -53,7 +53,7 @@ do_ref <- do %>%
              y = lnwage)) +
   geom_line(aes(group = id), alpha = 0.1) +
   geom_smooth(se = FALSE, colour = "pink") +
-  ggtitle("B refreshed data") +
+  labs(tag = "(B)", title = "Refreshed data") +
   theme_bw() +
   ylab("Hourly wage ($, natural log)") +
   xlab("Year of data collection") +
@@ -67,7 +67,7 @@ do_ref_mod <- do %>%
              linetype = hgc12)) +
   geom_smooth(method = "lm", se = FALSE,
               colour = "black") +
-  ggtitle("B refreshed data") +
+  labs(title = "Refreshed data", tag = "(B)") +
   scale_linetype("") +
   theme_bw() +
   ylab("Hourly wage ($, natural log)") +
@@ -128,7 +128,7 @@ sw_wages_agree <- sw_agree %>%
              y = ln_wages)) +
   geom_line(aes(group = id),alpha = 0.1) +
   geom_smooth(se = FALSE) +
-  ggtitle("A") +
+  labs(tag = "(A)") +
   theme_bw() +
   ylab("ln(Hourly wage) ($)") +
   xlab("Experience (years)") +
@@ -141,7 +141,7 @@ do_ref_agree <- do_agree %>%
              y = lnwage)) +
   geom_line(aes(group = id), alpha = 0.1) +
   geom_smooth(se = FALSE) +
-  ggtitle("B") +
+  labs(tag = "B") +
   theme_bw() +
   ylab("ln(Hourly wage) ($)") +
   xlab("Year") +

### 2021-09-29 data-raw/ida.R

@@ -93,7 +93,7 @@ spag <- wages_before %>%
                      breaks = seq(1980, 2020, 10),
                      #labels = c("80", "90", "00", "10", "20"),
                      minor_breaks = seq(1980, 2020, 5)) +
-  ggtitle("A") +
+  labs(tag = "(A)") +
   theme_bw() +
   ylab("Hourly wage ($)") #+
   #theme(plot.title = element_text(size = 10))
@@ -109,7 +109,7 @@ wages_feat_long <- wages_three_feat %>%
 
 feature <- ggplot(wages_feat_long) +
   geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
-  ggtitle("B") +
+  labs(tag = "(B)") +
   theme_bw() +
   theme(#plot.title = element_text(size = 10),
         #axis.text.x = element_text(angle = 10, size = 6),
@@ -120,12 +120,10 @@ feature_bp <- ggplot(wages_feat_long,
                          fill = feature, color = feature)) +
   geom_boxplot() +
   theme_bw() +
-  #ggtitle("C") +
-  ggtitle("B")  +
+  labs(tag = "(B)")  +
   ylab("Hourly wage ($)") +
   xlab("Feature") +
-  theme(legend.position = "none") #,
-  #      plot.title = element_text(size = 10))
+  theme(legend.position = "none")
 
 plot_high <- ggplot(filter(wages_high2, id == 39)) +
   geom_line(aes(x = year,
@@ -141,10 +139,7 @@ plot_high <- ggplot(filter(wages_high2, id == 39)) +
                      minor_breaks = seq(1980, 2020, 5)) +
   ylab("Hourly wage ($)") +
   theme_bw() +
-  ggtitle("C") #+
-  #theme(axis.text.x = element_text(angle = 10, size = 6),
-   #     plot.title = element_text(size = 10))
-  #ggtitle("D")
+  labs("(C)")
 
 #spag + feature + feature_bp + plot_high + plot_layout(nrow = 1, guides = "collect") &
 #  theme(legend.position = "bottom")
@@ -187,8 +182,7 @@ ggplot(wages_compare) +
   #xlab("Year") +
   ylab("Hourly wage ($)") +
 #  scale_colour_brewer("", palette = "Dark2", direction = -1,
-  scale_colour_grey("",
-                        labels = c("Before", "After"))
[…]

### 2021-09-29 paper/paper-jsdse/paper-jsdse.tex
warning text about textbook data not being the best textbook data and future updates possible to come

@@
 \noindent%
-{\it Keywords:}  Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression
+{\it Keywords:}  Data cleaning; Data tidying; Reproducible workflow; Longitudinal data; NLSY79; Initial data analysis;
 \vfill
@@
 The last step required would be to inflation-adjust wages, but this is better to do with each wave of new data added, so that it is relative to the last date in the data.
+
+Some readers may disagree with our decisions made to produce the refreshed textbook data and may have better insight than us in producing a more appropriate textbook data.
+We do not assert that we have produced the best textbook data, but rather we describe our journey to provide a reasonable textbook data.
+All code and documentation are provided for transparency and future updates of \texttt{yowie} package may contain additional versions of the data if we come to have better understanding of a valid data.
 

### 2021-09-29 paper/paper.Rmd
warning text about textbook data not being the best textbook data and future updates possible to come

@@
 
+Some readers may disagree with our decisions made to produce the refreshed textbook data and may have better insight than us in producing a more appropriate textbook data.
+We do not assert that we have produced the best textbook data, but rather we describe our journey to provide a reasonable textbook data.
+All code and documentation are provided for transparency and future updates of `yowie` package may contain additional versions of the data if we come to have better understanding of a valid data.
+
 Finally, for the data providers we recommend that a validation system with clear rules is added on data entry, and that alternative output formats, such as a tidy format would help users make better use of the resource.

### 2021-09-29 paper/paper.tex
warning text about textbook data not being the best textbook data and future updates possible to come

@@
 
+Some readers may disagree with our decisions made to produce the refreshed textbook data and may have better insight than us in producing a more appropriate textbook data.
+We do not assert that we have produced the best textbook data, but rather we describe our journey to provide a reasonable textbook data.
+All code and documentation are provided for transparency and future updates of \texttt{yowie} package may contain additional versions of the data if we come to have better understanding of a valid data.
+
 Finally, for the data providers we recommend that a validation system with clear rules is added on data entry, and that alternative output formats, such as a tidy format would help users make better use of the resource.

### 2021-09-29 paper/paper-jsdse/paper-jsdse.tex
corrected a few typos

@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/Users/etan0038/yowie/paper/figures/flowchart}
+{\centering \includegraphics[width=0.85\linewidth]{/Users/cookd/Rpackages/yowie/paper/figures/flowchart}
 
@@
 Some readers may disagree with our decisions made to produce the refreshed textbook data and may have better insight than us in producing a more appropriate textbook data.
-We do not assert that we have produced the best textbook data, but rather we describe our journey to provide a reasonable textbook data.
-All code and documentation are provided for transparency and future updates of \texttt{yowie} package may contain additional versions of the data if we come to have better understanding of a valid data.
+We do not assert that we have produced the best textbook data, but rather we describe our journey to provide a reasonable textbook data set.
+All code and documentation are provided for transparency, and future updates of the \texttt{yowie} package may contain additional variables, or filters of the full set, if it is deemed important.
 
@@
 \item
-  \textbf{Codes} : R script to reproduce data tidying and cleaning are available in this \href{https://numbats.github.io/yowie/articles/process-data.html}{page}.
+  \textbf{Codes}: R script to reproduce data tidying and cleaning are available in this \href{https://numbats.github.io/yowie/articles/process-data.html}{page}.
 \item
-  \textbf{R Package \texttt{yowie}}:\texttt{yowie} is a data container R package that contains 3 datasets, namely the high school mean hourly wage data, high school dropouts mean hourly wage data, and demographic data of the NLSY79 cohort.
+  \textbf{R Package}:\texttt{yowie} is a data container R package that contains 3 datasets, namely the high school mean hourly wage data, high school dropouts mean hourly wage data, and demographic data of the NLSY79 cohort.
 This package could be accessed \href{https://github.com/numbats/yowie}{here}.

### 2021-09-29 paper/paper.Rmd
corrected a few typos

--- 2021-09-29 paper/paper.Rmd
@@
 
+Some readers may disagree with our decisions made to produce the refreshed textbook data and may have better insight than us in producing a more appropriate textbook data.
+We do not assert that we have produced the best textbook data, but rather we describe our journey to provide a reasonable textbook data set.
+All code and documentation are provided for transparency, and future updates of the `yowie` package may contain additional variables, or filters of the full set, if it is deemed important.
+
 Finally, for the data providers we recommend that a validation system with clear rules is added on data entry, and that alternative output formats, such as a tidy format would help users make better use of the resource.
@@
 
-- **Codes** : R script to reproduce data tidying and cleaning are available in this [page](https://numbats.github.io/yowie/articles/process-data.html).
+- **Codes**: R script to reproduce data tidying and cleaning are available in this [page](https://numbats.github.io/yowie/articles/process-data.html).
 
-- **R Package `yowie`**:`yowie` is a data container R package that contains 3 datasets, namely the high school mean hourly wage data, high school dropouts mean hourly wage data, and demographic data of the NLSY79 cohort.
+- **R Package**:`yowie` is a data container R package that contains 3 datasets, namely the high school mean hourly wage data, high school dropouts mean hourly wage data, and demographic data of the NLSY79 cohort.
 This package could be accessed [here](https://github.com/numbats/yowie).

### 2021-09-29 data-raw/ida.R

@@ -137,9 +137,8 @@ plot_high <- ggplot(filter(wages_high2, id == 39)) +
                      breaks = seq(1980, 2020, 10),
                      #labels = c("80", "90", "00", "10", "20"),
                      minor_breaks = seq(1980, 2020, 5)) +
-  ylab("Hourly wage ($)") +
   theme_bw() +
-  labs("(C)")
+  labs(tag = "(C)", y = "Hourly wage ($)")
 
 #spag + feature + feature_bp + plot_high + plot_layout(nrow = 1, guides = "collect") &
 #  theme(legend.position = "bottom")

### 2021-09-29 paper/paper-jsdse/paper-jsdse.Rmd
cleaned the references up

@@
 bibliography: ../references.bib
-biblio-style: unsrt
 #preamble: >

### 2021-09-29 paper/paper-jsdse/paper-jsdse.tex
cleaned the references up

--- 2021-09-29 paper/paper-jsdse/paper-jsdse.tex
@@
 
-The NLSY79 is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth born between 1957-1964 (The U.S.
-Bureau of Labor Statistics, n.d.).
+The NLSY79 is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth born between 1957-1964 (Bureau of Labor Statistics, U.S.
+Department of Labor 2021).
 The cohort originally included 12,686 respondents aged 14-22 when first interviewed in 1979.
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-nlsy79}{}%
+Bureau of Labor Statistics, U.S.
+Department of Labor.
+2021.
+{``National Longitudinal Survey of Youth 1979 Cohort, 1979-2016 (Rounds 1-28).''} Produced and distributed by the Center for Human Resource Research (CHRR), The Ohio State University.
+Columbus, OH.
+
 \leavevmode\hypertarget{ref-shiny}{}%
-Chang, Winston, Joe Cheng, JJ Allaire, Yihui Xie, and Jonathan McPherson. 2020. \emph{Shiny: Web Application Framework for r}. \url{https://CRAN.R-project.org/package=shiny}.
+Chang, Winston, Joe Cheng, JJ Allaire, Yihui Xie, and Jonathan McPherson. 2020. \emph{{shiny: Web Application Framework for R}}. \url{https://CRAN.R-project.org/package=shiny}.
 
@@
 \leavevmode\hypertarget{ref-janitor}{}%
-Firke, Sam. 2020. \emph{Janitor: Simple Tools for Examining and Cleaning Dirty Data}. \url{https://CRAN.R-project.org/package=janitor}.
+Firke, Sam. 2020. \emph{{janitor: Simple Tools for Examining and Cleaning Dirty Data}}. \url{https://CRAN.R-project.org/package=janitor}.
 
@@
 \leavevmode\hypertarget{ref-purrr}{}%
-Henry, Lionel, and Hadley Wickham. 2020. \emph{Purrr: Functional Programming Tools}. \url{https://CRAN.R-project.org/package=purrr}.
+Henry, Lionel, and Hadley Wickham. 2020. \emph{{purrr: Functional Programming Tools}}. \url{https://CRAN.R-project.org/package=purrr}.
 
@@
 \leavevmode\hypertarget{ref-KollerManuel2016rARP}{}%
-Koller, Manuel. 2016. {``Robustlmm: An r Package for Robust Estimation of Linear Mixed-Effects Models.''} \emph{Journal of Statistical Software} 75 (6): 1--24.
+Koller, Manuel. 2016. {``{robustlmm: An R Package for Robust Estimation of Linear Mixed-Effects Models}.''} \emph{Journal of Statistical Software} 75 (6): 1--24.
 
@@
 \leavevmode\hypertarget{ref-patchwork}{}%
-Pedersen, Thomas Lin. 2020. \emph{Patchwork: The Composer of Plots}. \url{https://CRAN.R-project.org/package=patchwork}.
+Pedersen, Thomas Lin. 2020. \emph{{patchwork: The Composer of Plots}}. \url{https://CRAN.R-project.org/package=patchwork}.
 
@@
 
-\leavevmode\hypertarget{ref-nlsy79}{}%
-The U.S. Bureau of Labor Statistics. n.d. {``National Longitudinal Survey of Youth 1979.''} Available at \url{https://www.nlsinfo.org/content/cohorts/nlsy79} (2021/25/02).
-
 \leavevmode\hypertarget{ref-brolgar}{}%
-Tierney, Nicholas, Di Cook, and Tania Prvan. 2020. \emph{Brolgar: BRowse over Longitudinal Data Graphically and Analytically in r}. \url{https://github.com/njtierney/brolgar}.
+Tierney, Nicholas, Di Cook, and Tania Prvan. 2020. \emph{{brolgar: BRowse Over Longitudinal data Graphically and Analytically in R}}. \url{https://github.com/njtierney/brolgar}.
 
@@
 \leavevmode\hypertarget{ref-rlm}{}%
-UCLA: Statistical Consulting Group. 2021. {``Robust Regression \textbar{} r Data Analysis Examples.''} February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.
+UCLA: Statistical Consulting Group. 2021. {``{Robust Regression \textbar{} R Data Analysis Examples}.''} February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.
 
@@
 \leavevmode\hypertarget{ref-ggplot2}{}%
----------. 2016. \emph{Ggplot2: Elegant Graphics for Data Analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.
+---------. 2016. \emph{{ggplot2: Elegant Graphics for Data Analysis}}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.
 
 \leavevmode\hypertarget{ref-stringr}{}%
----------. 2019. \emph{Stringr: Simple, Consistent Wrappers for Common String Operations}. \url{https://CRAN.R-project.org/package=stringr}.
+---------. 2019. \emph{{stringr: Simple, Consistent Wrappers for Common String Operations}}. \url{https://CRAN.R-project.org/package=stringr}.
 
 \leavevmode\hypertarget{ref-tidyr}{}%
----------. 2020. \emph{Tidyr: Tidy Messy Data}. \url{https://CRAN.R-project.org/package=tidyr}.
+---------. 2020. \emph{{tidyr: Tidy Messy Data}}. \url{https://CRAN.R-project.org/package=tidyr}.
 
@@
 \leavevmode\hypertarget{ref-dplyr}{}%
-Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2020. \emph{Dplyr: A Grammar of Data Manipulation}. \url{https://CRAN.R-project.org/package=dplyr}.
+Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2020. \emph{{dplyr: A Grammar of Data Manipulation}}. \url{https://CRAN.R-project.org/package=dplyr}.
 
 \leavevmode\hypertarget{ref-readr}{}%
-Wickham, Hadley, and Jim Hester. 2020. \emph{Readr: Read Rectangular Text Data}. \url{https://CRAN.R-project.org/package=readr}.
+Wickham, Hadley, and Jim Hester. 2020. \emph{{readr: Read Rectangular Text Data}}. \url{https://CRAN.R-project.org/package=readr}.
 
@@
 \leavevmode\hypertarget{ref-kableExtra}{}%
-Zhu, Hao. 2019. \emph{kableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.
+Zhu, Hao. 2019. \emph{{kableExtra: Construct Complex Table with 'kable' and Pipe Syntax}}. \url{https://CRAN.R-project.org/package=kableExtra}.
 

### 2021-09-29 paper/cover-letter/cover-letter.Rmd

---
from: Dewi Amaliah
fromqualifications: Master of Business Analytics
fromWWW: "[https://numbats.github.io/yowie/](numbats.github.io/yowie/)"
fromemail: dlamaleeah@gmail.com
fromphone: "+61 3 9905 2608"
fromsignature: sigfile.png
toname: Prof. Nick Horton
toaddress: Editor, JSDSE 
opening: Dear Prof. Horton
closing: Regards
spacing: 1.4
output: MonashEBSTemplates::letter
---

Please consider our paper titled "The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth" for the JDSDE Special Issue on "Teaching Reproducibility and Responsible Workflow".

The paper is co-authored by Prof. Di Cook, Dr Emi Tanaka, Kate Hyde and Dr Nick Tierney. We believe it is a good fit for the special issue because it covers various aspects of data management and curation component of data acumen, including data preparation and cleaning, missing and conflicting data. The code to reproduce the data cleaning, and make further updates of the resulting data is provided.

### 2021-09-30 paper/cover-letter/cover-letter.Rmd
Edit phone number in cover letter.

@@
 fromemail: dlamaleeah@gmail.com
-fromphone: "+61 3 9905 2608"
+fromphone: "+61 435 444 025"
 fromsignature: sigfile.png

### 2021-09-30 paper/paper-jsdse/paper-jsdse.Rmd
blind option on

@@
     email: ""
+blind: 1 # 0 for no
 bibliography: ../references.bib

### 2021-09-30 paper/paper-jsdse/paper-jsdse.tex
blind option on

@@
 % NOTE: To produce blinded version, replace "0" with "1" below.
-\newcommand{\blind}{0}
+\newcommand{\blind}{1}
 
@@
   \begin{center}
-    {\LARGE\bf Title}
+    {\LARGE\bf The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth}
 \end{center}
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/Users/cookd/Rpackages/yowie/paper/figures/flowchart}
+{\centering \includegraphics[width=0.85\linewidth]{/Users/etan0038/yowie/paper/figures/flowchart}
 

### 2021-09-30 paper/paper-jsdse/paper-jsdse.tex
censoring implemented for pkgs and links

@@
 \usepackage{amsmath}
+%\usepackage{longtable}
 \usepackage{graphicx,psfrag,epsf}
@@
 Thus, this paper demonstrates the process -- extracting, tidying, cleaning, and exploring with documentation -- to make refreshed data available for education or research.
-Three new data sets, and the code to produce them are provided in an accompanying open source R package, called \texttt{yowie}.
+Three new data sets, and the code to produce them are provided in an accompanying open source R package, called {[}CENSORED{]}.
 As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
@@
 
+M.
 van der Loo and de Jonge (2018) describe the notion of a ``statistical value chain'' where the production stages of data cleaning process are earmarked as raw data (data from the data source), input data (data organised with correct type and identified variables) and valid data (data that faithfully represent the variables).
@@
 The long form of the data makes it possible to do these data transformations efficiently, and it is an intermediate step towards the final target data.
-The code for tidying the data are demonstrated at \url{https://numbats.github.io/yowie/articles/raw-to-input-data.html} but also described in the subsequent subsections.
+The code for tidying the data are demonstrated at \url{CENSORED/articles/raw-to-input-data.html} but also described in the subsequent subsections.
 
@@
 Hence, we only used the highest grade completed ever and derived the year when it is completed.
-The code for deriving this variable is also available at \url{https://numbats.github.io/yowie/articles/raw-to-input-data.html}.
+The code for deriving this variable is also available at \url{CENSORED/articles/raw-to-input-data.html}.
 
@@
 
-Figure \ref{fig:flow-chart} summarizes the steps taken to go from raw to input to valid data (van der Loo and de Jonge 2018) to create a refreshed wages data set.
+Figure \ref{fig:flow-chart} summarizes the steps taken to go from raw to input to valid data (M. van der Loo and de Jonge 2018) to create a refreshed wages data set.
 
@@
 
-\caption{The stages of data cleaning from the raw data to get three datasets contained yowie.
+\caption{The stages of data cleaning from the raw data to get three datasets contained \texttt{[CENSORED]}.
 \# of individuals means the number of respondents included in each stage, while \# of observations means the number of rows in the data.
-The color represents the stage of data cleaning in statistical value chain (van der Loo \& de Jonge, 2018).
+The color represents the stage of data cleaning in statistical value chain (M.
+P.
+J.
+van der Loo and de Jonge 2021).
 Pink, blue, and green represents the raw, input, and valid data, respectively.}\label{fig:flow-chart}
@@
 The refreshed data set for comparison is also wages on high school drop outs, over 1979 to 2018.
-The original set is available in the R package \texttt{brolgar} and the refreshed data is available in the R package \texttt{yowie}.
+The original set is available in the R package \texttt{brolgar} and the refreshed data is available in the R package \texttt{[CENSORED]}.
 
@@
 Determining the appropriate robustness weight to threshold the anomalous observations is documented, and a \texttt{shiny} (Chang et al. 2020) app is provided to assist.
-The current subset is made available in a new R package, called \texttt{yowie}.
+The current subset is made available in a new R package, called \texttt{[CENSORED]}.
 
@@
 We do not assert that we have produced the best textbook data, but rather we describe our journey to provide a reasonable textbook data set.
-All code and documentation are provided for transparency, and future updates of the \texttt{yowie} package may contain additional variables, or filters of the full set, if it is deemed important.
+All code and documentation are provided for transparency, and future updates of the \texttt{[CENSORED]} package may contain additional variables, or filters of the full set, if it is deemed important.
 
@@
 \item
-  \textbf{Codes}: R script to reproduce data tidying and cleaning are available in this \href{https://numbats.github.io/yowie/articles/process-data.html}{page}.
+  \textbf{Codes}: R script to reproduce data tidying and cleaning are available in this \href{CENSORED/articles/process-data.html}{page}.
 \item
-  \textbf{R Package}:\texttt{yowie} is a data container R package that contains 3 datasets, namely the high school mean hourly wage data, high school dropouts mean hourly wage data, and demographic data of the NLSY79 cohort.
-This package could be accessed \href{https://github.com/numbats/yowie}{here}.
+  \textbf{R Package}:\texttt{[CENSORED]} is a data container R package that contains 3 datasets, namely the high school mean hourly wage data, high school dropouts mean hourly wage data, and demographic data of the NLSY79 cohort.
+This package could be accessed \href{CENSORED}{here}.
 \item
   \textbf{shiny app}: An interactive \texttt{shiny} web app to visualise the effect of selecting different weight threshold for substituting the wages data to its predicted value from a fit of the robust linear regression model.
-This app could be accessed \href{https://ebsmonash.shinyapps.io/yowie_app/}{here} with the source code provided \href{https://github.com/numbats/yowie/tree/master/app}{here}.
+This app could be accessed \href{CENSORED}{here} with the source code provided \href{CENSORED/tree/master/app}{here}.
 \end{itemize}
@@
 
+\leavevmode\hypertarget{ref-validate}{}%
+van der Loo, Mark P. J., and Edwin de Jonge. 2021. {``Data Validation Infrastructure for {R}.''} \emph{Journal of Statistical Software} 97 (10): 1--31. \url{https://doi.org/10.18637/jss.v097.i10}.
+
 \leavevmode\hypertarget{ref-LooMarkvander2018Sdcw}{}%

### 2021-09-30 paper/paper.Rmd
censoring implemented for pkgs and links

@@
     email: ""
-abstract: > 
- The National Longitudinal Survey of Youth (NLSY79) is a prominent open data source that has been important for educational purposes and multidisciplinary research on longitudinal data.
-Subsets of this data can be found in numerous textbooks and research articles.
-However, the steps and decisions taken to get from the raw data to the textbook data is never clearly articulated.
-This article describes our journey when trying to re-create a textbook example data set from the original database, with the goal being to refresh the textbook data more regularly.
-Thus, this paper demonstrates the process -- extracting, tidying, cleaning, and exploring with documentation -- to make refreshed data available for education or research.
-Three new data sets, and the code to produce them are provided in an accompanying open source R package, called `yowie`.
-As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
 bibliography: references.bib
@@
 keywords: "Data cleaning; Data tidying; Reproducible workflow; Longitudinal data; NLSY79; Initial data analysis;"
+abstract: > 
+ The National Longitudinal Survey of Youth (NLSY79) is a prominent open data source that has been important for educational purposes and multidisciplinary research on longitudinal data.
+Subsets of this data can be found in numerous textbooks and research articles.
+However, the steps and decisions taken to get from the raw data to the textbook data is never clearly articulated.
+This article describes our journey when trying to re-create a textbook example data set from the original database, with the goal being to refresh the textbook data more regularly.
+Thus, this paper demonstrates the process -- extracting, tidying, cleaning, and exploring with documentation -- to make refreshed data available for education or research.
+Three new data sets, and the code to produce them are provided in an accompanying open source R package, called [CENSORED].
+As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
 ---
+
+```{r censor, include = FALSE}
+CENSOR <- TRUE
+yowie <- ifelse(CENSOR, "\\texttt{[CENSORED]}", "\\texttt{yowie}")
+yowie_repo <- ifelse(CENSOR, "CENSORED", "https://github.com/numbats/yowie")
+yowie_pkgdown <- ifelse(CENSOR, "CENSORED", "https://github.com/numbats/yowie")
+yowie_shiny <- ifelse(CENSOR, "CENSORED", "https://ebsmonash.shinyapps.io/yowie_app/")
+```
 
@@
 The long form of the data makes it possible to do these data transformations efficiently, and it is an intermediate step towards the final target data.
-The code for tidying the data are demonstrated at \url{https://numbats.github.io/yowie/articles/raw-to-input-data.html} but also described in the subsequent subsections.
+The code for tidying the data are demonstrated at \url{`r yowie_pkgdown`/articles/raw-to-input-data.html} but also described in the subsequent subsections.
 
@@
 Hence, we only used the highest grade completed ever and derived the year when it is completed.
-The code for deriving this variable is also available at \url{https://numbats.github.io/yowie/articles/raw-to-input-data.html}.
+The code for deriving this variable is also available at \url{`r yowie_pkgdown`/articles/raw-to-input-data.html}.
 
@@
 
-```{r flow-chart, fig.cap="The stages of data cleaning from the raw data to get three datasets contained yowie.
-\\# of individuals means the number of respondents included in each stage, while \\# of observations means the number of rows in the data.
-The color represents the stage of data cleaning in statistical value chain (van der Loo \\& de Jonge, 2018).
-Pink, blue, and green represents the raw, input, and valid data, respectively.", out.width="85%"}
+(ref:flow-chart) The stages of data cleaning from the raw data to get three datasets contained `r yowie`.
+# of individuals means the number of respondents included in each stage, while # of observations means the number of rows in the data.
+The color represents the stage of data cleaning in statistical value chain [@validate].
+Pink, blue, and green represents the raw, input, and valid data, respectively.
+
+```{r flow-chart, fig.cap="(ref:flow-chart)", out.width="85%"}
 ```
@@
 The refreshed data set for comparison is also wages on high school drop outs, over 1979 to 2018.
-The original set is available in the R package `brolgar` and the refreshed data is available in the R package `yowie`.
+The original set is available in the R package `brolgar` and the refreshed data is available in the R package `r yowie`.
 
@@
 Determining the appropriate robustness weight to threshold the anomalous observations is documented, and a `shiny`  [@shiny] app is provided to assist. 
-The current subset is made available in a new R package, called `yowie`.
+The current subset is made available in a new R package, called `r yowie`.
 
@@
 We do not assert that we have produced the best textbook data, but rather we describe our journey to provide a reasonable textbook data set.
-All code and documentation are provided for transparency, and future updates of the `yowie` package may contain additional variables, or filters of the full set, if it is deemed important.
+All code and documentation are provided for transparency, and future updates of the `r yowie` package may contain additional variables, or filters of the full set, if it is deemed important.
 
@@
 
-- **Codes**: R script to reproduce data tidying and cleaning are available in this [page](https://numbats.github.io/yowie/articles/process-data.html).
+- **Codes**: R script to reproduce data tidying and cleaning are available in this [page](`r yowie_pkgdown`/articles/process-data.html).
 
-- **R Package**:`yowie` is a data container R package that contains 3 datasets, namely the high school mean hourly wage data, high school dropouts mean hourly wage data, and demographic data of the NLSY79 cohort.
-This package could be accessed [here](https://github.com/numbats/yowie).
+- **R Package**:`r yowie` is a data container R package that contains 3 datasets, namely the high school mean hourly wage data, high school dropouts mean hourly wage data, and demographic data of the NLSY79 cohort.
+This package could be accessed [here](`r yowie_repo`).
 
 - **shiny app**: An interactive `shiny` web app to visualise the effect of selecting different weight threshold for substituting the wages data to its predicted value from a fit of the robust linear regression model.
-This app could be accessed [here](https://ebsmonash.shinyapps.io/yowie_app/) with the source code provided [here](https://github.com/numbats/yowie/tree/master/app).
+This app could be accessed [here](`r yowie_shiny`) with the source code provided [here](`r yowie_repo`/tree/master/app).
 

### 2021-09-30 paper/paper.tex
censoring implemented for pkgs and links

--- 2021-09-29 paper/paper.tex
@@
 Thus, this paper demonstrates the process -- extracting, tidying, cleaning, and exploring with documentation -- to make refreshed data available for education or research.
-Three new data sets, and the code to produce them are provided in an accompanying open source R package, called \texttt{yowie}.
+Three new data sets, and the code to produce them are provided in an accompanying open source R package, called \texttt{[CENSORED]}.
 As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
@@
 The last step required would be to inflation-adjust wages, but this is better to do with each wave of new data added, so that it is relative to the last date in the data.
+
+Some readers may disagree with our decisions made to produce the refreshed textbook data and may have better insight than us in producing a more appropriate textbook data.
+We do not assert that we have produced the best textbook data, but rather we describe our journey to provide a reasonable textbook data set.
+All code and documentation are provided for transparency, and future updates of the \texttt{[CENSORED]} package may contain additional variables, or filters of the full set, if it is deemed important.
 
@@
 \item
-  \textbf{Codes} : R script to reproduce data tidying and cleaning are available in this \href{https://numbats.github.io/yowie/articles/process-data.html}{page}.
+  \textbf{Codes}: R script to reproduce data tidying and cleaning are available in this \href{CENSORED/articles/process-data.html}{page}.
 \item
-  \textbf{R Package \texttt{yowie}}:\texttt{yowie} is a data container R package that contains 3 datasets, namely the high school mean hourly wage data, high school dropouts mean hourly wage data, and demographic data of the NLSY79 cohort.
-This package could be accessed \href{https://github.com/numbats/yowie}{here}.
+  \textbf{R Package}:\texttt{[CENSORED]} is a data container R package that contains 3 datasets, namely the high school mean hourly wage data, high school dropouts mean hourly wage data, and demographic data of the NLSY79 cohort.
+This package could be accessed \href{CENSORED}{here}.
 \item
   \textbf{shiny app}: An interactive \texttt{shiny} web app to visualise the effect of selecting different weight threshold for substituting the wages data to its predicted value from a fit of the robust linear regression model.
-This app could be accessed \href{https://ebsmonash.shinyapps.io/yowie_app/}{here} with the source code provided \href{https://github.com/numbats/yowie/tree/master/app}{here}.
+This app could be accessed \href{CENSORED}{here} with the source code provided \href{CENSORED/tree/master/app}{here}.
 \end{itemize}
@@
 \leavevmode\hypertarget{ref-rlm}{}%
-UCLA: Statistical Consulting Group. 2021. {``Robust Regression \textbar{} r Data Analysis Examples.''} February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.
+UCLA: Statistical Consulting Group. 2021. {``{Robust Regression \textbar{} R Data Analysis Examples}.''} February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.
+
+\leavevmode\hypertarget{ref-validate}{}%
+van der Loo, Mark P. J., and Edwin de Jonge. 2021. {``Data Validation Infrastructure for {R}.''} \emph{Journal of Statistical Software} 97 (10): 1--31. \url{https://doi.org/10.18637/jss.v097.i10}.
 

### 2021-09-30 paper/paper-jsdse/paper-jsdse.tex
added citations to grimshaw and kim

@@
 Singer and Willett (2003) used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling of wages on workforce experience, with covariates education and race.
-Our aim is to refresh this textbook data with data from 1994 through to the latest data reported in 2018.
+Our aim is to refresh this textbook data with data from 1994 through to the latest data reported in 2018, a purpose that is consistent with Grimshaw (2015)'s statistics education goal of embracing authentic data experiences.
 Here, we investigate the process of getting from the raw NLSY79 data to a textbook data set as similar as possible to that provided by Singer and Willett (2003).
@@
 The data is first organised into ``tidy data'' (Wickham 2014) and then further wrangled using the data pipeline and split-apply-combine approach (Wickham 2011).
-The resulting data is provided in a new R package called \texttt{yowie} which includes the code so that the process is reproducible, and could be used to further refresh the data as new records are made available in the NLSY79 database.
+(Tidy data shouldn't be confused with ``tame data'' which Kim, Ismay, and Chunn (2018) coined to refer to text book data sets suitable for teaching, particularly teaching statistics.) The resulting (tame) data is provided in a new R package called \texttt{yowie} which includes the code so that the process is reproducible, and could be used to further refresh the data as new records are made available in the NLSY79 database.
 
@@
 \caption{The stages of data cleaning from the raw data to get three datasets contained \texttt{[CENSORED]}.
-\# of individuals means the number of respondents included in each stage, while \# of observations means the number of rows in the data.
+``\# of individuals'' means the number of respondents included in each stage, while ``\# of observations'' means the number of rows in the data.
 The color represents the stage of data cleaning in statistical value chain (M.
@@
 \leavevmode\hypertarget{ref-racismnotrace}{}%
-Fullilove, M. T. 1998. {``Comment: Abandoning "Race" as a Variable in Public Health Research--an Idea Whose Time Has Come.''} \emph{American Journal of Public Health} 88 (9): 1297--98.
+Fullilove, M. T. 1998. {``Comment: Abandoning ``Race" as a Variable in Public Health Research--an Idea Whose Time Has Come.''} \emph{American Journal of Public Health} 88 (9): 1297--98.
+
+\leavevmode\hypertarget{ref-grimshaw}{}%
+Grimshaw, Scott D.
+2015.
+{``A Framework for Infusing Authentic Data Experiences Within Statistics Courses.''} \emph{The American Statistician} 69 (4): 307--14.
+\url{https://doi.org/10.1080/00031305.2015.1081106}.
 
@@
 {``Hidden Analyses: A Review of Reporting Practice and Recommendations for More Transparent Reporting of Initial Data Analyses.''} \emph{BMC Medical Research Methodology} 20 (1): 61--61.
+
+\leavevmode\hypertarget{ref-tamedata}{}%
+Kim, A.
+Y, C.
+Ismay, and J.
+Chunn.
+2018.
+{``The Fivethirtyeight {R} Package: ``Tame Data" Principles for Introductory Statistics and Data Science Courses.''} \emph{Technology Innovations in Statistics Education} 11 (1).
+\url{https://doi.org/10.5070/T511103589}.
 

### 2021-09-30 paper/paper.Rmd
added citations to grimshaw and kim

@@
 @SingerJudithD2003Alda used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling of wages on workforce experience, with covariates education and race.
-Our aim is to refresh this textbook data with data from 1994 through to the latest data reported in 2018.
+Our aim is to refresh this textbook data with data from 1994 through to the latest data reported in 2018, a purpose that is consistent with @grimshaw's statistics education goal of embracing authentic data experiences.
 Here, we investigate the process of getting from the raw NLSY79 data to a textbook data set as similar as possible to that provided by @SingerJudithD2003Alda.
@@
 The data is first organised into "tidy data" [@WickhamHadley2014TD] and then further wrangled using the data pipeline and split-apply-combine approach [@plyr].
-The resulting data is provided in a new R package called `yowie` which includes the code so that the process is reproducible, and could be used to further refresh the data as new records are made available in the NLSY79 database.
+(Tidy data shouldn't be confused with "tame data" which @tamedata coined to refer to text book data sets suitable for teaching, particularly teaching statistics.) The resulting (tame) data is provided in a new R package called `yowie` which includes the code so that the process is reproducible, and could be used to further refresh the data as new records are made available in the NLSY79 database.
 
@@
 (ref:flow-chart) The stages of data cleaning from the raw data to get three datasets contained `r yowie`.
-# of individuals means the number of respondents included in each stage, while # of observations means the number of rows in the data.
+"# of individuals" means the number of respondents included in each stage, while "# of observations" means the number of rows in the data.
 The color represents the stage of data cleaning in statistical value chain [@validate].

### 2021-09-30 paper/paper.Rmd
Check spelling.

@@
 Subsets of this data can be found in numerous textbooks and research articles.
-However, the steps and decisions taken to get from the raw data to the textbook data is never clearly articulated.
+However, the steps and decisions taken to get from the raw data to the textbook data are never clearly articulated.
 This article describes our journey when trying to re-create a textbook example data set from the original database, with the goal being to refresh the textbook data more regularly.
 Thus, this paper demonstrates the process -- extracting, tidying, cleaning, and exploring with documentation -- to make refreshed data available for education or research.
-Three new data sets, and the code to produce them are provided in an accompanying open source R package, called [CENSORED].
-As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
+Three new data sets and the code to produce them are provided in an accompanying open source R package, called [CENSORED].
+As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks and providing more convenient samples of the data to potential users.
 ---
@@
 @LooMarkvander2018Sdcw describe the notion of a "statistical value chain" where the production stages of data cleaning process are earmarked as raw data (data from the data source), input data (data organised with correct type and identified variables) and valid data (data that faithfully represent the variables).
-In this section, we outline the steps to download the raw data (Section \@ref(getdata)) and then tidying the raw data into input data, specifically for wages data (Section \@ref(tidywages)), demographic variables (Section \@ref(tidydemog)) and the employment variables (Section \@ref(tidyemp)), so that the resulting input data can be used downstream for validating the data as described in Section \@ref(ida).
+In this section, we outline the steps to download the raw data (Section \@ref(getdata)) and then tidying the raw data into input data, specifically for wages data (Section \@ref(tidyemp)), demographic variables (Section \@ref(tidydemog)) and the employment variables (Section \@ref(tidyemp)), so that the resulting input data can be used downstream for validating the data as described in Section \@ref(ida).
 
@@
 
-The next step is that explore the mean hourly wage data. In this case, we only explore the wages data of up to $12^{th}$ grade.  Here, we use visualisation techniques to do IDA.
+The next step is that explore the mean hourly wage data. In this case, we only explore the wages data of up to $12^{th}$ grade.  Here, we use visualization techniques to do IDA.
 
@@
 
+# Data Availability Statement
+
+The authors confirm that the data supporting the findings of this study are available within the supplementary materials.
+
 # References

### 2021-09-30 paper/paper-jsdse/paper-jsdse.tex
navigation step changed to figure

@@
 \usepackage{booktabs}
+\usepackage{float}
 \usepackage{url} % not crucial - just used below for the URL
@@
 
-The NLSY79 data used in Singer and Willett (2003).
-contains the longitudinal measurements on yearly mean hourly wages with years of workforce experience, and demographic variables education and race, from 1979 through to 1994.
+The NLSY79 data used in Singer and Willett (2003) contains the longitudinal measurements on yearly mean hourly wages with years of workforce experience, and demographic variables education and race, from 1979 through to 1994.
 In addition, the cohort is restricted to male high-school dropouts who first participated in the study at age 14-17 years.
@@
 M.
-van der Loo and de Jonge (2018) describe the notion of a ``statistical value chain'' where the production stages of data cleaning process are earmarked as raw data (data from the data source), input data (data organised with correct type and identified variables) and valid data (data that faithfully represent the variables).
-In this section, we outline the steps to download the raw data (Section \ref{getdata}) and then tidying the raw data into input data, specifically for wages data (Section \ref{tidywages}), demographic variables (Section \ref{tidydemog}) and the employment variables (Section \ref{tidyemp}), so that the resulting input data can be used downstream for validating the data as described in Section \ref{ida}.
+van der Loo and de Jonge (2018) describe the notion of a ``statistical value chain'' where the production stages of data cleaning process are earmarked as raw data (data as arrived to the desk of an analyst), input data (data organised with correct type and identified variables) and valid data (data that faithfully represent the variables).
+In this section, we outline the steps to download the raw data (Section \ref{getdata}) and then tidying the raw data into input data, specifically for the demographic variables (Section \ref{tidydemog}) and the employment variables (Section \ref{tidyemp}), so that the resulting input data can be used downstream for validating the data as described in Section \ref{ida}.
 
@@
 The NLSY79 data contains a large scope of variables but for our aim, the scope required is limited to demographic profiles and wages data.
-More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}, clicked on the direct link to NSLY79 data and navigated as described below.
+More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}, clicked on the direct link to NSLY79 data and navigated as described in Figure \ref{fig:source-nav}.
 
-\begin{tcolorbox}[label="box:data", title=Navigating the data source to download the raw data]
+\begin{figure}[Ht]
 
-\faDatabase~NLSY79 (\url{https://www.nlsinfo.org/investigator/pages/search?s=NLSY79})\\
+\begin{tcolorbox}
+\faDatabase\ NLSY79 (\url{https://www.nlsinfo.org/investigator/pages/search?s=NLSY79})\\
 \vspace{1mm}
-\faCheck~The CASEID will be always be selected.\\
+\faCheck\ The CASEID will be always be selected.  \\
 \vspace{1mm}
-\faCheck~The 3 recommended demographic variable (sample ID, race and sex) were selected.\\
+\faCheck\ The 3 recommended demographic variable (sample ID, race and sex) were selected.  \\
 \vspace{1mm}
-For the remaining variables, we went to the ``Variable Search'' tab and select variables as follows
-
+For the remaining variables, we went to the "Variable Search" tab and select variables as follows
 \begin{itemize}
@@
 \end{itemize}
-
 \end{tcolorbox}
+\caption{Above shows the steps taken to navigate the data source to download the raw data.\label{fig:source-nav}}
+\end{figure}
 

### 2021-09-30 paper/paper.Rmd
navigation step changed to figure

--- 2021-09-30 paper/paper.Rmd
@@
 
-The NLSY79 data used in @SingerJudithD2003Alda.
-contains the longitudinal measurements on yearly mean hourly wages with years of workforce experience, and demographic variables education and race, from 1979 through to 1994.
+The NLSY79 data used in @SingerJudithD2003Alda contains the longitudinal measurements on yearly mean hourly wages with years of workforce experience, and demographic variables education and race, from 1979 through to 1994.
 In addition, the cohort is restricted to male high-school dropouts who first participated in the study at age 14-17 years.
@@
 
-@LooMarkvander2018Sdcw describe the notion of a "statistical value chain" where the production stages of data cleaning process are earmarked as raw data (data from the data source), input data (data organised with correct type and identified variables) and valid data (data that faithfully represent the variables).
-In this section, we outline the steps to download the raw data (Section \@ref(getdata)) and then tidying the raw data into input data, specifically for wages data (Section \@ref(tidywages)), demographic variables (Section \@ref(tidydemog)) and the employment variables (Section \@ref(tidyemp)), so that the resulting input data can be used downstream for validating the data as described in Section \@ref(ida).
+@LooMarkvander2018Sdcw describe the notion of a "statistical value chain" where the production stages of data cleaning process are earmarked as raw data (data as arrived to the desk of an analyst), input data (data organised with correct type and identified variables) and valid data (data that faithfully represent the variables).
+In this section, we outline the steps to download the raw data (Section \@ref(getdata)) and then tidying the raw data into input data, specifically for the demographic variables (Section \@ref(tidydemog)) and the employment variables (Section \@ref(tidyemp)), so that the resulting input data can be used downstream for validating the data as described in Section \@ref(ida).
 
@@
 The NLSY79 data contains a large scope of variables but for our aim, the scope required is limited to demographic profiles and wages data.
-More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}, clicked on the direct link to NSLY79 data and navigated as described below.
+More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}, clicked on the direct link to NSLY79 data and navigated as described in Figure \ref{fig:source-nav}.
 
-::: {.tcolorbox data-latex="[label=\"box:data\", title=Navigating the data source to download the raw data]" }
-\faDatabase\ NLSY79 (\url{https://www.nlsinfo.org/investigator/pages/search?s=NLSY79})  
+\begin{figure}[t]
+
+\begin{tcolorbox}[title = Navigating the data source]
+\faDatabase\ NLSY79 (\url{https://www.nlsinfo.org/investigator/pages/search?s=NLSY79})\\
 \vspace{1mm}
-\faCheck\ The CASEID will be always be selected.  
+\faCheck\ The CASEID will be always be selected.  \\
 \vspace{1mm}
-\faCheck\ The 3 recommended demographic variable (sample ID, race and sex) were selected.  
+\faCheck\ The 3 recommended demographic variable (sample ID, race and sex) were selected.  \\
 \vspace{1mm}
@@
 \end{itemize}
-:::
+\end{tcolorbox}
+\caption{The above documents the steps taken to select variables of interest and download the raw data.\label{fig:source-nav}}
+\end{figure}
 

### 2021-09-30 paper/paper.tex
navigation step changed to figure

@@
 Thus, this paper demonstrates the process -- extracting, tidying, cleaning, and exploring with documentation -- to make refreshed data available for education or research.
-Three new data sets, and the code to produce them are provided in an accompanying open source R package, called \texttt{[CENSORED]}.
+Three new data sets, and the code to produce them are provided in an accompanying open source R package, called {[}CENSORED{]}.
 As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
@@
 The NLSY79 data contains a large scope of variables but for our aim, the scope required is limited to demographic profiles and wages data.
-More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}, clicked on the direct link to NSLY79 data and navigated as described below.
+More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}, clicked on the direct link to NSLY79 data and navigated as described in Figure \ref{fig:source-nav}.
 
-\begin{tcolorbox}[label="box:data", title=Navigating the data source to download the raw data]
+\begin{figure}[t]
 
-\faDatabase~NLSY79 (\url{https://www.nlsinfo.org/investigator/pages/search?s=NLSY79})\\
+\begin{tcolorbox}[title = Navigating the data source]
+\faDatabase\ NLSY79 (\url{https://www.nlsinfo.org/investigator/pages/search?s=NLSY79})\\
 \vspace{1mm}
-\faCheck~The CASEID will be always be selected.\\
+\faCheck\ The CASEID will be always be selected.  \\
 \vspace{1mm}
-\faCheck~The 3 recommended demographic variable (sample ID, race and sex) were selected.\\
+\faCheck\ The 3 recommended demographic variable (sample ID, race and sex) were selected.  \\
 \vspace{1mm}
-For the remaining variables, we went to the ``Variable Search'' tab and select variables as follows
-
+For the remaining variables, we went to the "Variable Search" tab and select variables as follows
 \begin{itemize}
@@
 \end{itemize}
-
 \end{tcolorbox}
+\caption{The above documents the steps taken to select variables of interest and download the raw data.\label{fig:source-nav}}
+\end{figure}
 

### 2021-09-30 data-raw/ida.R
Add script for add blinded flowchart.

@@ -356,3 +356,7 @@ include_graphics(here::here("paper/figures/flowchart.png"))
 #       [7]: 'yowie Package'
 #       ",
 #             height = 700)
+
+### ---- flow-chart-blind
+include_graphics(here::here("paper/figures/flowchart_blind.png"))
+

### 2021-09-30 paper/paper-jsdse/paper-jsdse.tex
demog data update

@@
 Subsets of this data can be found in numerous textbooks and research articles.
-However, the steps and decisions taken to get from the raw data to the textbook data is never clearly articulated.
+However, the steps and decisions taken to get from the raw data to the textbook data are never clearly articulated.
 This article describes our journey when trying to re-create a textbook example data set from the original database, with the goal being to refresh the textbook data more regularly.
 Thus, this paper demonstrates the process -- extracting, tidying, cleaning, and exploring with documentation -- to make refreshed data available for education or research.
-Three new data sets, and the code to produce them are provided in an accompanying open source R package, called {[}CENSORED{]}.
-As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
+Three new data sets and the code to produce them are provided in an accompanying open source R package, called {[}CENSORED{]}.
+As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks and providing more convenient samples of the data to potential users.
 \end{abstract}
@@
 
-\begin{figure}[Ht]
+\begin{figure}[t]
 
-\begin{tcolorbox}
+\begin{tcolorbox}[title = Navigating the data source]
 \faDatabase\ NLSY79 (\url{https://www.nlsinfo.org/investigator/pages/search?s=NLSY79})\\
@@
 \end{tcolorbox}
-\caption{Above shows the steps taken to navigate the data source to download the raw data.\label{fig:source-nav}}
+\caption{The above documents the steps taken to select variables of interest and download the raw data.\label{fig:source-nav}}
 \end{figure}
@@
 
-The next step is tidying the highest grade completed (\texttt{hgc}) variable.
-This variable came with several version in each year.
-We choose the revised May data because it seemed to have less missing and presumably has been checked.
+The next step is tidying to obtain \texttt{hgc} and \texttt{yr\_hgc}.
+The highest grade completed are recorded in \texttt{new\_data\_qnames} as variables beginning with \texttt{Q3-4} and \texttt{HGC} with suffix of the year it was recorded.
+In addition the variables beginning with \texttt{HGCREV} contain the revised data.
+We choose to use this revised May data because it seemed to have less missing and presumably has been checked.
 However, there is no revised May data for 2012, 2014, 2016, and 2018, thus, we use the ordinary May data for these years.
-The code for tidying this variable is provided in the supplementary materials.
 
-Furthermore, \texttt{hgc} is measured and could be updated in each period of the survey.
-Hence, we only used the highest grade completed ever and derived the year when it is completed.
-The code for deriving this variable is also available at \url{CENSORED/articles/raw-to-input-data.html}.
+The \texttt{hgc} is measured and could be updated in each period of the survey.
+We chose to only retain the highest grade completed for each individual and derived the year when they completed it (\texttt{yr\_hgc}) by finding the minimum year with the highest completed grade.
 
@@
 
-The next step is that explore the mean hourly wage data. In this case, we only explore the wages data of up to \(12^{th}\) grade. Here, we use visualisation techniques to do IDA.
+The next step is that explore the mean hourly wage data. In this case, we only explore the wages data of up to \(12^{th}\) grade. Here, we use visualization techniques to do IDA.
 
@@
 
+\hypertarget{data-availability-statement}{%
+\section{Data Availability Statement}\label{data-availability-statement}}
+
+The authors confirm that the data supporting the findings of this study are available within the supplementary materials.
+
 \hypertarget{references}{%

### 2021-09-30 paper/paper.Rmd
Edit the flowchart.
demog data update

@@
 
-The next step is tidying the highest grade completed (`hgc`) variable.
-This variable came with several version in each year.
-We choose the revised May data because it seemed to have less missing and presumably has been checked.
+The next step is tidying to obtain `hgc` and `yr_hgc`.
+The highest grade completed are recorded in `new_data_qnames` as variables beginning with `Q3-4` and `HGC` with suffix of the year it was recorded.
+In addition the variables beginning with `HGCREV` contain the revised data.
+We choose to use this revised May data because it seemed to have less missing and presumably has been checked.
 However, there is no revised May data for 2012, 2014, 2016, and 2018, thus, we use the ordinary May data for these years.
-The code for tidying this variable is provided in the supplementary materials.
 
@@
 
-Furthermore, `hgc` is measured and could be updated in each period of the survey.
-Hence, we only used the highest grade completed ever and derived the year when it is completed.
-The code for deriving this variable is also available at \url{`r yowie_pkgdown`/articles/raw-to-input-data.html}.
+The `hgc` is measured and could be updated in each period of the survey.
+We chose to only retain the highest grade completed for each individual and derived the year when they completed it (`yr_hgc`) by finding the minimum year with the highest completed grade.
 

### 2021-09-30 paper/paper.tex
demog data update

@@
 The NLSY79 data contains a large scope of variables but for our aim, the scope required is limited to demographic profiles and wages data.
-More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}, clicked on the direct link to NSLY79 data and navigated as described in Figure \ref{fig:source-nav}.
+More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}, clicked on the direct link to NSLY79 data and navigated as described in \newline\Figure \ref{fig:source-nav}.
 

### 2021-09-30 data-raw/data-preprocessing.R

@@ -102,6 +102,24 @@ full_demographics <- full_join(dob_tidy, demog_tidy, by = "id") %>%
   full_join(highest_year, by = "id") %>%
   mutate(age_1979 = 1979 - (dob_year + 1900))
 
+demog_nlsy79 <- full_demographics %>%
+  mutate(age_1979 = 1979 - (dob_year + 1900)) %>%
+  dplyr::select(id,
+                age_1979,
+                gender,
+                race,
+                hgc,
+                hgc_i,
+                yr_hgc) %>%
+  mutate(id = as.factor(id),
+         age_1979 = as.integer(age_1979),
+         hgc = as.factor(hgc),
+         yr_hgc = as.integer(yr_hgc))
+
+
+## ---- save-demog-data
+usethis::use_data(demog_nlsy79, overwrite = TRUE)
+
 ## ---- tidy-hours
 # make a list for years where we used the "QES-52A"
 year_A <- c(1979:1987, 1993)

### 2021-09-30 paper/paper-jsdse/paper-jsdse.tex
corrected the wages description

@@
 
-The employment data comprises of three variables, i.e., \texttt{total\ hours\ of\ work\ per\ week}, \texttt{number\ of\ jobs\ that\ an\ individual\ has}, and \texttt{mean\ hourly\ wage}.
+Our target variables for the employment are to obtain respondent's mean hourly wage (\texttt{mean\_hourly\_wage}), the number of jobs (\texttt{number\_of\_jobs}) and the total hours of work per week (\texttt{total\_hours}) for each survey year.
+As the data only reports up to 5 jobs for each respondent, the maximum number of jobs is capped at 5.
 
-\texttt{hours\ worked\ per\ week} initially has only one version per job, no choice from 1979 to 1987 (QES-52A).
-From 1988 onward, when we had more options, we chose the variable for total hours, including time spent working from home (QES-52D).
-However, in 1993, this variable did not have all the five D variables (the first one and the last one were missing), so we used QES-52A variable instead.
-In addition, 2008 only had jobs 1-4 for the QES-52D variable (whereas the other years had 1-5), so we use these.
+From 1979 to 1987, \texttt{new\_data\_qnames} only contain one version of hours worked per week for each job (in the variables with names starting with \texttt{QES-52A}).
+From 1988 onward, we selected the total hours worked per week, including hours working from home (\texttt{QES-52D}).
+However, in 1993, this variable was missing for the first and last job so we selected to use \texttt{QES-52A} instead.
+In addition, 2008 only had jobs 1-4 for the \texttt{QES-52D} variable, so we use only these.
 
-The same way was also deployed to tidy the rate of wage by year and by ID.
-The difference is that the hourly rate has only one version of each year.
-The hours of work and the hourly rate are then joined to calculate the number of jobs that a respondent has and their mean hourly wage.
-Some observations have 0 in their hourly rate and work hours, which is considered an invalid value.
-Thus, these observations are set to be NA.
-Besides, there are some observations with unusual hours of work.
-Here, we also censored the observations with \textgreater{} 84 hours of work a week to be NA.
+The hourly wages are in the variables beginning with \texttt{HRP} in \texttt{new\_data\_qnames}.
+As a respondent may have multiple jobs, the \texttt{mean\_hourly\_wage} is computed as a weighted average of the hourly wage for each job with the number of hours worked for each job as weights (provided that the information on number of hours is available); if number of hours worked for any job is missing, then the \texttt{mean\_hourly\_wage} is computed as a simple average of all available hourly wage.
+Prior to computing the mean hourly wage, we undertook a number of steps to treat unusual observations as described below:
 
-Since our ultimate goal is to calculate the mean hourly wage, the number of jobs is calculated based on the availability of the \texttt{rate\_per\_hour} information.
-For example, the number of jobs of ID 1, based on \texttt{hours\_work}, is 2.
-However, since the information of hourly rate of \texttt{job\_02} is not available, the number of jobs is considered as 1.
+\begin{itemize}
+\tightlist
+\item
+  If the hourly rate is recorded as 0, we set wage as missing.
+\item
+  If the total hours of worked for the corresponding job is greater than 84 hours, we set the wage and hour worked as missing.
+\end{itemize}
 
-Further, we calculated the mean hourly wage for each ID in each year using a weighted mean with the hours of work as the weight.
-However, there is a lot of missing value in \texttt{hours\_work} variable.
-In that case, we only calculated the mean hourly wage based on the arithmetic/regular mean method.
-Hence, we created a new variable to flag whether the mean hourly wage is weighted or regular.
-Additionally, if an ID only had one job, we directly used their hourly wages information and flagged it as arithmetic mean.
+The number of jobs (\texttt{number\_of\_jobs}) for each respondent per year is computed from the number of non-missing values of hourly wage \emph{and} hours worked.
+If either the hourly wage or hours worked is missing, we do not tally this.
 
@@
 The employment and demographic variables are then joined.
-We also filtered the data to only have the cohort who completed the education up to 12th grade and participated at least five rounds in the survey and save it to an object called \texttt{wages}.
+This data are further filtered to the cohort who completed the education up to 12th grade and participated at least five rounds in the survey.
+We save the resultant wage data on this cohort as \texttt{wages}.
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/Users/etan0038/yowie/paper/figures/flowchart}
+{\centering \includegraphics[width=0.85\linewidth]{/Users/etan0038/yowie/paper/figures/flowchart_blind}
 
@@
 van der Loo and de Jonge 2021).
-Pink, blue, and green represents the raw, input, and valid data, respectively.}\label{fig:flow-chart}
+Pink, blue, and green represents the raw, input, and valid data, respectively.}\label{fig:flow-chart-blind}
 \end{figure}

### 2021-09-30 paper/paper.Rmd
Add conditional chunk for blinded version.
corrected the wages description

@@
 
-The employment data comprises of three variables, i.e., `total hours of work per week`, `number of jobs that an individual has`, and `mean hourly wage`.
+Our target variables for the employment are to obtain respondent's mean hourly wage (`mean_hourly_wage`), the number of jobs (`number_of_jobs`) and the total hours of work per week (`total_hours`) for each survey year.
+As the data only reports up to 5 jobs for each respondent, the maximum number of jobs is capped at 5.
 
-`hours worked per week` initially has only one version per job, no choice from 1979 to 1987 (QES-52A).
-From 1988 onward, when we had more options, we chose the variable for total hours, including time spent working from home (QES-52D).
-However, in 1993, this variable did not have all the five D variables (the first one and the last one were missing), so we used QES-52A variable instead.
-In addition, 2008 only had jobs 1-4 for the QES-52D variable (whereas the other years had 1-5), so we use these.
+From 1979 to 1987, `new_data_qnames` only contain one version of hours worked per week for each job (in the variables with names starting with `QES-52A`).
+From 1988 onward, we selected the total hours worked per week, including hours working from home (`QES-52D`).
+However, in 1993, this variable was missing for the first and last job so we selected to use `QES-52A` instead.
+In addition, 2008 only had jobs 1-4 for the `QES-52D` variable, so we use only these.
 
@@
 
-The same way was also deployed to tidy the rate of wage by year and by ID.
-The difference is that the hourly rate has only one version of each year.
-The hours of work and the hourly rate are then joined to calculate the number of jobs that a respondent has and their mean hourly wage.
-Some observations have 0 in their hourly rate and work hours, which is considered an invalid value.
-Thus, these observations are set to be NA.
-Besides, there are some observations with unusual hours of work.
-Here, we also censored the observations with > 84 hours of work a week to be NA.
+The hourly wages are in the variables beginning with `HRP` in `new_data_qnames`.
+As a respondent may have multiple jobs, the `mean_hourly_wage` is computed as a weighted average of the hourly wage for each job with the number of hours worked for each job as weights (provided that the information on number of hours is available); if number of hours worked for any job is missing, then the `mean_hourly_wage` is computed as a simple average of all available hourly wage.
+Prior to computing the mean hourly wage, we undertook a number of steps to treat unusual observations as described below:
+
+* If the hourly rate is recorded as 0, we set wage as missing.
+* If the total hours of worked for the corresponding job is greater than 84 hours, we set the wage and hour worked as missing.
+
+The number of jobs (`number_of_jobs`) for each respondent per year is computed from the number of non-missing values of hourly wage *and* hours worked.
+If either the hourly wage or hours worked is missing, we do not tally this.
 
@@
 ```
-
-Since our ultimate goal is to calculate the mean hourly wage, the number of jobs is calculated based on the availability of the `rate_per_hour` information.
-For example, the number of jobs of ID 1, based on `hours_work`, is 2.
-However, since the information of hourly rate of `job_02` is not available, the number of jobs is considered as 1.
-
-Further, we calculated the mean hourly wage for each ID in each year using a weighted mean with the hours of work as the weight.
-However, there is a lot of missing value in `hours_work` variable.
-In that case, we only calculated the mean hourly wage based on the arithmetic/regular mean method.
-Hence, we created a new variable to flag whether the mean hourly wage is weighted or regular.
-Additionally, if an ID only had one job, we directly used their hourly wages information and flagged it as arithmetic mean.
 
@@
 The employment and demographic variables are then joined.
-We also filtered the data to only have the cohort who completed the education up to 12th grade and participated at least five rounds in the survey and save it to an object called `wages`.
+This data are further filtered to the cohort who completed the education up to 12th grade and participated at least five rounds in the survey.
+We save the resultant wage data on this cohort as `wages`.
 
@@
 
-```{r flow-chart, fig.cap="(ref:flow-chart)", out.width="85%"}
+```{r flow-chart-blind, fig.cap="(ref:flow-chart)", out.width="85%", eval = CENSOR}
 ```

### 2021-09-30 paper/paper.tex
corrected the wages description

@@
 The NLSY79 data contains a large scope of variables but for our aim, the scope required is limited to demographic profiles and wages data.
-More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}, clicked on the direct link to NSLY79 data and navigated as described in \newline\Figure \ref{fig:source-nav}.
+More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}, clicked on the direct link to NSLY79 data and navigated as described in Figure \ref{fig:source-nav}.
 

### 2021-09-30 data-raw/data-preprocessing.R

@@ -144,12 +144,11 @@ get_hour <- function(year) {
     rename(id = CASEID_1979)
 }
 
-
-
-## ---- tidy-rate
 # getting the hours of work of all observations
 years <- c(1979:1994, seq(1996, 2018, by = 2))
 hours_all <- map_dfr(years, get_hour)
+
+## ---- tidy-rate
 
 get_rate <- function(year) {
   new_data_qnames %>%
@@ -163,6 +162,8 @@ get_rate <- function(year) {
 }
 
 rates_all <- map_dfr(years, get_rate)
+
+## ---- tidy-rate-hour
 
 # join hours and rates variable
 hours_wages <- left_join(rates_all,

### 2021-09-30 paper/paper-jsdse/paper-jsdse.tex
clean up a bit for IDA part and censored link to input-to-valid html

@@
 To validate this, we used the \texttt{demog\_nlsy79}, i.e., the data with the survey years 1979 sample.
-Table \ref{tab:age-table} and Table \ref{tab:gender-race-table} suggest that the demographic data we had is consistent with the sample information in the database.
+Tables \ref{tab:age-table} and \ref{tab:gender-race-table} suggest that the demographic data we had is consistent with the sample information in the database.
 
@@
 
-The next step is that explore the mean hourly wage data. In this case, we only explore the wages data of up to \(12^{th}\) grade. Here, we use visualization techniques to do IDA.
+In the next step, we explore the mean hourly wage data.
+In this case, we only explore the wages data of respondents that have the highest completed grade of up to \(12^{th}\) grade.
+We employ visualization techniques to perform the IDA as described next.
 
@@
 Thus, the model for each ID is built utilizing the \texttt{nest} and \texttt{map} function from \texttt{tidyr} (Wickham 2020) and \texttt{purrr} (Henry and Wickham 2020), respectively.
+The full code for this is shown at \url{CENSORED/articles/input-to-valid-data.html} but also described in detail next.
 

### 2021-09-30 paper/paper.Rmd
clean up a bit for IDA part and censored link to input-to-valid html

@@
 To validate this, we used the `demog_nlsy79`, i.e., the data with the survey years 1979 sample.
-Table \ref{tab:age-table} and Table \ref{tab:gender-race-table} suggest that the demographic data we had is consistent with the sample information in the database.
+Tables \ref{tab:age-table} and \ref{tab:gender-race-table} suggest that the demographic data we had is consistent with the sample information in the database.
 
@@
 
-The next step is that explore the mean hourly wage data. In this case, we only explore the wages data of up to $12^{th}$ grade.  Here, we use visualization techniques to do IDA.
+In the next step, we explore the mean hourly wage data.
+In this case, we only explore the wages data of respondents that have the highest completed grade of up to $12^{th}$ grade.
+We employ visualization techniques to perform the IDA as described next.
 
@@
 Thus, the model for each ID is built utilizing the `nest` and `map` function from `tidyr` [@tidyr] and `purrr` [@purrr], respectively.
+The full code for this is shown at \url{`r yowie_pkgdown`/articles/input-to-valid-data.html} but also described in detail next.
 
@@
 ```{r flow-chart-blind, fig.cap="(ref:flow-chart)", out.width="85%", eval = CENSOR}
+```
+
+```{r flow-chart, fig.cap="(ref:flow-chart)", out.width="85%", eval = !CENSOR}
 ```

### 2022-01-12 paper/paper.Rmd
Adding reviewers comment to text, makes change to grammatical errors.

@@
 However, open data is often referred to as what we might call "wild data" because it requires substantial cleaning and tidying to tame it into textbook shape.
-@HuebnerMariannePhD2016Asat emphasize that making the data cleaning process accountable and transparent is imperative and essential for the integrity of downstream statistical analyses and model building [@HuebnerMarianne2020Haar].
+@HuebnerMariannePhD2016Asat emphasize that making the data cleaning process accountable and transparent is imperative and essential for the integrity of downstream statistical analyses and model building <!-- AE comment: It's unclear from context what is being referred to from the Huebner et al 2020 citation -->  [@HuebnerMarianne2020Haar].
 Data cleaning can be considered to be a part of what is called "initial data analysis" (IDA) [@Chatfield1985TIEo].
@@
 Here, we investigate the process of getting from the raw NLSY79 data to a textbook data set as similar as possible to that provided by @SingerJudithD2003Alda.
-We should also note that race is a variable in the original data set, and for compatibility it is also provided with the refreshed data, for the purposes of studying racism, not race [@racismnotrace],  <!--However, we cannot create the exact same data set as published in their book since we do not have the information of what age threshold they used to determine the high school dropouts.-->
+We should also note that race is a variable in the original data set, and for compatibility it is also provided with the refreshed data, for the purposes of studying racism, not race [@racismnotrace].
+<!--However, we cannot create the exact same data set as published in their book since we do not have the information of what age threshold they used to determine the high school dropouts.-->
 
@@
 The data is first organised into "tidy data" [@WickhamHadley2014TD] and then further wrangled using the data pipeline and split-apply-combine approach [@plyr].
-(Tidy data shouldn't be confused with "tame data" which @tamedata coined to refer to text book data sets suitable for teaching, particularly teaching statistics.) The resulting (tame) data is provided in a new R package called `yowie` which includes the code so that the process is reproducible, and could be used to further refresh the data as new records are made available in the NLSY79 database.
+<!--AE comment: this is an important point that shouldn't be relegated to a parenthetical statement.
+--> (Tidy data shouldn't be confused with "tame data" which @tamedata coined to refer to text book data sets suitable for teaching, particularly teaching statistics.) The resulting (tame) data is provided in a new R package called `yowie` which includes the code so that the process is reproducible, and could be used to further refresh the data as new records are made available in the NLSY79 database.
 
@@
 Section \@ref(cleaning) presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and IDA to find and repair anomalies.
-Our final subset is compared the old textbook subset in Section \@ref(compare).
+Our final subset is compared to the old textbook subset in Section \@ref(compare).
 Finally, Section \@ref(summary) summarizes the contribution and makes recommendations for the NLSY79 data curators.
@@
 
-@LooMarkvander2018Sdcw describe the notion of a "statistical value chain" where the production stages of data cleaning process are earmarked as raw data (data as arrived to the desk of an analyst), input data (data organised with correct type and identified variables) and valid data (data that faithfully represent the variables).
+@LooMarkvander2018Sdcw describe the notion of a "statistical value chain" where the production stages of the data cleaning process are earmarked as raw data (data as arrived to the desk of an analyst), input data (data organised with correct type and identified variables) and valid data (data that faithfully represent the variables).
 In this section, we outline the steps to download the raw data (Section \@ref(getdata)) and then tidying the raw data into input data, specifically for the demographic variables (Section \@ref(tidydemog)) and the employment variables (Section \@ref(tidyemp)), so that the resulting input data can be used downstream for validating the data as described in Section \@ref(ida).
@@
 We only alter the file path in `NLSY79.R` and run the script without any other alteration.
-This results in an initial processing of the raw data into two data set,  `categories_qnames` (where the observations are stored in categorical/interval values) and `new_data_qnames` (the observations are stored in integer form).
+This results in an initial processing of the raw data into two data sets,  `categories_qnames` (where the observations are stored in categorical/interval values) and `new_data_qnames` (the observations are stored in integer form).
 
@@
 According to @WickhamHadley2014TD, tidy data sets comply with three rules: (i) each variable forms a column, (ii) each observation forms a row, and (iii) each type of observational unit forms a table.
-The raw data does not comply these rules, thus we re-arrange and wrangle the data into tidy data form, columns corresponding to individual ID, year, job number, wage in dollars and the demographic variables.
-This is done using the `tidyverse` suite of packages [@tidyverse], `tidyr` [@tidyr] to pivot the data into long form, `dplyr` [@dplyr], and `stringr` [@stringr] for creating new variables, and levels of factors by text wrangling.
+The raw data does not comply with these rules, thus we re-arrange and wrangle the data into tidy data form, columns corresponding to individual ID, year, job number, wage in dollars and the demographic variables.<!-- AE comment: This sentence is unclear.
+What is dplyr used for?
+Are there tidyverse packages used but not mentioned?
+If not, splitting this sentence into two may make it clearer which package is used to what purpose.
+Also, I haven't tried it in code, but I imagine that the data could be tidied as described just using pivot_longer, at least for the job number, year, and wage data.
+In other words, I don't see why dplyr or stringr would be needed for the data described on page 6.--> This is done using the `tidyverse` suite of packages [@tidyverse], `tidyr` [@tidyr] to pivot the data into long form, `dplyr` [@dplyr], and `stringr` [@stringr] for creating new variables, and levels of factors by text wrangling.
 The long form of the data makes it possible to do these data transformations efficiently, and it is an intermediate step towards the final target data.
@@
 The hourly wages are in the variables beginning with `HRP` in `new_data_qnames`.
-As a respondent may have multiple jobs, the `mean_hourly_wage` is computed as a weighted average of the hourly wage for each job with the number of hours worked for each job as weights (provided that the information on number of hours is available); if number of hours worked for any job is missing, then the `mean_hourly_wage` is computed as a simple average of all available hourly wage.
+As a respondent may have multiple jobs, the `mean_hourly_wage` is computed as a weighted average of the hourly wage for each job with the number of hours worked for each job as weights (provided that the information on number of hours is available); if number of hours worked for any job is missing, then the `mean_hourly_wage` is computed as a simple average of all available hourly wages.
 Prior to computing the mean hourly wage, we undertook a number of steps to treat unusual observations as described below:
@@
 If either the hourly wage or hours worked is missing, we do not tally this.
+<!-- AE comment: "If either the hourly wage or hours worked is missing, we do not tally this." I take "this" to mean "number_of_jobs".
+But in row 2, total_hours is missing, yet number_of_jobs is 1.
+The number of jobs (1) was tallied even though hours worked was missing.
+-->
 
@@
 The employment and demographic variables are then joined.
-This data are further filtered to the cohort who completed the education up to 12th grade and participated at least five rounds in the survey.
+These data are further filtered to the cohort who completed the education up to 12th grade and participated in at least five rounds in the survey.
 We save the resultant wage data on this cohort as `wages`.
@@
 In this paper, we conduct an IDA or a preliminary data analysis to assess the consistency of the data with the cohort information that the NLSY provides.
-In addition, we also aim to find the anomaly in the wages values using this approach.
-We mainly use graphical summaries to do the IDA using `ggplot2` [@ggplot2] and `brolgar` [@brolgar].
+<!-- AE comments: This comes as a surprise to the reader.
+What anomaly?
+The wage anomalies should be mentioned in the introduction so as not to be a surprise at this point in the manuscript.
+ -->In addition, we also aim to find the anomaly in the wages values using this approach. We mainly use graphical summaries to do the IDA using `ggplot2` [@ggplot2] and `brolgar` [@brolgar].
 
@@
 To validate this, we used the `demog_nlsy79`, i.e., the data with the survey years 1979 sample.
-Tables \ref{tab:age-table} and \ref{tab:gender-race-table} suggest that the demographic data we had is consistent with the sample information in the database.
+<!--AE comment:  I don't know that Table 1 is necessary as I don't have anything to compare it to.
+If the numbers check out, a simple statement to that effect would suffice.
+Also, what does "(?CHECK)" mean?
+This needs to be resolved.--> Tables \ref{tab:age-table} and \ref{tab:gender-race-table} suggest that the demographic data we had is consistent with the sample information in the database.
 
@@
 
-We take 36 samples randomly from the data and plot them, as shown in Figure \ref{fig:sample-plot}.
-It implies that these respondents have a lot of variability in wages, for example, the IDs in panel numbers 5, 7, and 11.
+We take 20 samples randomly from the data and plot them, as shown in Figure \ref{fig:sample-plot}.<!--AE comment: "It" implies...
+Please specify what "it" refers to.
+This could be done by adding onto the first sentence...
+"as shown in Figure 2, which shows that some respondents have high variability in wages, particularly 5, 7, and 11." However, 7 and 11 don't seem to have particularly high variability, especially when compared to 18.
+This whole paragraph needs to be fixed/updated.--> It shows that these respondents have a lot of variability in wages, for example, the IDs in panel numbers 5, 7, and 11.
 The plot also implies that the samples have a different pattern of mean hourly wages.
-Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
+Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, <!--AE comment: Not panel 9 --> the IDs in panel 9.
 However, when checking the summary plots (Figure \ref{fig:feature-plot}), we found that some observations had exceptionally high wages.
@@
 
-The anomalies are also found in the total hours of work, where some observations work for 420 hours a week in total.
+<!--AE comment: Rewrite this sentence.
+"Observations" aren't working 420 hours.
+A "respondent" might have been reported as having worked 420 hours per week.--> The anomalies are also found in the total hours of work, where some observations work for 420 hours a week in total.
 According to @MichaelRPergamit2001DWTN, one of the flaws of the NLSY79 employment data is that the NLSY79 collects the information of the working hours since the last interview.
-Thus, it might be challenging for the respondents to track the within-job hours' changes between survey years, especially for the respondents with fluctuated working hours or seasonal jobs.
+Thus, it might be challenging for the respondents to track the within-job hours' changes between survey years, especially for the respondents with fluctuating working hours or seasonal jobs.
 It even has been more challenging since 1994, where the respondents had to recall two years periods.
@@
 
-As part of the IDA, which is the model formulation, we build a robust linear regression model to treat the extreme values in the data.
+As part of the IDA, <!--Awkward sentence.
+"which is the model formulation"?--> which is the model formulation, we build a robust linear regression model to treat the extreme values in the data.
 Robust linear regression yields an estimation robust to the influence of noise or contamination [@KollerManuel2016rARP].
@@
 
-Figure \ref{fig:flow-chart} summarizes the steps taken to go from raw to input to valid data [@LooMarkvander2018Sdcw] to create a refreshed wages data set.
+Figure \ref{fig:flow-chart-blind} summarizes the steps taken to go from raw to input to valid data [@LooMarkvander2018Sdcw] to create a refreshed wages data set.
 
@@
 Now is the time to see how close we have come to the original textbook data.
-The original set contains wages of high school dropouts wages [@SingerJudithD2003Alda] from 1979 through to 1994.
-The refreshed data set for comparison is also wages on high school drop outs, over 1979 to 2018.
+The original set contains wages of high school dropouts [@SingerJudithD2003Alda] from 1979 through to 1994.
+The refreshed data set for comparison is also wages on high school drop outs from 1979 to 2018.
 The original set is available in the R package `brolgar` and the refreshed data is available in the R package `r yowie`.
 
+<!--what does the natural log have to do with the y-axis?
+-->
+
 ```{r plotting-sw-do, fig.cap = "Comparison of original textbook example (A) with refreshed data (B).
-The original data was inflation-adjusted to 1990 prices and the individual's time of collection was converted to a length of experience in the workforce, which make it difficult to precisely compare the two sets.", fig.width = 9, fig.height = 5, out.width = "100%", cache = FALSE}
+The original data was inflation-adjusted to 1990 prices and the individual's time of collection was converted to a length of experience in the workforce, which makes it difficult to precisely compare the two sets.", fig.width = 9, fig.height = 5, out.width = "100%", cache = FALSE}
 ```
@@
 On education, in the refreshed data there are very few individuals with less than a 12th grade education, which varies from the original.
-In the original data there were 366 individuals with up to 8th grade and 522 9th-12th grade, as compared to 967 with 12th grade in the refreshed data.
+<!--AE comment: How many individuals up to 8th grade are in the refreshed dataset?
+What are plausible reasons for these differences?--> In the original data there were 366 individuals with up to 8th grade and 522 9th-12th grade, as compared to 967 with 12th grade in the refreshed data.
 
@@
 - there is no explicit variable in the database recording high school dropout, which means we needed to compare date of 12th grade with the individual's age.
-- calculating experience in the workforce, would require knowing the time of first job within the first year the individual was recorded.
+- calculating experience in the workforce would require knowing the time of first job within the first year the individual was recorded.
 

### 2022-01-13 paper/paper.Rmd
Add reviewer 2's comments, fix the copy-edited notes.

@@
 This paper demonstrates the steps of IDA, and documents the process, for a prominent open data source, National Longitudinal Survey of Youth 1979, henceforth referred to NLSY79.
-This data has been playing an important role in research in various disciplines, including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century [@MichaelRPergamit2001DWTN].
+This data has been playing an important role in research in various disciplines, including but not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century [@MichaelRPergamit2001DWTN].
 In addition, this is considered a carefully designed longitudinal survey with high retention rates, making it suitable for life course research [@MichaelRPergamit2001DWTN; @eliznlsy].
@@
 @SingerJudithD2003Alda used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling of wages on workforce experience, with covariates education and race.
+<!--AE comment: For example, I would like more discussion about the textbook selected as the exemplar.
+It sounds like this dataset is used in many books, why choose that one?
+The text should include the name of the book, so readers do not need to flip to the citations in order to learn what it was.
+From the name, I am guessing that the book focuses heavily on this dataset, which is why it makes sense to reproduce their subset.
+Be explicit about this.-->
 Our aim is to refresh this textbook data with data from 1994 through to the latest data reported in 2018, a purpose that is consistent with @grimshaw's statistics education goal of embracing authentic data experiences.
@@
 
-Our process of cleaning builds heavily on the tidyverse approach [@tidyverse].
+Our process of cleaning builds heavily on the `tidyverse` approach [@tidyverse].
 The data is first organised into "tidy data" [@WickhamHadley2014TD] and then further wrangled using the data pipeline and split-apply-combine approach [@plyr].
 <!--AE comment: this is an important point that shouldn't be relegated to a parenthetical statement.
---> (Tidy data shouldn't be confused with "tame data" which @tamedata coined to refer to text book data sets suitable for teaching, particularly teaching statistics.) The resulting (tame) data is provided in a new R package called `yowie` which includes the code so that the process is reproducible, and could be used to further refresh the data as new records are made available in the NLSY79 database.
+--> (Tidy data shouldn't be confused with "tame data" which @tamedata coined to refer to textbook data sets suitable for teaching, particularly teaching statistics.) The resulting (tame) data is provided in a new R package called `yowie` which includes the code so that the process is reproducible, and could be used to further refresh the data as new records are made available in the NLSY79 database.
 
@@
 
-The NLSY79 is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth born between 1957-1964 [@nlsy79].
+The NLSY79 is a longitudinal survey administered by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth born between 1957-1964 [@nlsy79].
 The cohort originally included 12,686 respondents aged 14-22 when first interviewed in 1979.
-It comprised of Blacks, Hispanics, economically disadvantaged non-Black non-Hispanics, and youth in the military.
-In 1984 and 1990, two sub-samples were dropped from the interview; the dropped subjects are the 1,079 members of the military sample and 1,643 members of the economically disadvantaged non-Black non-Hispanics, respectively.
+<!--AE comment: “It comprised of Blacks, Hispanics, economically disadvantaged non-Black non-Hispanics, and youth in the military.” Definitely needs to remove the term “Blacks” as that is considered to be offensive.
+Perhaps “It comprised of people sampled because they were Black, Hispanic, economically disadvantaged, or in the military.” --> It comprised of Blacks, Hispanics, economically disadvantaged non-Black non-Hispanics, and youth in the military.
+<!--AE comment: “1,643 members of the economically disadvantaged non-Black non-Hispanics, respectively” could perhaps be rephrased to “1,643 members who were selected because of their economic disadvantage, but were not Black or Hispanic.”-->In 1984 and 1990, two sub-samples were dropped from the interview; the dropped subjects were the 1,079 members of the military sample and 1,643 members of the economically disadvantaged non-Black non-Hispanics, respectively.
 Hence, 9,964 respondents remain in the eligible samples.
@@
 
+<!--AE comment: I found the semicolons distracting. Perhaps replace them with commas and use parenthetical for sub-categories, or use a bulleted list.-->
 Although the main focus area of the NLSY is labor and employment, the NLSY also covers several other topics, including education; training and achievement; household, geography, and contextual variables; dating, marriage, cohabitation; sexual activity, pregnancy, and fertility; children; income, assets and program participation; health; attitudes and expectations; and crime and substance use.
 
-There are two ways to conduct the interview of the NLSY79, which are face-to-face interviews or by telephone.
+There are two ways to conduct the interview of the NLSY79, which are face-to-face or by telephone interviews.
 In recent survey years, more than 90 percent of respondents were interviewed by telephone [@eliznlsy].
@@
 
+<!--AE comment: What does “yearly mean hourly wages with years of workforce experience” mean? Is that one variable or two?-->
+<!--AE comment: Was your goal to collect data from 1994-2018 or 1979-2018?
+Page 4 states 1994 onward, but from the figures later it appears that you collected data starting in 1979 (which makes more sense to me).-->
+<!--AE comment: “The NLSY79 data used in Singer and Willett (2003) contains the longitudinal measurements on yearly mean hourly wages with years of workforce experience, and demographic variables education and race, from 1979 through to 1994.” Might work better split into two sentences.
+-->
 The NLSY79 data used in @SingerJudithD2003Alda contains the longitudinal measurements on yearly mean hourly wages with years of workforce experience, and demographic variables education and race, from 1979 through to 1994.
 In addition, the cohort is restricted to male high-school dropouts who first participated in the study at age 14-17 years.
-Thus the target data set is to collect the same variables for the extended time frame of 1994 through to 2018, the most recent year reported.
+<!--AE comment: Thus the target of our study is to collect the same variables for the same time period and extending through to 2018, the most recent year reported.-->Thus the target data set is to collect the same variables for the extended time frame of 1994 through to 2018, the most recent year reported.
 
@@
 @LooMarkvander2018Sdcw describe the notion of a "statistical value chain" where the production stages of the data cleaning process are earmarked as raw data (data as arrived to the desk of an analyst), input data (data organised with correct type and identified variables) and valid data (data that faithfully represent the variables).
-In this section, we outline the steps to download the raw data (Section \@ref(getdata)) and then tidying the raw data into input data, specifically for the demographic variables (Section \@ref(tidydemog)) and the employment variables (Section \@ref(tidyemp)), so that the resulting input data can be used downstream for validating the data as described in Section \@ref(ida).
+In this section, we outline the steps to download the raw data (Section \@ref(getdata)) and then tidy the raw data into input data, specifically for the demographic variables (Section \@ref(tidydemog)) and the employment variables (Section \@ref(tidyemp)), so that the resulting input data can be used downstream for validating the data as described in Section \@ref(ida).
 
@@
 
-The NLSY79 data contains a large scope of variables but for our aim, the scope required is limited to demographic profiles and wages data.
+The NLSY79 data contains a large number of variables but for our aim, the scope required is limited to demographic profiles and wages data.
 More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}, clicked on the direct link to NSLY79 data and navigated as described in Figure \ref{fig:source-nav}.
@@
 
-- `NLSY79.csv`: csv format of the response data,
-- `NLSY79.dat`: dat format of the response data,
+- `NLSY79.csv`: Comma Separated Value format of the response data,
+- `NLSY79.dat`: .dat format of the response data,
 - `NLSY79.NLSY79`: Tagset of variables that can be uploaded to the website to recreate the data set, and
@@
 
-We only alter the file path in `NLSY79.R` and run the script without any other alteration.
+We alter only the file path in `NLSY79.R` and run the script without any other alteration.
 This results in an initial processing of the raw data into two data sets,  `categories_qnames` (where the observations are stored in categorical/interval values) and `new_data_qnames` (the observations are stored in integer form).
@@
 The record contains two versions of birth year and birth month as the survey recorded this in 1979 and 1981.
-We checked for consistency for the two versions and found no discrepancy where the responses were recorded in both 1979 and 1981.
+We checked for consistency between the two versions and found no discrepancy where the responses were recorded in both 1979 and 1981.
 The age was then calculated using the birth year.
@@
 
+<!--AE comment: I was confused by the following section: “We choose to use this revised May data because it seemed to have less missing and presumably has been checked.
+However, there is no revised May data for 2012, 2014, 2016, and 2018, thus, we use the ordinary May data for these years.” I was assuming the data was collected on a yearly basis, but this section suggests it was collected each month.
+Please clarify somewhere in the paper.
+-->
 The next step is tidying to obtain `hgc` and `yr_hgc`.
@@
 
-From 1979 to 1987, `new_data_qnames` only contain one version of hours worked per week for each job (in the variables with names starting with `QES-52A`).
+From 1979 to 1987, `new_data_qnames` only contains one version of hours worked per week for each job (in the variables with names starting with `QES-52A`).
 From 1988 onward, we selected the total hours worked per week, including hours working from home (`QES-52D`).
@@
 The employment and demographic variables are then joined.
-These data are further filtered to the cohort who completed the education up to 12th grade and participated in at least five rounds in the survey.
+These data are further filtered to the cohort who completed education up to 12th grade and participated in at least five rounds in the survey.
 We save the resultant wage data on this cohort as `wages`.
@@
 Also, what does "(?CHECK)" mean?
-This needs to be resolved.--> Tables \ref{tab:age-table} and \ref{tab:gender-race-table} suggest that the demographic data we had is consistent with the sample information in the database.
+This needs to be resolved.-->
+<!--AE comment: Table 1 includes the note ?CHECK which I assume was a note to self. -->
+Tables \ref{tab:age-table} and \ref{tab:gender-race-table} suggest that the demographic data we had is consistent with the sample information in the database.
 
@@
 ```
+
+<!--AE comment: Table 2 seems to have more precision than is necessary.
+Certainly 100% does not need to be reported as 100.00%, and probably all the percentages could be rounded to the nearest 1 percent.
+The caption could explain this, as there might be some rounding errors.
+-->
 
@@
 
+<!--AE comment: Figure 2’s caption should include something about how that data is the raw values from the survey, before inputing anomalous values so the figure and caption can stand alone.-->
+
 ```{r sample-plot, echo = FALSE, fig.cap="Longitudinal profiles of wages for a random sample of 20 individuals in the refreshed data.
@@
 
-We take 20 samples randomly from the data and plot them, as shown in Figure \ref{fig:sample-plot}.<!--AE comment: "It" implies...
+<!--AE comment: With the IDA about wage data, I would appreciate a bit more context.
+Why sample 36 observations?
+I am not a longitudinal data expert, so I was a bit surprised when I flipped to the plot and found it to be small-multiple time series, although upon further consideration it makes sense.
+In that section it would be good to describe what the figure shows in a bit more detail.
+Perhaps “We randomly sample 36 respondents from the data, and plot their average wage as a time series.
+Looking at the patterns over time, we see a lot of variability in wages.
+For example, the people shown in panels 5, 7, and 11 have [explain what is interesting here, again I’m not an expert].” The next sentence here, “Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.” does not seem to provide much additional information.
+What were you looking for here?
+Overall trends over time, increasing/decreasing?
+Any places where one year’s wage looked really different?-->
+
+We randomly take 20 samples from the data and plot them, as shown in Figure \ref{fig:sample-plot}.<!--AE comment: "It" implies...
 Please specify what "it" refers to.
@@
 Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, <!--AE comment: Not panel 9 --> the IDs in panel 9.
-However, when checking the summary plots (Figure \ref{fig:feature-plot}), we found that some observations had exceptionally high wages.
+<!--I think there is a narrative step missing here, which is that looking at those samples led you to look at plots that show values for the wage variable over all participants and all years.
+The phrase “the summary plots” was not enough description for me to know what to expect when I flipped to Figure 3.
+Please be explicit about what the three plots were, why they were made, and what you were looking for.
+I think there is another narrative step missing about Figure 3C, which is that after you saw there were outliers you tried to identify which respondents had the extreme values.
+Again, please spell this out in the text.
+When you return to the plots for the entire dataset in Figure 5, please make parallel comments.
+-->However, when checking the summary plots (Figure \ref{fig:feature-plot}), we found that some observations had exceptionally high wages.
 Some of them, for example respondents in Figure \ref {fig:feature-plot} C had experienced unusual wages only in certain year.
+
+<!--For both Figure 3 and Figure 5, I would prefer the letter label before the description, but that’s a matter of personal preference.-->
 
@@
 
+<!--The next paragraph, which begins “The anomalies are also found” should perhaps be “Similar anomalies were found.” Was it the same respondents with the extremely high wages that showed extremely high number of hours of work?
+I suspect not, but please be explicit.-->
 <!--AE comment: Rewrite this sentence.
 "Observations" aren't working 420 hours.
-A "respondent" might have been reported as having worked 420 hours per week.--> The anomalies are also found in the total hours of work, where some observations work for 420 hours a week in total.
+A "respondent" might have been reported as having worked 420 hours per week.--> The anomalies are also found in the total hours of work, where some observations reported as having worked for 420 hours a week in total.
 According to @MichaelRPergamit2001DWTN, one of the flaws of the NLSY79 employment data is that the NLSY79 collects the information of the working hours since the last interview.
 Thus, it might be challenging for the respondents to track the within-job hours' changes between survey years, especially for the respondents with fluctuating working hours or seasonal jobs.
-It even has been more challenging since 1994, where the respondents had to recall two years periods.
+It even has been more challenging since 1994, where the respondents had to recall two year periods.
 This shortcoming might also contribute to the fluctuation of one's wages data.
@@
 
-As part of the IDA, <!--Awkward sentence.
+<!--AE comment: Section 3.2.1 begins “As part of the IDA, which is the model formulation, we build a robust linear regression model to treat the extreme values in the data.” This could use clarification.
+In 3.2, you said that part of IDA is “ model formulation without any formal statistical inference.” To me, that sounds like using modeling as a descriptive technique, and it’s something I would consider to be more a part of EDA.
+But then in 3.2.1, while it is perhaps not using modeling in an inferential way, the model is being used for more than just description.
+I might rephrase the first sentence as follows: “In order to treat the extreme values in the data, we built a robust linear regression model.” -->
+As part of the IDA, <!--AE comment: Awkward sentence.
 "which is the model formulation"?--> which is the model formulation, we build a robust linear regression model to treat the extreme values in the data.
@@
 However, this method is too computationally and memory expensive, especially for a large data set, like the NLSY79 data.
-Thus, the model for each ID is built utilizing the `nest` and `map` function from `tidyr` [@tidyr] and `purrr` [@purrr], respectively.
+<!--I would pull the tidy and purrr citations to the end of the sentence to make it read more smoothly-->Therefore, the model for each ID is built utilizing the `nest` and `map` function from `tidyr` [@tidyr] and `purrr` [@purrr], respectively.
 The full code for this is shown at \url{`r yowie_pkgdown`/articles/input-to-valid-data.html} but also described in detail next.
@@
 Furthermore, we use M-Estimation with Huber weighting, where the observation with a slight residual gets a weight of 1, while the larger the residual, the smaller the weight (less than 1) [@rlm].
-However, the challenging part of detecting the anomaly using the robustness weight is determining the weight threshold in which the observations are considered outliers.
+<!--AE comment: “However, the challenging part of detecting the anomaly using the robustness weight is determining the weight threshold in which the observations are considered outliers.” I don’t think this sentence needs a “However” and probably should say “past which” rather than “in which”-->However, the challenging part of detecting the anomaly using the robustness weight is determining the weight threshold in which the observations are considered outliers.
 Moreover, it should be noted that not all the outliers are due to an error.
 Instead, it might be that one had reasonably increasing or decreasing wages in a particular period.
+
+<!--AE comment: At the bottom of page 11, where you describe the model formation, more detail would be appreciated.
+You are modeling the mean hourly wage based on year.
+Why?
+Is this standard practice for robust linear regression?
+What is a “slight” residual?
+Is it defined based on absolute magnitude, or standard deviations, or something else?
+Later, you say “To minimize the risk of mistakenly identifying an outlier as an “erroneous outlier”.” Is an erroneous outlier a technical term?
+If not, I think you are trying not to identify erroneous outliers, not mistakenly identifying outliers as erroneous outliers (that’s a double negative).
+The value of 0.12 feels very specific and out of place in a paper that has had few numerals.
+Is this a number an expert in RLR would be able to understand?
+Again, is that a raw value?
+Is it a standard deviation?
+Is there literature about how to choose a threshold?
+Overall, the sentence “We find that 0.12 is the most reasonable value to be the threshold to minimize that drawback’s risk because it still captures the sensible spikes in the data.” is vague and should be expanded upon.
+What is “that drawback’s risk”?
+What are “the sensible spikes in the data?” Be more explicit.-->
 
@@
 In Figure \ref{fig:fixed-feature-plot} B, we plot the three features of mean hourly wages, namely the minimum, median, and maximum value.
+<!--AE comment: What does “It implies that the fluctuation can still be observed in the data after the treatment.” mean?
+I suspect this sentence should be replaced with something along the lines of “The figure shows that fluctuations in wages still exist even in the imputed data.”-->
 We still can see some extreme values in maximum wages, but consider it as a natural variability of the data.
@@
 
-(ref:flow-chart) The stages of data cleaning from the raw data to get three datasets contained `r yowie`.
+(ref:flow-chart) The stages of data cleaning from the raw data to get three datasets contained in `r yowie`.
 "# of individuals" means the number of respondents included in each stage, while "# of observations" means the number of rows in the data.
 The color represents the stage of data cleaning in statistical value chain [@validate].
-Pink, blue, and green represents the raw, input, and valid data, respectively.
+Pink, blue, and green represent the raw, input, and valid data, respectively.
 
@@
 
+<!--AE comment: first sentence can be removed or rephrased.
+Next sentence includes the word wages too many times.
+Next sentence should say “dropouts” rather than “drop outs” to be consistent with the previous sentence.
+-->
 Now is the time to see how close we have come to the original textbook data.
@@
 
-<!--what does the natural log have to do with the y-axis?
--->
+<!--AE comment: what does the natural log have to do with the y-axis?-->
+<!--AE comment: For Figure 7, why did you choose to show the entire dataset here? It would be easier to make a comparison if the plot was cut off at 1994 on the x-axis.-->
 
@@
 
+<!--AE comment: In Section 4, you talk about the variables in the original data.
+“One would expect that there is a record of the day the individual first started a job, and this is used to adjust the year of collection.” I’m not sure what this means.
+Please clarify.-->
+<!--AE comment: “experience in the workforce” should have the period inside the quotations.-->
 There are two aspects of the original data that make a direct comparison difficult.
@@
 
+<!--AE comment: “The treatment of unlikely wages differ in the refreshed data.” I think this means “The treatment of anomalous wages differs between the original and refreshed data sets.” Then you say “We opted to use the weights from a robust linear regression to determine what should be set as missing value as described in Section 3.2.1.” I don’t think you were setting missing values though, were you?
+It was used to determine which values should be imputed, correct?
+Either way, please clarify.-->
 The treatment of unlikely wages differ in the refreshed data.
@@
 This suggests some individuals were removed in the original data.
-In addition, in the original set the racial break down was 246 black, 204 hispanic and 438 white, while in the refreshed data there are 346 black, 219 hispanic and 623 white, so the proportions have not been exactly reproduced.
-On education, in the refreshed data there are very few individuals with less than a 12th grade education, which varies from the original.
+In addition, in the original set the racial breakdown was 246 Black, 204 Hispanic and 438 White participants, while in the refreshed data there are 346 Black, 219 Hispanic, and 623 White participants, so the proportions have not been exactly reproduced.
+On education, in the refreshed data there are very few individuals with less than a 12th grade education, which is different from the original.
 <!--AE comment: How many individuals up to 8th grade are in the refreshed dataset?
@@
 In the first stage, we showed the steps performed to get the data from the NLSY79 database.
-The data format needed converting to tidy format, and this was described.
+The data format needed conversion to tidy format, and this was described.
 After that, an initial data analysis was conducted to investigate and screen the quality of the data.
@@
 The current subset is made available in a new R package, called `r yowie`.
+
+<!--AE comment: The bulleted difficulties in Section 5 are a bit jarring.
+Make them parallel (i.e.
+every bullet should start with a verb -ing so it is determining, calculating, etc) or turn them into a paragraph.
+I liked the specificity of $30,000/hour but it would be better up in the section where you discuss the anomalous values.-->
 
@@
 
+<!--AE comment: At the end of page 14, it would be good to explicitly say that if people don’t like your decisions, they can grab the code and adjust the parameters themselves!
+It’s implied, but should be made explicit.
+-->
 Some readers may disagree with our decisions made to produce the refreshed textbook data and may have better insight than us in producing a more appropriate textbook data.
@@
 The problem with many of the wages records is that there are implausible values, or confusion on how to record wages for multiple jobs.
-These values could be validated with simple checks at data entry.
+These values can be validated with simple checks at data entry.
 Providing an open data resource also is accompanied with the responsibility that the data, especially data that is as valuable as this, is reliable.
@@
 The entire analysis is conducted using `R` [@R] in RStudio IDE using these packages: `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tidyr` [@tidyr], `stringr` [@stringr], `purrr` [@purrr], `brolgar` [@brolgar], `patchwork` [@patchwork], `kableExtra` [@kableExtra], `MASS` [@mass], `janitor` [@janitor], and `tsibble` [@tsibble].
-The paper are generated using `knitr` [@knitr] and `rmarkdown` [@rmarkdown].
+The paper was generated using `knitr` [@knitr] and `rmarkdown` [@rmarkdown].
 
@@
 
-- **Codes**: R script to reproduce data tidying and cleaning are available in this [page](`r yowie_pkgdown`/articles/process-data.html).
+- **Codes**: R script to reproduce data tidying and cleaning is available in this [page](`r yowie_pkgdown`/articles/process-data.html).
 
 - **R Package**:`r yowie` is a data container R package that contains 3 datasets, namely the high school mean hourly wage data, high school dropouts mean hourly wage data, and demographic data of the NLSY79 cohort.
-This package could be accessed [here](`r yowie_repo`).
+This package can be accessed [here](`r yowie_repo`).
 
 - **shiny app**: An interactive `shiny` web app to visualise the effect of selecting different weight threshold for substituting the wages data to its predicted value from a fit of the robust linear regression model.
-This app could be accessed [here](`r yowie_shiny`) with the source code provided [here](`r yowie_repo`/tree/master/app).
+This app can be accessed [here](`r yowie_shiny`) with the source code provided [here](`r yowie_repo`/tree/master/app).
 

### 2022-01-28 paper/paper-jsdse/paper-jsdse.tex
Added work experience variable.

@@
 However, open data is often referred to as what we might call ``wild data'' because it requires substantial cleaning and tidying to tame it into textbook shape.
-Huebner Marianne, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative and essential for the integrity of downstream statistical analyses and model building (M.
-Huebner et al.
+Huebner, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative and essential for the integrity of downstream statistical analyses and model building (Huebner et al.
 2020).
@@
 There are few research papers that document the data cleaning (Wickham 2014).
-Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way and is only shared among restricted parties (M.
-Huebner et al.
+Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way and is only shared among restricted parties (Huebner et al.
 2020).
@@
 Here, we investigate the process of getting from the raw NLSY79 data to a textbook data set as similar as possible to that provided by Singer and Willett (2003).
-We should also note that race is a variable in the original data set, and for compatibility it is also provided with the refreshed data, for the purposes of studying racism, not race (Fullilove 1998),
+We should also note that race is a variable in the original data set, and for compatibility it is also provided with the refreshed data, for the purposes of studying racism, not race (Fullilove 1998).
 
-Our process of cleaning builds heavily on the tidyverse approach (Wickham et al.
+Our process of cleaning builds heavily on the tidyverse approach (Wickham, Averick, et al.
 2019).
@@
 Section \ref{cleaning} presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and IDA to find and repair anomalies.
-Our final subset is compared the old textbook subset in Section \ref{compare}.
+Our final subset is compared to the old textbook subset in Section \ref{compare}.
 Finally, Section \ref{summary} summarizes the contribution and makes recommendations for the NLSY79 data curators.
@@
 
-M.
-van der Loo and de Jonge (2018) describe the notion of a ``statistical value chain'' where the production stages of data cleaning process are earmarked as raw data (data as arrived to the desk of an analyst), input data (data organised with correct type and identified variables) and valid data (data that faithfully represent the variables).
+van der Loo and de Jonge (2018) describe the notion of a ``statistical value chain'' where the production stages of the data cleaning process are earmarked as raw data (data as arrived to the desk of an analyst), input data (data organised with correct type and identified variables) and valid data (data that faithfully represent the variables).
 In this section, we outline the steps to download the raw data (Section \ref{getdata}) and then tidying the raw data into input data, specifically for the demographic variables (Section \ref{tidydemog}) and the employment variables (Section \ref{tidyemp}), so that the resulting input data can be used downstream for validating the data as described in Section \ref{ida}.
@@
 \item
-  \texttt{NLSY79.csv}: csv format of the response data,
+  \texttt{NLSY79.csv}: Comma Separated Value format of the response data,
 \item
@@
 We only alter the file path in \texttt{NLSY79.R} and run the script without any other alteration.
-This results in an initial processing of the raw data into two data set, \texttt{categories\_qnames} (where the observations are stored in categorical/interval values) and \texttt{new\_data\_qnames} (the observations are stored in integer form).
+This results in an initial processing of the raw data into two data sets, \texttt{categories\_qnames} (where the observations are stored in categorical/interval values) and \texttt{new\_data\_qnames} (the observations are stored in integer form).
 
@@
 According to Wickham (2014), tidy data sets comply with three rules: (i) each variable forms a column, (ii) each observation forms a row, and (iii) each type of observational unit forms a table.
-The raw data does not comply these rules, thus we re-arrange and wrangle the data into tidy data form, columns corresponding to individual ID, year, job number, wage in dollars and the demographic variables.
-This is done using the \texttt{tidyverse} suite of packages (Wickham et al.
-2019), \texttt{tidyr} (Wickham 2020) to pivot the data into long form, \texttt{dplyr} (Wickham et al.
+The raw data does not comply with these rules, thus we re-arrange and wrangle the data into tidy data form, columns corresponding to individual ID, year, job number, wage in dollars and the demographic variables.
+This is done using the \texttt{tidyverse} suite of packages (Wickham, Averick, et al.
+2019), \texttt{tidyr} (Wickham 2020) to pivot the data into long form, \texttt{dplyr} (Wickham, François, et al.
 2020), and \texttt{stringr} (Wickham 2019) for creating new variables, and levels of factors by text wrangling.
@@
 The hourly wages are in the variables beginning with \texttt{HRP} in \texttt{new\_data\_qnames}.
-As a respondent may have multiple jobs, the \texttt{mean\_hourly\_wage} is computed as a weighted average of the hourly wage for each job with the number of hours worked for each job as weights (provided that the information on number of hours is available); if number of hours worked for any job is missing, then the \texttt{mean\_hourly\_wage} is computed as a simple average of all available hourly wage.
+As a respondent may have multiple jobs, the \texttt{mean\_hourly\_wage} is computed as a weighted average of the hourly wage for each job with the number of hours worked for each job as weights (provided that the information on number of hours is available); if number of hours worked for any job is missing, then the \texttt{mean\_hourly\_wage} is computed as a simple average of all available hourly wages.
 Prior to computing the mean hourly wage, we undertook a number of steps to treat unusual observations as described below:
@@
 The employment and demographic variables are then joined.
-This data are further filtered to the cohort who completed the education up to 12th grade and participated at least five rounds in the survey.
+These data are further filtered to the cohort who completed the education up to 12th grade and participated in at least five rounds in the survey.
 We save the resultant wage data on this cohort as \texttt{wages}.
@@
 
-According to Huebner Marianne, Vach, and Cessie (2016), initial data analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
+According to Huebner, Vach, and Cessie (2016), initial data analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
 Moreover, Chatfield (1985) argues that the two main objectives of IDA are data description, which is to assess the structure and the quality of the data, and model formulation without any formal statistical inference.
@@
 
-We take 36 samples randomly from the data and plot them, as shown in Figure \ref{fig:sample-plot}.
-It implies that these respondents have a lot of variability in wages, for example, the IDs in panel numbers 5, 7, and 11.
+We take 20 samples randomly from the data and plot them, as shown in Figure \ref{fig:sample-plot}.
+It shows that these respondents have a lot of variability in wages, for example, the IDs in panel numbers 5, 7, and 11.
 The plot also implies that the samples have a different pattern of mean hourly wages.
@@
 (2001), one of the flaws of the NLSY79 employment data is that the NLSY79 collects the information of the working hours since the last interview.
-Thus, it might be challenging for the respondents to track the within-job hours' changes between survey years, especially for the respondents with fluctuated working hours or seasonal jobs.
+Thus, it might be challenging for the respondents to track the within-job hours' changes between survey years, especially for the respondents with fluctuating working hours or seasonal jobs.
 It even has been more challenging since 1994, where the respondents had to recall two years periods.
@@
 
-To minimize the risk of mistakenly identifying an outlier as an ``erroneous outlier,'' we simulate some thresholds and study how they affect the data.
+To minimize the risk of mistakenly identifying an outlier as an ``erroneous outlier'', we simulate some thresholds and study how they affect the data.
 We find that 0.12 is the most reasonable value to be the threshold to minimize that drawback's risk because it still captures the sensible spikes in the data.
@@
 It implies that the fluctuation can still be observed in the data after the treatment.
-However, the large spikes, which are considered ``erroneous outliers,'' are already eliminated from the data.
+However, the large spikes, which are considered ``erroneous outliers'', are already eliminated from the data.
 Hence, the model produces a data set with a more reasonable degree of fluctuation.
@@
 
-Figure \ref{fig:flow-chart} summarizes the steps taken to go from raw to input to valid data (M. van der Loo and de Jonge 2018) to create a refreshed wages data set.
+Figure \ref{fig:flow-chart-blind} summarizes the steps taken to go from raw to input to valid data (van der Loo and de Jonge 2018) to create a refreshed wages data set.
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/Users/etan0038/yowie/paper/figures/flowchart_blind}
+{\centering \includegraphics[width=0.85\linewidth]{/Users/deeamaliah/Documents/yowie/paper/figures/flowchart_blind}
 
@@
 ``\# of individuals'' means the number of respondents included in each stage, while ``\# of observations'' means the number of rows in the data.
-The color represents the stage of data cleaning in statistical value chain (M.
-P.
-J.
-van der Loo and de Jonge 2021).
+The color represents the stage of data cleaning in statistical value chain (van der Loo and de Jonge 2021).
 Pink, blue, and green represents the raw, input, and valid data, respectively.}\label{fig:flow-chart-blind}
@@
 Now is the time to see how close we have come to the original textbook data.
-The original set contains wages of high school dropouts wages (Singer and Willett 2003) from 1979 through to 1994.
-The refreshed data set for comparison is also wages on high school drop outs, over 1979 to 2018.
+The original set contains wages of high school dropouts (Singer and Willett 2003) from 1979 through to 1994.
+The refreshed data set for comparison is also wages on high school drop outs from 1979 to 2018.
 The original set is available in the R package \texttt{brolgar} and the refreshed data is available in the R package \texttt{[CENSORED]}.
@@
 \caption{Comparison of original textbook example (A) with refreshed data (B).
-The original data was inflation-adjusted to 1990 prices and the individual's time of collection was converted to a length of experience in the workforce, which make it difficult to precisely compare the two sets.}\label{fig:plotting-sw-do}
+The original data was inflation-adjusted to 1990 prices and the individual's time of collection was converted to a length of experience in the workforce, which makes it difficult to precisely compare the two sets.}\label{fig:plotting-sw-do}
 \end{figure}
@@
 There are two aspects of the original data that make a direct comparison difficult.
-The time variable provided is ``experience in the workforce.'' It is not clear how this is calculated.
+The time variable provided is ``experience in the workforce''.
+It is not clear how this is calculated.
 One would expect that there is a record of the day the individual first started a job, and this is used to adjust the year of collection.
@@
 \item
-  calculating experience in the workforce, would require knowing the time of first job within the first year the individual was recorded.
+  calculating experience in the workforce would require knowing the time of first job within the first year the individual was recorded.
 \end{itemize}
@@
 
-The entire analysis is conducted using \texttt{R} (R Core Team 2020) in RStudio IDE using these packages: \texttt{tidyverse} (Wickham et al.
-2019), \texttt{ggplot2} (Wickham 2016), \texttt{dplyr} (Wickham et al.
+The entire analysis is conducted using \texttt{R} (R Core Team 2020) in RStudio IDE using these packages: \texttt{tidyverse} (Wickham, Averick, et al.
+2019), \texttt{ggplot2} (Wickham 2016), \texttt{dplyr} (Wickham, François, et al.
 2020), \texttt{readr} (Wickham and Hester 2020), \texttt{tidyr} (Wickham 2020), \texttt{stringr} (Wickham 2019), \texttt{purrr} (Henry and Wickham 2020), \texttt{brolgar} (Tierney, Cook, and Prvan 2020), \texttt{patchwork} (Pedersen 2020), \texttt{kableExtra} (Zhu 2019), \texttt{MASS} (Venables and Ripley 2002), \texttt{janitor} (Firke 2020), and \texttt{tsibble} (Wang, Cook, and Hyndman 2020).
@@
 \hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
 \leavevmode\hypertarget{ref-nlsy79}{}%
@@
 2021.
-{``National Longitudinal Survey of Youth 1979 Cohort, 1979-2016 (Rounds 1-28).''} Produced and distributed by the Center for Human Resource Research (CHRR), The Ohio State University.
+``National Longitudinal Survey of Youth 1979 Cohort, 1979-2016 (Rounds 1-28).'' Produced and distributed by the Center for Human Resource Research (CHRR), The Ohio State University.
 Columbus, OH.
@@
 \leavevmode\hypertarget{ref-shiny}{}%
-Chang, Winston, Joe Cheng, JJ Allaire, Yihui Xie, and Jonathan McPherson. 2020. \emph{{shiny: Web Application Framework for R}}. \url{https://CRAN.R-project.org/package=shiny}.
+Chang, Winston, Joe Cheng, JJ Allaire, Yihui Xie, and Jonathan McPherson. 2020. \emph{shiny: Web Application Framework for R}. \url{https://CRAN.R-project.org/package=shiny}.
 
 \leavevmode\hypertarget{ref-Chatfield1985TIEo}{}%
-Chatfield, C. 1985. {``The Initial Examination of Data.''} \emph{Journal of the Royal Statistical Society. Series A. General} 148 (3): 214--53.
+Chatfield, C. 1985. ``The Initial Examination of Data.'' \emph{Journal of the Royal Statistical Society. Series A. General} 148 (3): 214--53.
 
@@
 2017.
-{``Using the National Longitudinal Surveys of Youth (NLSY) to Conduct Life Course Analyses.''} In \emph{Handbook of Life Course Health Development}, edited by Richard M.
+``Using the National Longitudinal Surveys of Youth (Nlsy) to Conduct Life Course Analyses.'' In \emph{Handbook of Life Course Health Development}, edited by Richard M.
 Lerner Neal Halfon Christoper B.
@@
 Cham: Springer.
-https://doi.org/\url{https://doi.org/10.1007/978-3-319-47143-3_23}.
+\url{https://doi.org/https://doi.org/10.1007/978-3-319-47143-3_23}.
 
@@
 \leavevmode\hypertarget{ref-janitor}{}%
-Firke, Sam. 2020. \emph{{janitor: Simple Tools for Examining and Cleaning Dirty Data}}. \url{https://CRAN.R-project.org/package=janitor}.
+Firke, Sam. 2020. \emph{janitor: Simple Tools for Examining and Cleaning Dirty Data}. \url{https://CRAN.R-project.org/package=janitor}.
 
 \leavevmode\hypertarget{ref-racismnotrace}{}%
-Fullilove, M. T. 1998. {``Comment: Abandoning ``Race" as a Variable in Public Health Research--an Idea Whose Time Has Come.''} \emph{American Journal of Public Health} 88 (9): 1297--98.
+Fullilove, M. T. 1998. ``Comment: Abandoning ``Race" as a Variable in Public Health Research--an Idea Whose Time Has Come.'' \emph{American Journal of Public Health} 88 (9): 1297--8.
 
@@
 2015.
-{``A Framework for Infusing Authentic Data Experiences Within Statistics Courses.''} \emph{The American Statistician} 69 (4): 307--14.
+``A Framework for Infusing Authentic Data Experiences Within Statistics Courses.'' \emph{The American Statistician} 69 (4): 307--14.
 \url{https://doi.org/10.1080/00031305.2015.1081106}.
@@
 \leavevmode\hypertarget{ref-purrr}{}%
-Henry, Lionel, and Hadley Wickham. 2020. \emph{{purrr: Functional Programming Tools}}. \url{https://CRAN.R-project.org/package=purrr}.
+Henry, Lionel, and Hadley Wickham. 2020. \emph{purrr: Functional Programming Tools}. \url{https://CRAN.R-project.org/package=purrr}.
 
@@
 2016.
-{``A Systematic Approach to Initial Data Analysis Is Good Research Practice.''} \emph{The Journal of Thoracic and Cardiovascular Surgery} 151 (1): 25--27.
+``A Systematic Approach to Initial Data Analysis Is Good Research Practice.'' \emph{The Journal of Thoracic and Cardiovascular Surgery} 151 (1): 25--27.
 
@@
 2020.
-{``Hidden Analyses: A Review of Reporting Practice and Recommendations for More Transparent Reporting of Initial Data Analyses.''} \emph{BMC Medical Research Methodology} 20 (1): 61--61.
+``Hidden Analyses: A Review of Reporting Practice and Recommendations for More Transparent Reporting of Initial Data Analyses.'' \emph{BMC Medical Research Methodology} 20 (1): 61--61.
 
@@
 2018.
-{``The Fivethirtyeight {R} Package: ``Tame Data" Principles for Introductory Statistics and Data Science Courses.''} \emph{Technology Innovations in Statistics Education} 11 (1).
+``The Fivethirtyeight R Package: ``Tame Data" Principles for Introductory Statistics and Data Science Courses.'' \emph{Technology Innovations in Statistics Education} 11 (1).
 \url{https://doi.org/10.5070/T511103589}.
@@
 \leavevmode\hypertarget{ref-KollerManuel2016rARP}{}%
-Koller, Manuel. 2016. {``{robustlmm: An R Package for Robust Estimation of Linear Mixed-Effects Models}.''} \emph{Journal of Statistical Software} 75 (6): 1--24.
+Koller, Manuel. 2016. ``robustlmm: An R Package for Robust Estimation of Linear Mixed-Effects Models.'' \emph{Journal of Statistical Software} 75 (6): 1--24.
 
 \leavevmode\hypertarget{ref-opendata}{}%
-Open Knowledge Foundation. 2021. {``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.''} 2021. \url{http://opendefinition.org/od/2.1/en/}.
+Open Knowledge Foundation. 2021. ``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.'' 2021. \url{http://opendefinition.org/od/2.1/en/}.
 
 \leavevmode\hypertarget{ref-patchwork}{}%
-Pedersen, Thomas Lin. 2020. \emph{{patchwork: The Composer of Plots}}. \url{https://CRAN.R-project.org/package=patchwork}.
+Pedersen, Thomas Lin. 2020. \emph{patchwork: The Composer of Plots}. \url{https://CRAN.R-project.org/package=patchwork}.
 
@@
 2001.
-{``Data Watch: The National Longitudinal Surveys.''} \emph{The Journal of Economic Perspectives} 15 (2): 239--53.
+``Data Watch: The National Longitudinal Surveys.'' \emph{The Journal of Economic Perspectives} 15 (2): 239--53.
 
@@
 \leavevmode\hypertarget{ref-brolgar}{}%
-Tierney, Nicholas, Di Cook, and Tania Prvan. 2020. \emph{{brolgar: BRowse Over Longitudinal data Graphically and Analytically in R}}. \url{https://github.com/njtierney/brolgar}.
+Tierney, Nicholas, Di Cook, and Tania Prvan. 2020. \emph{brolgar: BRowse Over Longitudinal data Graphically and Analytically in R}. \url{https://github.com/njtierney/brolgar}.
 
@@
 \leavevmode\hypertarget{ref-rlm}{}%
-UCLA: Statistical Consulting Group. 2021. {``{Robust Regression \textbar{} R Data Analysis Examples}.''} February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.
+UCLA: Statistical Consulting Group. 2021. ``Robust Regression \textbar{} R Data Analysis Examples.'' February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.
+
+\leavevmode\hypertarget{ref-LooMarkvander2018Sdcw}{}%
+van der Loo, Mark, and Edwin de Jonge. 2018. \emph{Statistical Data Cleaning with Applications in R}.
 
 \leavevmode\hypertarget{ref-validate}{}%
-van der Loo, Mark P. J., and Edwin de Jonge. 2021. {``Data Validation Infrastructure for {R}.''} \emph{Journal of Statistical Software} 97 (10): 1--31. \url{https://doi.org/10.18637/jss.v097.i10}.
-
-\leavevmode\hypertarget{ref-LooMarkvander2018Sdcw}{}%
-van der Loo, Mark, and Edwin de Jonge. 2018. \emph{Statistical Data Cleaning with Applications in r}.
+van der Loo, Mark P. J., and Edwin de Jonge. 2021. ``Data Validation Infrastructure for R.'' \emph{Journal of Statistical Software} 97 (10): 1--31. \url{https://doi.org/10.18637/jss.v097.i10}.
 
 \leavevmode\hypertarget{ref-mass}{}%
-Venables, W. N., and B. D. Ripley. 2002. \emph{Modern Applied Statistics with s}. Fourth. New York: Springer. \url{http://www.stats.ox.ac.uk/pub/MASS4}.
+Venables, W. N., and B. D. Ripley. 2002. \emph{Modern Applied Statistics with S}. Fourth. New York: Springer. \url{http://www.stats.ox.ac.uk/pub/MASS4}.
 
@@
 2020.
-{``A New Tidy Data Structure to Support Exploration and Modeling of Temporal Data.''} \emph{Journal of Computational and Graphical Statistics} 29 (3): 466--78.
+``A New Tidy Data Structure to Support Exploration and Modeling of Temporal Data.'' \emph{Journal of Computational and Graphical Statistics} 29 (3): 466--78.
 \url{https://doi.org/10.1080/10618600.2019.1695624}.
@@
 \leavevmode\hypertarget{ref-plyr}{}%
-Wickham, Hadley. 2011. {``The Split-Apply-Combine Strategy for Data Analysis.''} \emph{Journal of Statistical Software, Articles} 40 (1): 1--29. \url{https://doi.org/10.18637/jss.v040.i01}.
+Wickham, Hadley. 2011. ``The Split-Apply-Combine Strategy for Data Analysis.'' \emph{Journal of Statistical Software, Articles} 40 (1): 1--29. \url{https://doi.org/10.18637/jss.v040.i01}.
 
 \leavevmode\hypertarget{ref-WickhamHadley2014TD}{}%
----------. 2014. {``Tidy Data.''} \emph{Journal of Statistical Software} 59 (10): 1--23.
+---------. 2014. ``Tidy Data.'' \emph{Journal of Statistical Software} 59 (10): 1--23.
 
 \leavevmode\hypertarget{ref-ggplot2}{}%
----------. 2016. \emph{{ggplot2: Elegant Graphics for Data Analysis}}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.
+---------. 2016. \emph{ggplot2: Elegant Graphics for Data Analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.
 
 \leavevmode\hypertarget{ref-stringr}{}%
----------. 2019. \emph{{stringr: Simple, Consistent Wrappers for Common String Operations}}. \url{https://CRAN.R-project.org/package=stringr}.
+---------. 2019. \emph{stringr: Simple, Consistent Wrappers for Common String Operations}. \url{https://CRAN.R-project.org/package=stringr}.
 
 \leavevmode\hypertarget{ref-tidyr}{}%
----------. 2020. \emph{{tidyr: Tidy Messy Data}}. \url{https://CRAN.R-project.org/package=tidyr}.
+---------. 2020. \emph{tidyr: Tidy Messy Data}. \url{https://CRAN.R-project.org/package=tidyr}.
 
@@
 2019.
-{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software} 4 (43): 1686.
+``Welcome to the tidyverse.'' \emph{Journal of Open Source Software} 4 (43): 1686.
 \url{https://doi.org/10.21105/joss.01686}.
@@
 \leavevmode\hypertarget{ref-dplyr}{}%
-Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2020. \emph{{dplyr: A Grammar of Data Manipulation}}. \url{https://CRAN.R-project.org/package=dplyr}.
+Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2020. \emph{dplyr: A Grammar of Data Manipulation}. \url{https://CRAN.R-project.org/package=dplyr}.
 
 \leavevmode\hypertarget{ref-readr}{}%
-Wickham, Hadley, and Jim Hester. 2020. \emph{{readr: Read Rectangular Text Data}}. \url{https://CRAN.R-project.org/package=readr}.
+Wickham, Hadley, and Jim Hester. 2020. \emph{readr: Read Rectangular Text Data}. \url{https://CRAN.R-project.org/package=readr}.
 
@@
 2014.
-{``Knitr: A Comprehensive Tool for Reproducible Research in {R}.''} In \emph{Implementing Reproducible Computational Research}, edited by Victoria Stodden, Friedrich Leisch, and Roger D.
+``Knitr: A Comprehensive Tool for Reproducible Research in R.'' In \emph{Implementing Reproducible Computational Research}, edited by Victoria Stodden, Friedrich Leisch, and Roger D.
 Peng.
@@
 \leavevmode\hypertarget{ref-kableExtra}{}%
-Zhu, Hao. 2019. \emph{{kableExtra: Construct Complex Table with 'kable' and Pipe Syntax}}. \url{https://CRAN.R-project.org/package=kableExtra}.
-
-\end{CSLReferences}
+Zhu, Hao. 2019. \emph{kableExtra: Construct Complex Table with 'kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.
 

### 2022-01-28 paper/paper.tex
Added work experience variable.

@@
 % Pandoc citation processing
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-% for Pandoc 2.8 to 2.10.1
-\newenvironment{cslreferences}%
-  {}%
-  {\par}
-% For Pandoc 2.11+
-\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc} % for calculating minipage widths
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 
-Our process of cleaning builds heavily on the tidyverse approach (Wickham et al.
+Our process of cleaning builds heavily on the tidyverse approach (Wickham, Averick, et al.
 2019).
@@
 
-M.
 van der Loo and de Jonge (2018) describe the notion of a ``statistical value chain'' where the production stages of data cleaning process are earmarked as raw data (data as arrived to the desk of an analyst), input data (data organised with correct type and identified variables) and valid data (data that faithfully represent the variables).
@@
 The raw data does not comply these rules, thus we re-arrange and wrangle the data into tidy data form, columns corresponding to individual ID, year, job number, wage in dollars and the demographic variables.
-This is done using the \texttt{tidyverse} suite of packages (Wickham et al.
-2019), \texttt{tidyr} (Wickham 2020) to pivot the data into long form, \texttt{dplyr} (Wickham et al.
+This is done using the \texttt{tidyverse} suite of packages (Wickham, Averick, et al.
+2019), \texttt{tidyr} (Wickham 2020) to pivot the data into long form, \texttt{dplyr} (Wickham, François, et al.
 2020), and \texttt{stringr} (Wickham 2019) for creating new variables, and levels of factors by text wrangling.

### 2022-01-28 data-raw/NLSY79/NLSY79.R

@@ -1,5 +1,12 @@
-new_data <- read.table(here::here('data-raw/NLSY79/NLSY79.dat'), sep=' ')
-names(new_data) <- c('R0000100',
+
+# Set working directory
+# setwd()
+
+
+new_data <- read.table('data-raw/NLSY79/NLSY79.dat', sep=' ')
+names(new_data) <- c('E0010101',
+  'E0010102',
+  'R0000100',
   'R0000300',
   'R0000500',
   'R0017300',
@@ -17,6 +24,7 @@ names(new_data) <- c('R0000100',
   'R0173600',
   'R0214700',
   'R0214800',
+  'R0215300',
   'R0216700',
   'R0216701',
   'R0229200',
@@ -33,6 +41,7 @@ names(new_data) <- c('R0000100',
   'R0384910',
   'R0406400',
   'R0406401',
+  'R0406700',
   'R0410100',
   'R0410300',
   'R0417400',
@@ -49,6 +58,7 @@ names(new_data) <- c('R0000100',
   'R0599010',
   'R0618900',
   'R0618901',
+  'R0645800',
   'R0664500',
   'R0702510',
   'R0839900',
@@ -61,6 +71,7 @@ names(new_data) <- c('R0000100',
   'R0880310',
   'R0892300',
   'R0893410',
+  'R0897300',
   'R0898200',
   'R0898201',
   'R0905900',
@@ -77,6 +88,7 @@ names(new_data) <- c('R0000100',
   'R1141010',
   'R1145000',
   'R1145001',
+  'R1145800',
   'R1205800',
   'R1256010',
   'R1463000',
@@ -91,6 +103,7 @@ names(new_data) <- c('R0000100',
   'R1516310',
   'R1520200',
   'R1520201',
+  'R1521000',
   'R1605100',
   'R1650810',
   'R1809800',
@@ -105,6 +118,7 @@ names(new_data) <- c('R0000100',
   'R1861510',
   'R1890900',
   'R1890901',
+  'R1891700',
   'R1905600',
   'R1923410',
   'R2171500',
@@ -119,6 +133,7 @@ names(new_data) <- c('R0000100',
   'R2226810',
   'R2258000',
   'R2258001',
+  'R2258800',
   'R2306500',
   'R2318210',
   'R2376300',
@@ -133,6 +148,7 @@ names(new_data) <- c('R0000100',
[…]

### 2022-01-28 data-raw/data-preprocessing.R

@@ -24,23 +24,18 @@ dob_tidy <- new_data_qnames %>%
          year_1981 = `Q1-3_A~Y_1981`,
          month_1979 = `Q1-3_A~M_1979`,
          month_1981 = `Q1-3_A~M_1981`) %>%
-  mutate(dob_year = case_when(
-                      # if the years recorded in both sets match, take 79 data
-                      year_1979 == year_1981 ~ year_1979,
+  mutate(dob_year = case_when(year_1979 == year_1981 ~ year_1979,
                       # if the year in the 81 set is missing, take the 79 data
                       is.na(year_1981) ~ year_1979,
                       # if the sets don't match for dob year, take the 79 data
                       year_1979 != year_1981 ~ year_1979),
-         dob_month = case_when(
-                      # if months recorded in both sets match, take 79 data
-                      month_1979 == month_1981 ~ month_1979,
+         dob_month = case_when(month_1979 == month_1981 ~ month_1979,
                       # if month in 81 set is missing, take the 79 data
                       is.na(month_1981) ~ month_1979,
                       # if sets don't match for dob month, take the 79 data
                       month_1979 != month_1981 ~ month_1979),
          # flag if there is a conflict between dob recorded in 79 and 81
-         dob_conflict = case_when(
-                      !is.na(month_1981) && (month_1979 != month_1981) ~ TRUE,
+         dob_conflict = case_when(!is.na(month_1981) && (month_1979 != month_1981) ~ TRUE,
                       !is.na(year_1981) && (year_1979 != year_1981) ~ TRUE,
                       (year_1979 == year_1981) & (month_1979 == month_1981) ~ FALSE,
                       is.na(month_1981) | is.na(year_1981) ~ FALSE)) %>%
@@ -80,8 +75,8 @@ demog_education <- new_data_qnames %>%
 # Get the highest year completed
 highest_year <- demog_education %>%
   group_by(id) %>%
-  filter(grade == max(grade),
-         year == min(year)) %>%
+  filter(grade == max(grade)) %>%
+  filter(year == min(year)) %>%
   rename(yr_hgc = year,
          hgc_i = grade) %>%
   select(id, yr_hgc, hgc_i) %>%
[…]

### 2022-02-24 R/demog_nlsy79.R
edit the documentation.

@@ -5,7 +5,7 @@
 #'
 #' @source The U.S. Bureau of Labor Statistics. (2021, January 6). *National Longitudinal Survey of Youth 1979*. https://www.nlsinfo.org/content/cohorts/nlsy79/get-data
 #'
-#' @format A data frame contains 12,686 rows and 7 variables:
+#' @format A data frame contains 12,686 rows and 8 variables:
 #' \describe{
 #'  \item{id}{A unique individual's ID number.}
 #'  \item{age_1979}{The age of the subject in 1979.}
@@ -13,6 +13,7 @@
 #'  \item{race}{Race of the subject, NON-BLACK,NON-HISPANIC; HISPANIC; BLACK.}
 #'  \item{hgc}{Highest grade completed.}
 #'  \item{hgc_i}{Integer of highest grade completed.}
+#'  \item{hgc_1979}{The highest grade completed in 1979 (integer value).}
 #'  \item{ged}{Wether the respondent had a high school diploma or Graduate Equivalency Degree (GED). 1: High school diploma; 2: GED; 3: Both}
 #' }
 #'

### 2022-02-24 R/wages.R

@@ -3,13 +3,11 @@
 #' A data set contains longitudinal data of mean hourly wages along with several
 #' demographic variables of Americans from the National Longitudinal Survey of Youth
 #' (NLSY79) held by the U.S. Bureau of Labor Statistics from Round 1 (1979 survey year)
-#' to Round 28 (2018 survey year). The cohort provided in this data set is
-#' the highest grade completed, up to 12th grade, and participated in
-#' at least five rounds of surveys.
+#' to Round 28 (2018 survey year). The cohort provided in this data set of the NLSY79 cohort who participated in at least 3 survey rounds.
 #'
 #' @source The U.S. Bureau of Labor Statistics. (2021, January 6). *National Longitudinal Survey of Youth 1979*. https://www.nlsinfo.org/content/cohorts/nlsy79/get-data
 #'
-#' @format A tsibble with 206,795 rows and 16 variables:
+#' @format A tsibble with 206,795 rows and 18 variables:
 #' \describe{
 #'   \item{id}{A unique individual's ID number. This is the `key` of the data.}
 #'   \item{year}{The year the observation was taken. This could be the `index` of the data.}
@@ -23,8 +21,10 @@
 #'   \item{age_1979}{The age of the subject in 1979.}
 #'   \item{gender}{Gender of the subject, FEMALE and MALE.}
 #'   \item{race}{Race of the subject, NON-BLACK,NON-HISPANIC; HISPANIC; BLACK.}
+#'   \item{grade}{Integer value of the highest grade completed corresponding to `year`.}
 #'   \item{hgc}{Highest grade completed.}
 #'   \item{hgc_i}{Integer of highest grade completed.}
+#'   \item{hgc_1979}{The highest grade completed in 1979 (integer value).}
 #'   \item{ged}{Wether the respondent had a high school diploma or Graduate Equivalency Degree (GED). 1: High school diploma; 2: GED; 3: Both}
 #'   \item{njobs}{Number of jobs that an individual has.}
 #'   \item{hours}{The total number of hours the individual usually works per week.}

### 2022-02-24 R/wages_hs_do.R

@@ -4,13 +4,11 @@
 #' several demographic variables from the National Longitudinal Survey of Youth (NLSY79)
 #' held by the U.S. Bureau of Labor Statistics from Round 1 (1979 survey year)
 #' to Round 28 (2018 survey year). The cohort provided in this data set is high school
-#' dropouts, i.e., the male aged between 14 and 17 years and only completed either
-#' 9th, 10th, 11th grade, or aged at least 19 years old when completed high school and participated
-#' in at least five rounds of surveys.
+#' dropouts in NLSY79 cohort and participated in at least 3 survey rounds.
 #'
 #' @source The U.S. Bureau of Labor Statistics. (2021, January 6). *National Longitudinal Survey of Youth 1979*. https://www.nlsinfo.org/content/cohorts/nlsy79/get-data
 #'
-#' @format A tsibble contains 14,174 rows and 16 variables:
+#' @format A tsibble contains 14,174 rows and 18 variables:
 #'
 #' #' \describe{
 #'   \item{id}{A unique individual's ID number. This is the `key` of the data.}
@@ -25,8 +23,10 @@
 #'   \item{age_1979}{The age of the subject in 1979.}
 #'   \item{gender}{Gender of the subject, FEMALE and MALE.}
 #'   \item{race}{Race of the subject, NON-BLACK,NON-HISPANIC; HISPANIC; BLACK.}
+#'   \item{grade}{Integer value of the highest grade completed corresponding to `year`.}
 #'   \item{hgc}{Highest grade completed.}
 #'   \item{hgc_i}{Integer of highest grade completed.}
+#'   \item{hgc_1979}{The highest grade completed in 1979 (integer value).}
 #'   \item{ged}{Wether the respondent had a high school diploma or Graduate Equivalency Degree (GED). 1: High school diploma; 2: GED; 3: Both}
 #'   \item{njobs}{Number of jobs that an individual has.}
 #'   \item{hours}{The total number of hours the individual usually works per week.}

### 2022-02-24 data-raw/NLSY79/NLSY79.R

@@ -3,11 +3,12 @@
 # setwd()
 
 
-new_data <- read.table('data-raw/NLSY79/NLSY79.dat', sep=' ')
+new_data <- read.table(here::here('data-raw/NLSY79/NLSY79.dat'), sep=' ')
 names(new_data) <- c('E0010102',
   'R0000100',
   'R0000300',
   'R0000500',
+  'R0017300',
   'R0018300',
   'R0047010',
   'R0070500',
@@ -24,6 +25,9 @@ names(new_data) <- c('E0010102',
   'R0214700',
   'R0214800',
   'R0215300',
+  'R0216700',
+  'R0216701',
+  'R0229200',
   'R0230000',
   'R0263710',
   'R0337800',
@@ -36,9 +40,12 @@ names(new_data) <- c('E0010102',
   'R0373410',
   'R0383800',
   'R0384910',
+  'R0406400',
+  'R0406401',
   'R0406700',
   'R0410100',
   'R0410300',
+  'R0417400',
   'R0418200',
   'R0446810',
   'R0545600',
@@ -51,7 +58,10 @@ names(new_data) <- c('E0010102',
   'R0585910',
   'R0598000',
   'R0599010',
+  'R0618900',
+  'R0618901',
   'R0645800',
+  'R0664500',
   'R0665300',
   'R0702510',
   'R0839900',
@@ -65,6 +75,9 @@ names(new_data) <- c('E0010102',
   'R0892300',
   'R0893410',
   'R0897300',
+  'R0898200',
+  'R0898201',
+  'R0905900',
   'R0906700',
   'R0945610',
   'R1087300',
@@ -77,7 +90,10 @@ names(new_data) <- c('E0010102',
   'R1127810',
   'R1140100',
   'R1141010',
+  'R1145000',
+  'R1145001',
   'R1145800',
+  'R1205800',
   'R1206600',
   'R1256010',
   'R1463000',
@@ -90,7 +106,10 @@ names(new_data) <- c('E0010102',
   'R1503210',
   'R1515400',
   'R1516310',
+  'R1520200',
+  'R1520201',
   'R1521000',
+  'R1605100',
   'R1605900',
   'R1650810',
   'R1809800',
@@ -103,7 +122,10 @@ names(new_data) <- c('E0010102',
   'R1848810',
   'R1860600',
   'R1861510',
+  'R1890900',
[…]

### 2022-02-24 data-raw/data-preprocessing.R

@@ -15,6 +15,8 @@ new_data_qnames %>%
   str(list.len = 10)
 
 
+#===== TIDY DEMOGRAPHIC VARIABLES =====
+
 ## ---- dob-tidy
 dob_tidy <- new_data_qnames %>%
   select(id = CASEID_1979,
@@ -53,6 +55,26 @@ demog_tidy <- categories_qnames %>%
          gender = SAMPLE_SEX_1979)
 
 ## ---- demog-ed
+
+# tidy highest grade completed for each round of the survey
+demog_education <- new_data_qnames %>%
+  # in 2018, the variable's name is Q3-4_2018, instead of HGC_2018
+  rename(HGC_2018 = `Q3-4_2018`) %>%
+  select(id = CASEID_1979,
+         starts_with("HGCREV"),
+         "HGC_2012",
+         "HGC_2014",
+         "HGC_2016",
+         "HGC_2018") %>%
+  pivot_longer(!id,
+               names_to = "var",
+               values_to = "grade") %>%
+  separate("var", c("var", "year"), sep = "_") %>%
+  mutate(year = as.numeric(year)) %>%
+  #filter(!is.na(grade)) %>%
+  select(-var)
+
+# get the highest grade completed ever
 hgc_ever <- new_data_qnames %>%
   select(id = CASEID_1979,
          hgc_i = HGC_EVER_XRND) %>%
@@ -69,6 +91,73 @@ hgc_ever <- new_data_qnames %>%
 
 ## ---- tidy-hgc
 
+# Get the highest year completed from the calculation using `grade` variable
+# This aims to check the consistency between grade and hgc_ever in the database
+
+hgc_calc <- demog_education %>%
+  filter(!is.na(grade)) %>%
+  group_by(id) %>%
+  filter(grade == max(grade)) %>%
+  filter(year == min(year)) %>%
+  rename(yr_hgc = year,
+         hgc_i = grade) %>%
+  select(id, yr_hgc, hgc_i) %>%
+  ungroup() %>%
+  mutate(hgc = case_when(hgc_i == 0 ~ "NONE",
+                         hgc_i == 1 ~ "1ST GRADE",
+                         hgc_i == 2 ~ "2ND GRADE",
+                         hgc_i == 3 ~ "3RD GRADE",
+                         hgc_i >= 4 & hgc_i <= 12 ~ paste0(hgc_i, "TH GRADE"),
+                         hgc_i == 13 ~ "1ST YEAR COL",
+                         hgc_i == 14 ~ "2ND YEAR COL",
[…]

### 2022-02-24 data-raw/eda.R

@@ -1,198 +1,117 @@
 ## ---- sw_wages
 
-sw <- brolgar::wages
-
-sw_wages <- sw %>%
-  ggplot(aes(x = xp,
-             y = ln_wages)) +
-  geom_line(aes(group = id), alpha = 0.1) +
-  geom_smooth(se = FALSE, colour = "pink") +
-  labs(tag = "(A)", title = "Original") +
-  theme_bw() +
-  ylab("Hourly wage ($, natural log)") +
-  xlab("Experience in years") +
-  #theme(plot.title = element_text(size = 10)) +
-  ylim(0, 5)
-
-sw_wages_mod <- sw %>%
-  as_tibble() %>%
-  mutate(hgc = ifelse(high_grade < 9, "8TH", "12TH")) %>%
-  mutate(race = case_when(black == 1 ~ "black",
-                          hispanic == 1 ~ "hispanic",
-                          TRUE ~ "white")) %>%
-  mutate(race = factor(race)) %>%
-  ggplot(aes(x = xp,
-             y = ln_wages,
-             linetype = hgc)) +
-  geom_smooth(method = "lm", se = FALSE,
-              colour = "black") +
-  labs(tag = "(A)", title = "Original") +
-  scale_linetype("") +
-  theme_bw() +
-  ylab("Hourly wage ($, natural log)") +
-  xlab("Experience in years") +
-  theme(legend.position = "bottom",
-        legend.direction = "horizontal")
-
-## ---- do_refreshed
-
-# singer-willet did not stated how they calculate experience
-# in database, there is work experience topic
-# (https://www.nlsinfo.org/content/cohorts/nlsy79/topical-guide/employment/work-experience)
-# but there is no variable explicitly mention about experience.
-# in this case, we approximate it with year-1979
-
-# filter do data up to 1994 since the textbook data only covers that period.
-
-do <- yowie::wages_hs_do %>%
-  mutate(lnwage = log(wage)) #%>%
-  #filter(year < 1995)
-
-do_ref <- do %>%
-  ggplot(aes(x = year,
-             y = lnwage)) +
-  geom_line(aes(group = id), alpha = 0.1) +
-  geom_smooth(se = FALSE, colour = "pink") +
-  labs(tag = "(B)", title = "Refreshed data") +
-  theme_bw() +
-  ylab("Hourly wage ($, natural log)") +
-  xlab("Year of data collection") +
[…]

### 2022-02-24 data-raw/ida.R

@@ -256,19 +256,17 @@ wages_clean <- wages_cleaned %>%
 
 # rename and select the wages in tidy
 wages <- wages_clean %>%
-  select(id, year, mean_hourly_wage, age_1979, gender, race, hgc, hgc_i, ged = dip_or_ged,
-                number_of_jobs, total_hours, stwork_year, years_in_workforce, exp, is_wm, is_pred) %>%
+  select(id, year, mean_hourly_wage, age_1979, gender, race, grade, hgc, hgc_i, hgc_1979, ged = dip_or_ged,
+         number_of_jobs, total_hours, stwork, yr_wforce, exp, is_wm, is_pred) %>%
   mutate(id = as.factor(id),
          hgc = as.factor(hgc),
          year = as.integer(year),
-         ged = as.factor(ged),
          age_1979 = as.integer(age_1979),
+         ged = as.factor(ged),
          number_of_jobs = as.integer(number_of_jobs)) %>%
   rename(wage = mean_hourly_wage,
          njobs = number_of_jobs,
-         hours = total_hours,
-         stwork = stwork_year,
-         yr_wforce = years_in_workforce) %>%
+         hours = total_hours) %>%
   as_tsibble(key = id,
              index = year,
              regular = FALSE)

### 2022-06-06 R/adj_inflation.R
Add inflation_adjust function and change the vignette directory.

@@ -1,0 +1,86 @@
+globalVariables("series_id")
+globalVariables("period")
+globalVariables("value")
+globalVariables("year")
+globalVariables("adj_value")
+globalVariables("wage")
+
+#' A Function to get inflation-adjusted wages.
+#'
+#' This function aims to adjust the `wage` column in `wages` and  `wages_hs_do` with the
+#' inflation rate of the base year of your choice.
+#' The inflation rate is calculated from the CPI data published by the US Bureau of Labor Statistics (BLS).
+#' This function needs internet connection to download the data from the BLS website.
+#' The code is adapted from Kris Eberwein in R Bloggers website. https://www.r-bloggers.com/2015/12/calculate-inflation-with-r/
+#' @param base_year The base year of your choice (4 digits-year without quotation mark).
+#'
+#' @return wages and wages_hs_do data with a new column of inflation-adjusted wages value.
+#'
+#' @examples
+#' \dontrun{
+#'    inflation_adjust(1990)
+#' }
+#'
+#' @export
+inflation_adjust <- function(base_year){
+  filter <- dplyr::filter
+  select <- dplyr::select
+  summarise <- dplyr:: summarise
+  as_tsibble <- tsibble::as_tsibble
+  group_by_key <- tsibble::group_by_key
+  index_by <- tsibble::index_by
+
+  if (nchar(base_year) == 4){
+    # Download file from BLS
+    temp <- tempfile()
+    download.file("http://download.bls.gov/pub/time.series/cu/cu.data.1.AllItems",temp)
+    cpi <- read.table(temp,
+                        header=FALSE,
+                        sep="\t",
+                        skip=1,
+                        stringsAsFactors=FALSE,
+                        strip.white=TRUE)
+    colnames(cpi)<-c("series_id", "year", "period", "value", "footnote_codes")
+    unlink(temp)
+    #Get rid of annual time periods and add real dates.
+    cpi <- filter(cpi,series_id == "CUSR0000SA0" &
+                    period !="M13" &
+                    period !="S01" &
+                    period!="S02" &
+                    period!="S03")
+
+    cpi$date <-as.Date(paste(cpi$year, cpi$period,"01",sep="-"),"%Y-M%m-%d")
+
+    cpi <- select(cpi, c(series_id, date, value)) %>%
+      as_tsibble(key = series_id, index = date)
+
[…]

### 2022-06-06 R/utils-pipe.R

@@ -1,0 +1,13 @@
+#' Pipe operator
+#'
+#' See \code{magrittr::\link[magrittr:pipe]{\%>\%}} for details.
+#'
+#' @name %>%
+#' @rdname pipe
+#' @keywords internal
+#' @export
+#' @importFrom magrittr %>%
+#' @usage lhs \%>\% rhs
+#' @importFrom utils download.file
+#' @importFrom utils read.table
+NULL

### 2022-06-07 R/adj_inflation.R
change the inflation adjust function name.

@@ -22,7 +22,7 @@ globalVariables("wage")
 #' }
 #'
 #' @export
-inflation_adjust <- function(base_year){
+adj_inflation <- function(base_year){
   filter <- dplyr::filter
   select <- dplyr::select
   summarise <- dplyr:: summarise
